### 2012-03-02 fgcs/rebuttal.txt
initial files from http://projects.arcs.org.au/svn/podd/trunk/doc/publications/

Firstly, we'd like to thank both reviewers for their insightful comments!

Reviewer 1

When we designed the base ontology, we made careful decisions to only include general concepts in it.

Domain-specific concepts (e.g., those specific to phenomics) are defined in lower-level ontologies that can be incorporated through links from the base ontology.

For example, in the base ontology, we only include generic concepts such as Project, Publication, Investigation, Measurement, Analysis, and so on. Concepts specific to phenomics, such as Genotype, Gene, Sequence, Phenotype, 
are all defined in a separate ontology. This layering/linking of ontologies is discussed in detail in Section 4 with the phenomics ontology as a case study.

The purpose of this architecture is to enable different domains to reuse concepts from the base ontology and extend it with their own vocabularies. Hence, it is the design goal that the system can easily be applied to other domains by developing and plugging in a new domain-specific ontology – little/no redevelopment work will be required to adapt the system to a new domain.

Both reviewers commented on the relevance of Section 5. We think it is very relevant to the architecture and in this revision we have made that connection more obvious. We have also created a new Section 3.3 that discusses the ontology-centric operations: versioning, reasoning and dynamic composition. These operations are central to making the architecture extensible. Their formal definitions and semantics are provided in Section 5.

Domain independence and extensibility are two orthogonal axes of the proposed architecture here. However, they both rely on hierarchical/linked ontologies (Section 4) and ontology-centric operations (Section 3.3 and 5). Domain independence stems from the fact that new ontologies can be easily defined or existing ontologies integrated - by extending the base ontology. Extensibility stems from the fact that concept/object ontologies can be versioned, dynamically composed and reasoned about. Hence, we feel that Section 5 is an integral part of the architecture and deserves a detailed treatment.

Reviewer 2

(1) We added a new subsection, Sec. 4.4, to discuss ways to address the need for both local ontologies and external vocabularies.

(2) Thank you for pointing this out. We have rephrased relevant parts of the paper to provide greater emphasis on our architecture.

(3) We rephrased the discussion of FuGe on p. 6.

(4) A new subsection, Section 3.3, has been created to discuss the essential operations: versioning, ontology reasoning and dynamic composition. The original Figure 6 has been moved to this new section as it is a better place to explain these concepts.

The base ontology is described separately, in Section 4.

(5) More descriptions of the CCLRC Scientific Metadata Model have been added to Section 2.2. Its focus and main drawbacks are also discussed.

ISO 19156 has been described at the end of Section 2.2.

The Observation class has been added to the original Figure 2 (now Figure 3). It was a mistake omitting it. Thank you for picking this up! The missing Protocol class has also been added.

(6) In the Conclusion section, we added 3 paragraphs discussing how Linked Data datasets can be integrated in our ontology-centric architecture as sources of annotation on domain concepts and objects.

(7) We emphasized the significance of Section 5 by linking the new Section 3.3 to Section 5. Section 5 provides the semantic foundations for the operations defined in Section 3.3 - that make the overall architecture extensible and domain-independent. We feel that the semantics are important as they establish the foundation of the work described here. The relevance of Section 5 is also discussed above in response to reviewer 1's comments.

We also clarified the definition of locality in Section 5.3.

All typos have also been corrected. Thank you!

### 2012-03-02 escience2010/latex/podd.tex
initial files from http://projects.arcs.org.au/svn/podd/trunk/doc/publications/

\documentclass[conference,10pt]{IEEEtran}

\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{zed-cm}
\usepackage{graphicx}
\usepackage{dblfloatfix}

%\renewcommand \thefigure{\Roman{figure}}

\begin{document}

\title{PODD - Towards An Extensible, Domain-agnostic Scientific Data Management System}

\author{\IEEEauthorblockN{Yuan-Fang Li, Gavin Kennedy, Faith Davies and Jane Hunter}
\IEEEauthorblockA{School of ITEE, The University of Queensland\\
Brisbane, Australia\\
Email: \{uqyli4,g.kennedy1,f.davies,j.hunter\}@uq.edu.au}
}

\maketitle

\begin{abstract}
Data management has become a critical challenge faced by a wide array of scientific disciplines in which the provision of sound data management is pivotal to the achievements and impact of research projects. Massive and rapidly expanding amounts of data combined with data models that evolve over time contribute to making data management an increasingly challenging task that warrants a rethinking of its design. In this paper we present PODD, an ontology-centric architecture for data management systems that is extensible and domain independent. In this architecture, the behaviors of domain concepts and objects are captured entirely by ontological entities, around which all data management tasks are carried out. The open and semantic nature of ontology languages also makes PODD amenable to greater data reuse and interoperability. To evaluate the PODD architecture, we have applied it to the challenge of managing phenomics data. 
\end{abstract}

\begin{IEEEkeywords}
PODD, data management systems, OWL, ontology-centric architecture, phenomics
\end{IEEEkeywords}

\section{Introduction}\label{sec:intro}
Data management is the practice of managing (digital) data and resources, encompassing a wide range of activities including acquisition, storage, retrieval, discovery, access control, publication, integration, curation and archival. For many data-intensive scientific disciplines such as life sciences and bioinformatics, sound data management informs and enables research and has become an indispensable component~\cite{1107503}.

The need for effective data management is, in a large part, due to the fact that massive amounts of digital data are being generated by modern instruments. Furthermore, the fast evolution of technologies/processes and discovery of new scientific knowledge require flexibility in handling dynamic data and models in data management systems. Among others, there are three core challenges for effective data management in scientific research.

\begin{itemize}
\item The ability to provide a data management service that can manage large quantities of heterogeneous data in multiple formats (text, image, and video) and not be constrained to a finite set of experimental, imaging and measurement platforms or data formats.

\item The ability to support metadata-related services to provide context and structure for data within the data management service to facilitate effective search, query and dissemination.

\item The ability to accommodate evolving and emerging knowledge, technologies and processes.
\end{itemize}

Database systems have traditionally been used successfully to manage research data~\cite{brm2007} in which database schemas are used as domain models to capture attributes and relationships of domain concepts. One implication of the above approach is that domain models need to stay relatively stable as database extension and migration is often an error-prone and laborious task. Consequently, this approach is not suitable for domains where data and model evolution is the norm rather than the exception.

Semantic Web ontology languages such as RDF Schema and OWL possess expressive, rigorously-defined semantics and non-ambiguous syntaxes. Moreover, they have been designed to be open and extensible and to support knowledge and data exchange on the Web scale~\cite{linkeddata,aue07dbpedia}. These intrinsic characteristics make them an ideal conceptual platform on which a flexible scientific data management system can be built.

Ontology language OWL has been widely used in a number of domains, notably in life sciences and biotechnology~\cite{journals/bib/RuttenbergRSM09,citeulike:1882392,citeulike:212874} as a modeling language for its expressivity and extensibility. There is also growing tool support for tasks including reasoning, querying and visualization, making it a viable option for the modeling and representation of scientific domain concepts.

Moreover, with the rapid progression of Semantic Web-based data integration through the community-driven Linked Data project~\cite{citeulike:5008761}, it is advantageous for data management systems to support Semantic Web languages and standards natively to benefit from the rapidly expanding, integrated open datasets.

In this paper, we present our work in designing PODD (Phenomics Ontology Driven Data Management), an extensible, domain-agnostic architecture for scientific data management that uses an ontology-centric approach. In our architecture, we support data and model changes through ontology-based domain modeling. Ontologies are at the core of the system - the behaviors of abstract domain concepts and concrete domain objects are entirely defined by ontological vocabularies. Logical structure of data is therefore maintained and enforced via ontological definitions and reasoning and not via database schemas and associated constraints.

The ontology-based domain model is at the core of PODD as it drives the creation, storage, validation, query and search of data and metadata. In contrast to traditional data management systems that use database schemas as the underlying model, the layered object-oriented approach to developing ontology models and the versioning of ontological definitions make PODD highly extensible.

Based on the ontology-centric architecture, we have developed the PODD repository~\cite{podd_icadl} to meet the above challenges facing the Australian phenomics research community. Our aim is to provide efficient and flexible repository functionalities for large-scale phenomics data, and to provide a mechanism for maintaining structured and precise metadata around the raw data so that they can be stored, distributed and published in a reusable fashion.

We would like to emphasize that although the PODD system is geared towards phenomics research, the ontology-centric architecture we propose in this paper is domain-independent and can be applied in any scientific discipline where research output can be conceptually organized in a structured manner.

The rest of the paper is organized as follows. In Section~\ref{sec:overview} we present related work and give a brief overview of the motivation and goals of the PODD project. Section~\ref{sec:arch} presents the ontology-based architecture for data management systems. In Section~\ref{sec:ont}, we discuss the PODD ontologies in more detail and show how the ontology-based modeling approach is used in the life cycle of repository concepts and objects. In Section~\ref{sec:sys}, we describe the implementation of the PODD data management system and evaluation results to date. Finally, Section~\ref{sec:concl} concludes the paper and identifies future directions.

\section{Related Work, Motivation \& Goals}\label{sec:overview}
Over the years attempts have been made to develop content repository systems and architectures to meet institutional and personal data management needs. In this section, we introduce a number of such systems and architectures. With a survey of related work, we present the motivation behind the ontology-centric architecture and the goals we wish to achieve with the PODD data management system.

\subsection{Scientific Data and Resource Management Systems}
A number of open-source content repository specifications and scientific data management systems have been developed.

Fedora Commons\footnote{\url{http://www.fedora-commons.org/}} is an open-source digital resource management system based on the principles of modularity, interoperability and extensibility. In Fedora Commons, abstract concepts are defined as models, on which inter-relationships and behaviors can be further defined. Data in Fedora Commons repositories are represented as objects, which contain datastreams that store either metadata or data. Fedora Commons makes heavy use of Semantic Web technologies through the use of common RDF vocabularies and the integration with the Mulgara triple store\footnote{\url{http://www.mulgara.org/}}, which can be used for metadata storage and query through SPARQL.

Apache Jackrabbit\footnote{\url{http://jackrabbit.apache.org/}} is an open-source implementation of the Content Repository for Java Technology (JCR) API\footnote{\url{http://jcp.org/en/jsr/detail?id=283}}. In JCR, data is stored in a tree of nodes, which can hold properties of arbitrary values, which is conceptually similar to Fedora Commons. Types can be defined on nodes to place certain restrictions on them.

Fedora Commons and JCR both support fairly basic mechanisms for defining object relationships. Hence, they are usually used as the underlying repository solution on which complex data and document management systems are built. These systems include the National Science Digital Library (NSDL)\footnote{\url{http://nsdl.org/}}, PLoS ONE\footnote{\url{http://www.plosone.org/}}, Biodiversity Heritage Library\footnote{\url{http://www.biodiversitylibrary.org/}} and Fez\footnote{\url{http://fez.library.uq.edu.au/}}, among others. As stated previously, these systems use database schemas as their domain models.

Data management systems have also been developed to support a number of scientific disciplines including high-energy physics, bioinformatics and Earth observation~\cite{652836}.

Bioinformatics Resource Manager (BRM)~\cite{brm2007} is one example of client-server style data management software for bioinformatics research. The client software is installed on users' computers to access (microarray and proteomic) resources stored on BRM server in a PostgreSQL relational database. The BRM server supports data acquisition from external sources such as NCBI~\cite{ncbi} and UniProt~\cite{citeulike:503151}. It also supports annotation using public datasets and connectivity to analytics tools. Data in BRM is stored under the \emph{Project} concept and is mostly flat, i.e., it does not support hierarchical domain concepts such as investigation and publication. SciPort~\cite{1326001} is a peer-to-peer based platform for scientific data integration.

Besides data management systems, grid-based middleware systems have also been developed to provide distributed storage solutions. Such systems include the Storage Resource Broker (SRB)~\cite{783165} and the CERN Data Grid~\cite{652836} and other systems that make use of Globus\footnote{\url{http://www.globus.org/}} middleware. These systems store data in a distributed environment and usually support authentication, replication, redundancy, etc. However, they are mostly only concerned about data storage and replication and hence do not provide full-fledged data management capabilities. Interested readers are referred to~\cite{565296} for a detailed survey of grid resource management systems. Semantic Grid\footnote{\url{http://www.semanticgrid.org/}} is an extension of Grid technology in which rich metadata is made available to and managed explicitly by applications in the grid. A reference architecture for semantic grid, S-OGSA~\cite{citeulike:1021910}, has been proposed that defines a model and capabilities and mechanisms for the Semantic Grid.

More recently, ontology-based approaches have been taken in VIVO~\cite{vivo10} to model, organize and integrate research activities and researcher profile in an institutional setting.

\subsection{Domain Modeling in Scientific Research}
A number of specifications and ontologies have been proposed to model scientific research activities. In 2004, Council for the Central Laboratory of the Research Councils (CCLRC) of UK developed a CCLRC Scientific Metadata Model~\cite{Spallation_metadatamodel} that models scientific activities in free text. An OWL ontology, EXPO, was developed~\cite{citeulike:3735746} to capture metadata about scientific experiments. EXPO was developed in a top-down manner by extending concepts in the Suggested Upper Merged Ontology (SUMO)\footnote{\url{http://www.ontologyportal.org/}}. Although very comprehensive, these models are fairly verbose and not very suitable as models for developing data management systems.

In biological and particularly 'omics research, a large number of databases have been developed to host a variety of information such as genes (Ensembl\footnote{\url{http://www.ensembl.org/}}), proteins (UniProt\footnote{\url{http://www.uniprot.org/}}), publications (PubMed\footnote{\url{http://www.ncbi.nlm.nih.gov/pubmed/}}) and microarray (GEO\footnote{\url{http://www.ncbi.nlm.nih.gov/geo/}}). These databases are generally characterized by the fact that they specialize in a particular kind of data (protein sequences, publications, etc.) and that their conceptual domain models, such as genes~\cite{citeulike:212874} and microarray experiments~\cite{citeulike:151946}, are well understood.

For a scientific data management system to be effective, models of domain concepts need to be integrated with models of scientific activities and workflows. However, models of biological and clinical investigations are less well understood.

The Ontology for Biomedical Investigations (OBI)\footnote{\url{http://purl.obolibrary.org/obo/obi}} is an ongoing effort aimed at developing an integrative ontology for biological and clinical investigations. It takes a top-down approach by reusing high-level, abstract concepts from other ontologies. OBI includes 2,600+ OWL classes and 10,000+ axioms (in the import closure of the OBI ontology). Although OBI is very comprehensive, its size and complexity makes reasoning and querying of OBI-based ontologies and RDF graphs computationally expensive and time consuming, making it impractical as a domain model for data management systems where such reasoning may need to be performed repeatedly.

Functional Genomics Experiment Model (FuGe)~\cite{citeulike:1756058} is an extensible modeling framework for high-throughput functional genomics experiments, aimed at improving the consistency and efficiency of experimental data modeling for the molecular biology research community. Centered around the concept of experiments, it encompasses domain concepts such as \emph{Protocol}, \emph{Sample} and \emph{Data}. FuGe is developed using UML from which XML Schemas and database definitions are derived. The FuGe model not only covers concepts specific to scientific research such as \emph{Analysis}, \emph{Sequence} and \emph{Investigation}; it also defines commonly used concepts such as \emph{Audit}, \emph{Reference} and \emph{Measurement}. Extensions in FuGe are defined through inheritance of UML classes.

We feel that the extensibility we require is not met by FuGe as any addition of new concepts would require amendment of database schemas and code. Moreover, the concrete objects reside in relational databases, making subsequent integration and dissemination more difficult.

\vspace{-8pt}
\subsection{The PODD Repository - Motivation \& Goals}
Phenomics is a fast-growing, data-intensive discipline with new technologies and processes rapidly emerging and evolving. As a result, its domain model and data management systems must also be able to evolve to handle the complexity, dynamics and scale of the data.

In phenomics, data is usually captured by both high- and low-throughput phenotyping devices. The scale of measurement can be from the micro or cellular level, through the level of a single organism, and up to the macro or field level. Imaging, measurement and analysis of organisms on such a large scale will produce an enormous amount of data.

Phenomics research makes use of a large variety of imaging and measurement platforms. For example, in mouse histopathology and organ pathology research, the Zeiss ``Mirax Scan'' scanner is used to scan microscope slides. In clinical pathology, a Flow Cytometer is used to capture laser diffraction images of blood samples. In plant research, the Lemnatec Scanalyzer is used to capture RGB images of plants in growth cabinets. The Fluorogroscan system is used in quenching analysis: the partitioning of light energy used in photosynthesis on model plants such as Arabidopsis. Other devices, such as the Infrared Thermography Camera are used to capture leaf temperature and the SPAD Meter is used to measure the chlorophyll content of plant leaves. New devices and instruments will also be employed as they become available. Moreover, existing instruments may be upgraded so that they can capture more information. The PODD domain model needs to be flexible to accommodate these continual changes in formats, resolution and source of the data.

Because an organism's phenotype is often the product of the organism's genetic makeup, its development stage, disease conditions and its environment, any measurement made against an organism needs to be recorded in the context of these other metadata. Consequently the opportunity exists to create a repository to record the data, the contextual data (metadata) and data classifiers in the form of ontological or structured vocabulary terms. The structured nature of this repository will support both manual and autonomous data discovery as well as provide the infrastructure for data based collaborations with domestic and international research institutions. Currently there are no such integrated systems available. The goals of PODD are to capture, manage, annotate and distribute the data generated by mouse and plant phenomics research activities. Note that currently PODD is designed to be a data management system, it doesn't model scientific workflows.

%\vspace{-6pt}
\section{The Architecture of the Ontology-centric Data Management System}\label{sec:arch}

%\vspace{+4pt}
\subsection{Requirements of Data Management Systems}

For any scientific data management system, a number of requirements need to be satisfied.

\begin{list}{\labelitemi}{\leftmargin=1em}
\item[\textbf{Data storage and management}] Research activities in data-intensive discipline such as 'omics often generate huge amounts of data. The ability to efficiently acquire, store and manage large volumes of data is essential.

\item[\textbf{Data contextualization}] Sufficient contextual information needs to be maintained for more effective organization, understanding and discovery of raw data. Contextual information includes both conceptual domain models, such as how research activities are organized and carried out; and metadata such as provenance information.

\item[\textbf{Data security}] There are many dimensions to data security, including access control and archival. An effective data management system needs to ensure data security through the use of authentication and authorization and sound versioning and backup solutions.

\item[\textbf{Data identification and longevity}] In order to support the dissemination of scientific findings, data in the repository needs to be publicly accessible after being published. Hence, a persistent and unique naming scheme is required. Moreover, valuable scientific data also need to be stored in perpetuity.

\item[\textbf{Data reuse and integration}] Contextual information helps to make sense of raw data. Moreover, it also needs to be made discoverable, through mechanisms such as full-text search, faceted browsing and complex query answering, to allow raw data to be integrated and reused.

\item[\textbf{Model extensibility}] A data management system may need to manage a wide variety of data, which may be generated by different software and captured by different platforms. An expressive and extensible domain model is therefore essential to cater for modification, addition and deletion of domain concepts. The data management system also needs to be designed to minimize service disruption when such a model change occurs.
\end{list}

\subsection{The PODD Ontology-centric Architecture For Data Management Systems}
The most distinguishing characteristic of PODD is the central role that ontologies play. In this architecture, raw data is not stored in a flat structure but is attached to domain objects organized in a logical, hierarchical system, defined according to the domain model that represents the structure of research activities.

Current document management systems such as Fez typically define a relatively static domain model and hardwire it as relational schemas and foreign key constraints in a custom relational database independent from the underlying repository system. Consequently, the information pertinent to each concrete object is stored in this custom database as well. As stated in the previous section, this approach is unsuitable for dynamic environments where conceptual changes are common.

To effectively support a dynamic conceptual framework, the domain model in the proposed architecture is defined using OWL ontologies, in which: OWL classes represent domain concepts; OWL properties define concept attributes and their relationships; OWL restrictions specify constraints on concepts; and finally, OWL individuals define concrete domain objects where attributes and relationships are defined using OWL assertions. Raw data files are attached to concrete domain objects.

Such a conceptual architecture alleviates the problem of imposing hard relational constraints in a database which is difficult to extend/change.

It is worth noting that referential integrity is not sacrificed in achieving flexibility: ontological reasoning involving relevant concepts and objects are performed before object modification to ensure that all constraints are satisfied.

Another drawback of existing systems is that there can be only one domain model. When a concept needs to be updated, all the existing objects need to be updated accordingly, which may be undesirable, inappropriate and time-consuming. This is, unfortunately, unavoidable as long as the domain model is defined using database schemas. In our proposed architecture, as concept and object definitions are stored in the repository, such changes can be versioned so that existing instance objects can remain legitimate when integrity validation is performed as they can still refer to the previous conceptual definitions.

Similarly, the data and metadata (including ontological definitions) of each object can be modified, and the modifications should be versioned so that they can be rolled back.

\vspace{-8pt}
\begin{figure}[htb]
\centering
\includegraphics[trim = 60mm 30mm 50mm 36mm, clip,height=60mm]{architecture.pdf}
\vspace{-20pt} \caption{A high-level view of the ontology-centric architecture.}\label{fig:arch}
\end{figure}

\vspace{-8pt}
\begin{figure*}[htb]
\centering
\includegraphics[trim = 0mm 30mm 0mm 36mm, clip,width=150mm]{top_ont.pdf}
\vspace{-8pt}
\caption{A top-level view of the parent-child relationships between significant domain concepts.}\label{fig:ont}
\end{figure*}

The high-level design of ontology-centric architecture takes a modular and layered approach, as can be seen in Figure~\ref{fig:arch}. At the foundation is the \textbf{data access layer}, consisting of an underlying repository system, an RDF triple store, an in-house database that stores essential information and a full-text search engine. This layer is responsible for low-level tasks when the creation, modification and deletion of concepts and objects occur. The \textbf{business logic layer} in the middle is responsible for managing concepts and objects, such as versioning, object conversion and integrity validation. The \textbf{security layer} controls access (authentication and authorization) to concepts and objects and guards all operations on them. In this architecture, authorization is based on users' access-related attributes, which have two dimensions. Firstly, each user has a system-wide role, such as \emph{registered user} or \emph{system administrator}, which is used to determine access rights across the system. Secondly, a project-wide role, such as \emph{project administrator} and \emph{project observer}, can be assigned to a user so that he can have project-specific access rights. At the top of the stack is the \textbf{interface layer}, where the data management system can be accessed using a number of interfaces such as a Web browser or API calls.

In developing the ontology-centric architecture, the following design decisions have been made to balance expressivity, flexibility and conceptual clarity. These decisions have also been based on a survey of user requirements from scientists within a range of research organizations including the Australian Plant Phenomics Facility (APPF), the Australian Phenomics Network (APN) as well as the Institute of Molecular Biology (IMB), working on collaborative research projects that involve large scale data and distributed teams.

\begin{itemize}
\item There is a top-level domain concept, called Project\footnote{The choice of concept names in the domain ontology is actually irrelevant to the proposed architecture. Names such as Project and others are chosen as they are general and representative enough.}, under which other concepts (such as Investigation and Material) reside in a hierarchical manner.

\item For simplicity reasons, access control (authorization) is defined on the Project level but not on an individual object level, i.e., a given user will have the same access rights for all objects within a given project.
\item Within a Project hierarchy, objects are in a parent-child relationship in a tree structure such that each child can only have one parent. This ensures that access rights are properly propagated from parent to child and there is no chance of confusion.
\item Additionally, inter-object, many-to-many reference relationships can be defined to enhance flexibility of the architecture as it allows arbitrary links between objects to be established.
\item Objects cannot be shared across Projects. Instead, objects must be copied from one project and pasted into another one. Such a rule simplifies object management with the elimination of possible side-effects caused by sharing object between projects.

\item There should be no interference between different versions of a given concept and between objects that are instances of different concept versions.
\end{itemize}

\section{Ontology-based Domain Modeling}\label{sec:ont}
As we emphasized previously, the domain model should be flexible enough to accommodate the rapid changes and dynamic nature of scientific research. In this section, we present the base ontology and discuss the roles it plays in the ontology-centric architecture. It should be noted that the architecture proposed here is domain-independent and it can be applied to any scientific discipline that shares a similar high-level domain model.

\subsection{The Base Domain Ontology}
Inspired by FuGe and OBI, we created the base domain ontology in OWL to define essential concepts, their attributes and inter-relationships in an object-oriented fashion. As stated in the previous section, concepts will be modeled as OWL classes; relationships between concepts and object attributes will be modeled as OWL object and datatype properties. Concrete objects will be modeled as OWL individuals.

For an overview, inter-relationships of some of the concepts in the base ontology are shown in Figure~\ref{fig:ont}. For brevity reasons, only parent-child relationships are shown here. The OWL object properties and cross references between classes are not shown. We also defined the following design principles for the base ontology.

\begin{figure*}[!b]
\vspace{-16pt}
\centering
\includegraphics[trim = 0mm 30mm 0mm 48mm, clip,width=160mm]{podd_ont.pdf}
\vspace{-32pt}
\caption{Extended Domain Ontology for phenomics.}\label{fig:podd_ont}
\end{figure*}

\begin{itemize}
\item All essential concepts are modeled as subclasses of an abstract top-level OWL class \emph{PODDConcept} that captures common attributes and relationships.

\item All relationships between concepts are captured by domain properties, which can be further divided into two property hierarchies, one for parent-child relationships and the other for reference relationships. Each of the two hierarchies have an abstract top-level property, called \emph{contains} and \emph{refersTo}, respectively.

\item All parent-child relationships are modeled in a property hierarchy as sub-properties of the abstract property \emph{contains}, and all reference relationships are modeled in another property hierarchy as sub-properties of the abstract property \emph{refersTo}.

\item For each domain concept $C$, one property is defined in each of the above hierarchies with its range defined to be $C$. The domains of such properties are not specified so that they can be used by any applicable domain concept to establish a relationship between them.

\item Class attributes are modeled using OWL restrictions.

\item Essential domain concepts can be sub-classed to provide more specialized and refined information.

\item To ensure that each object can have at most one parent object, the inverse property of \emph{contains}, \emph{isContainedBy}, is defined so that a max cardinality restriction can be added to the top-level concept \emph{PODDConcept} to enforce it. 
\end{itemize}

The definitions of the top-level constructs are summarized in Figure~\ref{for:concept}, in OWL DL syntax~\cite{hoph03a}.

\vspace{-8pt}
\begin{figure}[htb]
\centering\small
\begin{zed}
PODDConcept \sqsubseteq \top\\
\top \sqsubseteq \forall contains.PODDConcept\\
isContainedBy \sqsubseteq (^-contains)\\
PODDConcept \sqsubseteq \leq 1~ isContainedBy\\
\top \sqsubseteq \forall refersTo PODDConcept
\end{zed}

\vspace{-12pt}
\caption{Top-level ontology constructs in the base ontology.}\label{for:concept}
\end{figure}

\vspace{-8pt}
\subsection{Roles of Domain Ontologies in Object Life Cycle}
The base ontology defines essential concepts independent of the domain. Domain-specific knowledge can be incorporated by extending the base ontology for discipline-specific systems.

As stated in Section~\ref{sec:intro}, the ontology-based domain model is at the center of the whole life cycle of objects. In this subsection, we briefly describe the roles that the domain ontologies perform at various stages of the object life cycle.

\begin{list}{\labelitemi}{\leftmargin=1em}
\item[\textbf{Ingestion}] When an object is created, its definition is expressed in ontological terms. Such definitions will be used to (a) guide the rendering of object creation interfaces and (b) validate the attributes and inter-object relationships the user has entered before the object is ingested. When an object is ingested, its definitions are stored as RDF assertions.

\item[\textbf{Retrieval \& update}] When an object is retrieved from the repository, its attributes and inter-object relationships are retrieved from its RDF assertions, which are used to drive on-screen rendering. When any value is updated, it is validated and updated in this object's RDF assertions.

\item[\textbf{Query \& search}] An object's assertions will be stored in an RDF triple store, which can be queried using the SPARQL query language~\cite{sparql}. Similarly, RDF assertions are indexed to provide functionalities such as full-text search and faceted browsing.

\item[\textbf{Publication \& export}] When an object is published or exported, its metadata, in RDF, will be retrieved and exported.
\end{list}

\subsection{Integration with Existing Domain Ontologies}
As stated before, ontologies such as the Gene Ontology~\cite{citeulike:212874} and Plant Ontology~\cite{citeulike:3008167} are widely used in biomedical research to capture domain knowledge about genes, proteins, sequences and organism phenotypes. In our ontology-centric approach, these ontologies are used to extend our top-level ontology, add annotations on domain objects to enrich their semantic descriptions and enable cross-application reference and integration.

To describe domain knowledge in phenomics, we extend the base ontology by defining additional concepts including \emph{Genotype}, \emph{Gene}, \emph{Phenotype} and \emph{Sequence} as subclasses of \emph{PODDConcept}. Additional OWL object and datatype properties are also defined to model the attributes and relationships of these concepts, as shown in Figure~\ref{fig:podd_ont}. Note that \emph{Phenotype} is a subclass of \emph{Observation}.

\section{The PODD Data Management System}\label{sec:sys}
\subsection{Implementation}
Based on the ontology-centric architecture presented in Section~\ref{sec:arch} and the base ontology presented in Section~\ref{sec:ont} we implemented the PODD data management system - with the aim being to meet the data management challenges faced by the Australian phenomics research community.

In developing the PODD system, we chose to employ a number of mature technologies, as can be seen in Figure V. (1) We use Fedora Commons for the storage and retrieval of domain objects. Together with raw data files, the OWL (for concepts) and RDF (for objects) definitions of each concept and object are stored in a versioned datastream \texttt{PODD}, which is used by the PODD system in various tasks such as object creation, rendering, validation, update and visualization. Moreover, Fedora supports distributed storage through plugins, making it possible to increase the repository storage with demand. (2) We incorporate the Sesame\footnote{\url{http://www.openrdf.org/}} triple store to support complex query answering using SPARQL. Sesame \emph{contexts} are used to give scope to the RDF triples for each domain object. As described in Section III, access control needs to be enforced on a per project basis, which also needs to be enforced for query answering in the triple store. By identifying triples of individual objects, we are able to control contexts a user can access through query expansion. (3) We use the Solr\footnote{\url{http://lucene.apache.org/solr/}} open-source search engine platform to provide full-text search and faceted browsing capabilities. Similar to the structure of the Sesame triple store, there is a one-to-one correspondence between domain objects in the repository and the Solr \emph{documents}, the logical indexing units. (4) Lastly, we use a MySQL database to store user and access control related information, as it is orthogonal to other domain concepts.

\vspace{-8pt}
\begin{figure}[htb]
\centering
\includegraphics[trim = 60mm 30mm 50mm 24mm, clip,height=72mm]{podd_arch.pdf}

\vspace{-8pt} \caption{The architecture and main components of the PODD system.}\label{fig:podd_arch}
\end{figure}

\vspace{-8pt}
Although the architecture and the system are based on ontologies, the interface is designed to hide ontology-related complexity from the user and present information in an easy to use manner for all repository functions. For example, Figure~\ref{fig:browser} shows the browser view of a plant phenomics project that investigates salt tolerance in wheat breeding lines. In this view, the objects are shown in a tree-like structure by following parent-child assertions of subproperties of \emph{contains} defined in  the base and domain ontologies.

\begin{figure}[!t]
\centering
\includegraphics[trim = 0mm 0mm 0mm 0mm, clip,width=88mm]{browser.png}
\vspace{-16pt} \caption{The browser view of a plant project in the PODD repository.}\label{fig:browser}
\end{figure}

%\vspace{-8pt}
\subsection{Preliminary Evaluation}
We have started to deploy the PODD system in Australian phenomics research centers including APPF and APN and begun engaging users in the evaluation of the performance, flexibility, usability and scalability of the system. User feedback to date has shown that the system is intuitive and efficient.

It is well known that native RDF triple stores are not yet as efficient as relational database systems, especially in terms of query performance~\cite{10.1109/ICDE.2009.28,Bizer09theberlin}. In the PODD system, we employ a number of approaches to alleviate this problem. Firstly, we give scope to RDF triples of individual objects so a SPARQL query will only be matched against a (potentially very small) subset of the entire triple store, therefore improving query performance. Secondly, we index all RDF datatype property assertions and RDF type assertions in the Solr index to enable faceted searching and filtering. Together with the use of logical operators in search queries, full-text search can be used effectively to perform complex discovery tasks in most cases.

In summary, following the ontology-centric architecture and through the use of mature technologies, we have successfully developed PODD, a scalable, extensible data management system. At the same time, data management tasks such as versioning, logical organization of data, authentication \& authorization, and discovery, can all be supported effectively. The PODD system demonstrates the feasibility and practicality of the proposed ontology-centric architecture.

\section{Conclusion}\label{sec:concl}
In this work, our contribution to scientific data management is three-fold: firstly, the proposal of the ontology-centric architecture for developing highly extensible and domain-agnostic data management systems that supports the evolution of domain concepts; secondly, the development of a base ontology that defines essential concept in a domain-independent fashion; and thirdly, the development of a phenomics ontology and the data management system (based on both existing and new technologies) that serves as a validation of the feasibility of the proposed architecture.

We have identified a number of future work directions that we would like to pursue. (1) We will investigate the integration with existing domain ontologies such as the Gene Ontology and the Plant Ontology. Potential approaches include using terms defined in these ontologies to annotate metadata objects, and establishing semantic links between concepts/properties through ontology alignment/mapping. (2) We would like to investigate the generalization of the ontology-centric approach and the further abstraction of the PODD base ontology so that the architecture can be applied to other domains including archeology, geological and environmental sciences. (3) Automated population of the PODD system of existing data can improve adoption of the system and alleviate the data entry burden for scientists. We have started working on developing tools for importing existing data in relational databases into PODD. (4) We will continue the development of the PODD system to provide additional functionalities such as data visualization, automated data integration, finer-grained representation of repository objects (file contents and metadata) in RDF, and Linked Data-style data discovery and publication. (5) Last but not least, we will investigate the feasibility of extending the PODD ontology and architecture to support scientific workflow modeling and composition.

\section*{Acknowledgment}
The authors wish to acknowledge the support of the National eResearch Architecture Taskforce (NeAT) and the Integrated Biological Sciences Steering Committee (IBSSC) Australia. The authors also wish to thank Dr. Xavier Sirault, Mr. Philip Wu and Dr. Kai Xu for the discussion and assistance in the development of the PODD ontologies and the PODD repository system.

\bibliography{all}
\bibliographystyle{IEEEtran}

\end{document}

### 2012-03-02 fgcs/podd.tex
initial files from http://projects.arcs.org.au/svn/podd/trunk/doc/publications/

%%
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01
%%
%%
%%
%% $Id: podd.tex 1849 2011-05-20 00:39:13Z liyf@itee.uq.edu.au $
%%
%%
\documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{zed-cm}
\usepackage{graphicx}
\usepackage{dblfloatfix}
\usepackage{boxedminipage}
\usepackage{proof}
%\usepackage{todonotes}

%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

\journal{Future Generation Computer Systems}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{An Ontology-centric Architecture for Extensible Scientific Data Management Systems\tnoteref{t1}}

\tnotetext[t1]{This paper is based on an extended abstract published in the sixth IEEE e-Science Conference, 2010~\cite{podd_escience}.}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author[monash]{Yuan-Fang Li}
\ead{yuanfang.li@monash.edu}

\author[uq,csiro]{Gavin Kennedy}
\ead{Gavin.Kennedy@csiro.au}

\author[uq]{Faith Ngoran}
\ead{f.davies@uq.edu.au}

\author[anu]{Philip Wu}
\ead{philip.wu@anu.edu.au}

\author[uq]{Jane Hunter}
\ead{j.hunter@uq.edu.au}

\address[monash]{Clayton School of IT, Monash University}
\address[uq]{School of ITEE, The University of Queensland}
\address[csiro]{High Resolution Plant Phenomics Centre, Canberra}
\address[anu]{The John Curtin School of Medical Research, Australian National University}

\begin{abstract}
Data management has become a critical challenge faced by a wide array of scientific disciplines in which the provision of sound data management is pivotal to the achievements and impact of research projects. Massive and rapidly expanding amounts of data combined with data models that evolve over time contribute to making data management an increasingly challenging task that warrants a new approach. In this paper we present an ontology-centric architecture for data management systems that is extensible and domain independent. In this architecture, the behaviors of domain concepts and objects are captured entirely by ontological entities, around which all data management tasks are carried out. The open and semantic nature of ontology languages also makes this architecture amenable to greater data reuse and interoperability. To evaluate the proposed architecture, we have applied it to the challenge of managing phenomics data.
\end{abstract}

\begin{keyword}
PODD \sep data management systems \sep OWL \sep ontology-centric architecture \sep phenomics \sep bioinformatics
\end{keyword}
\end{frontmatter}

%%
%% Start line numbering here if you want
%%
% \linenumbers

%% main text
\section{Introduction}\label{sec:intro}
Data management is the practice of managing (digital) data and resources, encompassing a wide range of activities including acquisition, storage, retrieval, discovery, access control, publication, integration, curation and archival. For many data-intensive scientific disciplines such as life sciences and bioinformatics, sound data management informs and enables research and has become an indispensable component~\cite{1107503}.

The need for effective data management is, in a large part, due to the fact that massive amounts of digital data are being generated by modern instruments. Furthermore, the fast evolution of technologies/processes and the discovery of new scientific knowledge require flexibility in handling dynamic data and models in data management systems. Among others, there are three core challenges for effective data management in scientific research.

\begin{itemize}
\item The ability to provide a data management service that can manage large quantities of heterogeneous data in multiple formats (text, image, and video) and not be constrained to a finite set of experimental, imaging and measurement platforms or data formats.

\item The ability to support metadata-related services to provide context and structure for data within the data management service to facilitate effective search, query and dissemination.

\item The ability to accommodate evolving and emerging knowledge, technologies and processes.
\end{itemize}

Database systems have traditionally been used successfully to manage research data~\cite{brm2007} in which database schemas are used as domain models to capture attributes and relationships of domain concepts. One implication of the above approach is that domain models need to stay relatively stable as database extension and migration is often an error-prone and laborious task. Consequently, this approach is not suitable for domains where data and model evolution is the norm rather than the exception.

Semantic Web ontology languages such as RDF Schema and OWL possess expressive, rigorously-defined semantics and non-ambiguous syntaxes. Moreover, they have been designed to be open and extensible and to support knowledge and data exchange on the Web~\cite{linkeddata,aue07dbpedia}. These intrinsic characteristics make them an ideal conceptual platform on which a flexible scientific data management system can be built.

The ontology language OWL is being widely used as a modeling language in a number of domains, notably in life sciences and biotechnology~\cite{journals/bib/RuttenbergRSM09,citeulike:1882392,citeulike:212874}, due to its expressivity and extensibility. There is also an increasing number of tools to support tasks including reasoning, querying and visualization, making it a viable option for the modeling and representation of scientific domain concepts.

Moreover, with the rapid progression of Semantic Web-based data integration through the community-driven Linked Data project~\cite{citeulike:5008761}, it is advantageous for data management systems to support Semantic Web languages and standards natively to benefit from the rapidly expanding, integrated open datasets.

In this paper, we present our work in designing PODD (Phenomics Ontology Driven Data Management), an extensible, domain-agnostic architecture for scientific data management that uses an ontology-centric approach. In our architecture, we support data and model changes through ontology-based domain modeling. Ontologies are at the core of the system - the behaviors of abstract domain concepts and concrete domain objects are entirely defined by ontological vocabularies. Logical structure of data is therefore maintained and enforced via ontological definitions and reasoning and not via database schemas and associated constraints.

The ontology-based domain model is at the core of PODD as it drives the creation, storage, validation, query and search of data and metadata. The object-oriented approach to developing layers of ontology models and the versioning of ontological definitions make PODD highly extensible.

Based on the ontology-centric architecture, we have developed the PODD repository~\cite{podd_icadl} to meet the above challenges facing the Australian phenomics research community. Our aim is to provide efficient and flexible data repository functionality for large-scale phenomics research data, and to provide a mechanism for maintaining structured and precise metadata around the raw data so that the combined data/metadata can be stored, distributed and published in a reusable fashion.

We would like to emphasize that although the PODD system is geared towards phenomics research, the ontology-centric architecture we propose in this paper is domain-independent and can be applied in any scientific discipline where research output can be conceptually organized in a structured manner.

The rest of the paper is organized as follows. In Section~\ref{sec:overview} we present related work and give a brief overview of the motivation and goals of the PODD project. Section~\ref{sec:arch} presents the ontology-based architecture for data management systems. In Section~\ref{sec:ont}, we discuss the PODD ontologies in more detail and show how the ontology-based modeling approach is used in the life cycle of repository concepts and objects. 
%In this architecture, we rely heavily on ontology reasoning for validation and other purposes. 
Ontology versioning and dynamic composition is central to the support of co-evolution of knowledge and data, and to the maintenance of knowledge and data consistency. The semantics of these operations are formally defined in Section~\ref{sec:spec} using first-order logic and set theory. In Section~\ref{sec:sys}, we describe the implementation of the PODD data management system and evaluation results to date. Finally, Section~\ref{sec:concl} concludes the paper and identifies future directions.

\section{Related Work, Motivation \& Goals}\label{sec:overview}
Over the years attempts have been made to develop content repository systems and architectures to meet institutional and personal data management needs. In this section, we introduce a number of such systems and architectures. With this survey of related work, we present the motivation behind the ontology-centric architecture and the goals we wish to achieve with the PODD data management system.

\subsection{Scientific Data and Resource Management Systems \& Tools}
Fedora Commons\footnote{\url{http://www.fedora-commons.org/}} is an open-source digital resource management system based on the principles of modularity, interoperability and extensibility. In Fedora Commons, abstract concepts are defined as models, on which inter-relationships and behaviors can be further defined. Data in Fedora Commons repositories are represented as objects, which contain datastreams that store either metadata or data. Fedora Commons makes heavy use of Semantic Web technologies through the use of common RDF vocabularies and the integration with the Mulgara triple store\footnote{\url{http://www.mulgara.org/}}, which can be used for metadata storage and queried using SPARQL\footnote{\url{http://www.w3.org/TR/rdf-sparql-query/}}.

Apache Jackrabbit\footnote{\url{http://jackrabbit.apache.org/}} is an open-source implementation of the Content Repository for Java Technology (JCR) API\footnote{\url{http://jcp.org/en/jsr/detail?id=283}}. In JCR, data is stored in a tree of nodes, which can hold properties of arbitrary values, which is conceptually similar to Fedora Commons. Types can be defined on nodes to place certain restrictions on them.

Fedora Commons and JCR both support fairly basic mechanisms for defining object relationships. Hence, they are usually used as the underlying repository solution on which complex data and document management systems are built. These systems include the National Science Digital Library (NSDL)\footnote{\url{http://nsdl.org/}}, PLoS ONE\footnote{\url{http://www.plosone.org/}}, Biodiversity Heritage Library\footnote{\url{http://www.biodiversitylibrary.org/}} and Fez\footnote{\url{http://fez.library.uq.edu.au/}}, among others. In contrast to our ontology-centric approach, these systems are reliant upon database schemas as their domain models.

%Data management systems have also been developed to support a number of scientific disciplines including high-energy physics, bioinformatics and Earth observation~\cite{652836}.

Bioinformatics Resource Manager (BRM)~\cite{brm2007} is one example of client-server style data management software for bioinformatics research. The client software is installed on users' computers to access (microarray and proteomic) resources stored on BRM server in a PostgreSQL relational database. The BRM server supports data acquisition from external sources such as NCBI~\cite{ncbi} and UniProt~\cite{citeulike:503151}. It also supports annotation using public datasets and connectivity to analytics tools. Data in BRM is stored under the \emph{Project} concept and is mostly flat, i.e., it does not support hierarchical domain concepts such as investigation and publication. %SciPort~\cite{1326001} is a peer-to-peer based platform for scientific data integration.

ArrayExpress~\cite{citeulike:3501974} is a public database of microarray gene expression data. Its three components, the Repository, the Warehouse and the Atlas, contains a large amount of data about functional genomics experiments, gene expression profiles and summaries and analytical tools for gene expressions across experiments and biological conditions. ArrayExpress is built on top of a number of relational databases. Recently, an Experimental Factor Ontology (EFO), a controlled vocabulary of diseases, multi-species anatomy, compounds and cell-type terms has been developed to support the annotation of data in ArrayExpress. The EFO ontology contains mapping to other ontologies including the Disease Ontology\footnote{\url{http://diseaseontology.sourceforge.net/}}, the NCI Thesaurus\footnote{\url{http://ncit.nci.nih.gov/}} and the Cell Type ontology\footnote{\url{http://lists.sourceforge.net/lists/listinfo/obo-cell-type}}.

Besides data management systems, grid-based middleware systems have also been developed to provide distributed storage solutions. Such systems include the Storage Resource Broker (SRB)~\cite{783165} and the CERN Data Grid~\cite{652836} and other systems that make use of Globus\footnote{\url{http://www.globus.org/}} middleware. These systems store data in a distributed environment and usually support authentication, replication, redundancy, etc. However, they are primarily  concerned with data storage and replication and hence do not provide full-fledged data management capabilities. Interested readers are referred to~\cite{565296} for a detailed survey of grid resource management systems. Semantic Grid\footnote{\url{http://www.semanticgrid.org/}} is an extension of Grid technology in which rich metadata is made available to and managed explicitly by applications in the grid. A reference architecture for semantic grid, S-OGSA~\cite{citeulike:1021910}, has been proposed that defines a model and capabilities and mechanisms for the Semantic Grid.

More recently, ontology-based approaches have been taken in VIVO~\cite{vivo10} to model, organize and integrate research activities and researcher profiles in an institutional setting. DOKMS~\cite{Pirro201038} is a distributed, ontology-based knowledge management framework that serves similar purposes as VIVO. DOKMS operates in a P2P environment with desktop clients instead of browsed-based as in VIVO.

\subsection{Domain Modeling in Scientific Research}
A number of specifications and ontologies have been proposed to model scientific research activities. In 2004, the Council for the Central Laboratory of the Research Councils (CCLRC) of UK developed a CCLRC Scientific Metadata Model~\cite{Spallation_metadatamodel} that models \emph{data holdings} in scientific activities in free text. Similar to what we propose here, the CCLCR Scientific Metadata Model logically organises data into objects, whose  definitions are supplied by UML diagrams. The emphasis of this model is on cataloging data, but not integrating data. Compared to this model, our architecture is based on a language with formal semantics. Our architecture also focuses on extensibility and domain independence.

An OWL ontology, EXPO, was developed~\cite{citeulike:3735746} to capture metadata about scientific experiments. EXPO was developed in a top-down manner by extending concepts in the Suggested Upper Merged Ontology (SUMO)\footnote{\url{http://www.ontologyportal.org/}}. Although very comprehensive, these models are quite verbose and consequently are not very suitable as models for developing data management systems.

For a scientific data management system to be effective, models of domain concepts need to be integrated with models of scientific activities and workflows. In biological and particularly 'omics research, a large number of databases have been developed to host a variety of information such as genes (Ensembl\footnote{\url{http://www.ensembl.org/}}), proteins (UniProt\footnote{\url{http://www.uniprot.org/}}), publications (PubMed\footnote{\url{http://www.ncbi.nlm.nih.gov/pubmed/}}) and microarray (GEO\footnote{\url{http://www.ncbi.nlm.nih.gov/geo/}}). These databases are generally characterized by the fact that they specialize in a particular kind of data (protein sequences, publications, etc.) and that their conceptual domain models, such as genes~\cite{citeulike:212874} and microarray experiments~\cite{citeulike:151946}, are well understood. However, models of biological and clinical investigations are less well understood.

The Ontology for Biomedical Investigations (OBI)\footnote{\url{http://purl.obolibrary.org/obo/obi}} is an ongoing effort aimed at developing an integrative ontology for biological and clinical investigations. It takes a top-down approach by reusing high-level, abstract concepts from other ontologies. OBI includes 2,600+ OWL classes and 10,000+ axioms (in the import closure of the OBI ontology). Although OBI is very comprehensive, its size and complexity makes reasoning and querying of OBI-based ontologies and RDF graphs computationally expensive and time consuming, making it impractical as a domain model for data management systems where such reasoning may need to be performed repeatedly.

The Functional Genomics Experiment Model (FuGe)~\cite{citeulike:1756058} is an extensible modeling framework for high-throughput functional genomics experiments, aimed at improving the consistency and efficiency of experimental data modeling for the molecular biology research community. Centered around the concept of experiments, it encompasses domain concepts such as \emph{Protocol}, \emph{Sample} and \emph{Data}. FuGe is developed using UML from which XML Schemas and database definitions are derived. The FuGe model not only covers concepts specific to scientific research such as \emph{Analysis}, \emph{Sequence} and \emph{Investigation}; it also defines commonly used concepts such as \emph{Audit}, \emph{Reference} and \emph{Measurement}. Extensions in FuGe are defined through inheritance of UML classes.

FuGe provides \emph{extension points} to support the addition of technology-specific data formats~\cite{citeulike:4540763}. The \emph{Audit} package in FuGe provides experimental versioning support to track changes to data objects. However, it has not yet been defined how models can be versioned~\cite{citeulike:4540763}.

Observations and Measurements is an International Standardization effort (ISO/PRF 19156)\footnote{\url{http://www.iso.org/iso/catalogue_detail.htm?csnumber=32574}} under development that defines a conceptual model and encoding for observations and measurements for geographic information systems (GIS). This standard, is a core part of the Open Geospatial Consortium (OGC) Sensor Web Enablement that aims at providing annotations for sensor data. A core component of this standard is the Observation Schema, which defines a set of core features for an observation, including feature of interest, observed property, observation procedure and result. This schema takes a user-centric point of view, emphasizing on the meaning of the observation. An XML Schema\footnote{\url{http://schemas.opengis.net/om/2.0/}} and an OWL implementation\footnote{\url{https://www.seegrid.csiro.au/subversion/xmml/ontologies/Ogc/}} have been developed to formally define this schema.

\subsection{The Phenomics Domain as a Motivation for PODD}
Phenomics is a fast-growing, data-intensive discipline in biology with new technologies and processes rapidly emerging and evolving. As a result, its domain model and data management systems must also be able to evolve to handle the complexity, dynamics and scale of the data.

In phenomics, data is captured by both high- and low-throughput phenotyping devices. The scale of measurement can be from the micro or cellular level, through the level of a single organism, and up to the macro or field level. Imaging, measurement and analysis of organisms on such a large scale produces an enormous amount of data.

In plant phenomics, the objective is to capitalise on emergent technologies to comprehensively measure and study the phenotypes that arise from the plant's genome combined with its development stage, environment and disease factors. New imaging technologies deployed at the High Resolution Plant Phenomics Centre (HRPPC), a node of the Australian Plant Phenomics Facility (APPF), allow for comprehensive imaging and analysis of plant morphology and function. Through visual imaging plant characteristics such as leaf area, biomass, structure and damage can be identified and measured. Far infrared imaging allows profiles of leaf temperature and water usage to be obtained. Near infrared imaging can measure actual water content in plant tissue and soil substrates. FTIR and hyperspectral imaging determine the presence and distribution of chemical compounds at the cellular level. When combined, these platforms can build up a comprehensive digital profile of a plants phenome, especially when recorded over multiple time points or developmental stages.

At the HRPPC many plant lines with distinct genotypes can be analyzed in this way to determine their phenotypes. Information on the environmental conditions and treatments of the plant lines are also recorded and available as subsequent metadata and data to support the primary phenotyping data and analysis. This includes data from controlled environment plant growth rooms and cabinets, sensor data from field trials and automated watering and treatment regimes.

These comprehensive data stacks for a single plant line are then represented in PODD as a combination of metadata and data streams. The PODD domain model describes the data generation platforms (platform), the plant genotypes (genotype), the individual plants that are phenotyped (material), their growth environments (environment) and their treatment regimes (treatment material and treatment). PODD then provides the concept of data objects (data) to wrap data file collections generated by the platforms and to associate with the plant materials. Finally PODD provides analysis objects (analysis) for encapsulating secondary data derived from the analysis of the primary data files.

Similarly, mouse phenomics research makes use of a large variety of imaging and measurement platforms. The Australian Phenomics Network (APN) utilises PODD for the management of primary phenotyping data and metadata. For example, in mouse histopathology and organ pathology based research, the digital slide scanners are used to scan microscope slides, producing high resolution images of stained organ cross-sections. In clinical pathology, a flow cytometer uses laser diffraction to identify and quantify cell types in mouse blood samples and analyse cell types that are the consequence of an ENU mutation. For each of these platforms metadata on the mutagenized mouse lines are recorded in PODD along with visual recordings of the individual mouse specimens, and summary pathology reports.

In both plant and mouse phenomics the scientific workflows are also incorporated into the ontology as process and protocol objects, allowing researchers accessing the online materials to understand the experimental processes that have been used to derive the phenotypes described.

Because an organism's phenotype is often the product of the organism's genetic makeup, its development stage, disease conditions and its environment, any measurement made against an organism needs to be recorded in the context of these other metadata. Consequently the opportunity exists to create a repository to record the data, the contextual data (metadata) and data classifiers in the form of ontological or structured vocabulary terms. The structured nature of this repository will support both manual and autonomous data discovery as well as provide the infrastructure for data based collaborations with domestic and international research institutions. Currently there are no such integrated systems available. The goals of PODD are to capture, manage, annotate and distribute the data generated by mouse and plant phenomics research activities.

%\vspace{-6pt}
\section{The Architecture of the Ontology-centric Data Management System}\label{sec:arch}

%\vspace{+4pt}
\subsection{Requirements of Data Management Systems}

For any scientific data management system, a number of requirements need to be satisfied.

\begin{list}{\labelitemi}{\leftmargin=30pt}
\item[\textbf{Data storage and management}] Research activities in data-intensive disciplines such as 'omics often generate huge amounts of data. The ability to efficiently acquire, store and manage large volumes of data is essential.

\item[\textbf{Data contextualization}] Sufficient contextual information needs to be maintained for more effective organization, understanding and discovery of raw data. Contextual information includes both conceptual domain models, such as how research activities are organized and carried out; and metadata such as provenance information.

\item[\textbf{Data security}] There are many dimensions to data security, including access control and archival. An effective data management system needs to ensure data security through the use of authentication and authorization and sound versioning and backup solutions.

\item[\textbf{Data identification and longevity}] In order to support the dissemination of scientific findings, data in the repository needs to be publicly accessible after being published. Hence, a persistent and unique naming scheme is required. Moreover, valuable scientific data also need to be stored in perpetuity.

\item[\textbf{Data reuse and integration}] Contextual information helps to make sense of raw data. Moreover, it also needs to be made discoverable, through mechanisms such as full-text search, faceted browsing and complex query answering, to allow raw data to be integrated and reused.

\item[\textbf{Model extensibility}] A data management system may need to manage a wide variety of data, which may be generated by different software and captured by different platforms. An expressive and extensible domain model is therefore essential to cater for modification, addition and deletion of domain concepts. The data management system also needs to be designed to minimize service disruption when such a model change occurs.
\end{list}

\subsection{The Ontology-centric Architecture For Data Management Systems}\label{sec:arch_desc}
With the data management requirements identified, we design an ontology-centric architecture that satisfies these requirements effectively and efficiently. The most distinguishing characteristic of this architecture is the central role that ontologies play. In our architecture, raw data is not stored in a flat structure but is attached to domain objects organized in a logical, hierarchical system, defined according to the domain model that represents the structure of research activities.

Current document management systems such as Fez typically define a relatively static domain model and hardwire it as relational schemas and foreign key constraints in a custom relational database independent from the underlying repository system. Consequently, the information pertinent to each concrete object is stored in this custom database as well. As stated in the previous section, we find this approach less flexible for dynamic environments where conceptual changes are common.

To effectively support a dynamic conceptual framework, the domain model in the proposed architecture is defined using OWL ontologies, in which: OWL classes represent domain concepts; OWL properties define concept attributes and their relationships; OWL restrictions specify constraints on concepts; and finally, OWL individuals define concrete domain objects where attributes and relationships are defined using OWL assertions. Raw data files are attached to concrete domain objects.

Such a conceptual architecture alleviates the problem of imposing hard relational constraints in a database which is difficult to extend/change.

It is worth noting that referential integrity is not sacrificed in achieving flexibility: ontological reasoning involving relevant concepts and objects is performed before object modification to ensure that all constraints are satisfied.

%Another drawback of existing systems is that there can be only one domain model. When a concept needs to be updated, all the existing objects need to be updated accordingly, which may be undesirable, inappropriate and time-consuming. This is, unfortunately, unavoidable as long as the domain model is defined using database schemas.

Compared to existing relational data management frameworks, concept changes are naturally supported in our ontology-centric architecture. In our proposed architecture, concepts (schema) and objects (data) are organic entities and their definitions can evolve independently over time. Changes to concept and object definitions are versioned to maintain a complete history. As a result, existing instance objects can remain legitimate when integrity validation is performed as they can still refer to the version of conceptual definition. Details of such operations will be discussed in Section~\ref{sec:spec}.

%\vspace{-8pt}
\begin{figure}[htb]
\centering
\includegraphics[trim = 60mm 30mm 50mm 36mm, clip,height=60mm]{architecture.pdf}
\vspace{-20pt} \caption{A high-level view of the ontology-centric architecture.}\label{fig:arch}
\end{figure}

The high-level design of ontology-centric architecture takes a modular and layered approach, as can be seen in Figure~\ref{fig:arch}. At the foundation is the \textbf{data access layer}, consisting of an underlying repository system, an RDF triple store, a relational database that stores system and user-related information and a full-text search engine. This layer is responsible for low-level tasks supporting the creation, modification and deletion of concepts and objects. The \textbf{business logic layer} in the middle is responsible for managing concepts and objects, such as versioning, object composition and integrity validation. These operations will be further discussed in Section~\ref{sec:op}. The \textbf{security layer} controls access (authentication and authorization) to concepts and objects and guards all operations on them.

In this architecture, authorization is based on users' access-related attributes, which have two dimensions. Firstly, each user has a system-wide role, such as \emph{registered user} or \emph{system administrator}, which is used to determine access rights across the system. Secondly, a project-wide role, such as \emph{project administrator} and \emph{project observer}, can be assigned so a user can have project-specific access rights. At the top of the stack is the \textbf{interface layer}, where the data management system can be accessed using a number of interfaces such as a Web browser or API calls.

In developing the ontology-centric architecture, the following design decisions have been made to balance expressivity, flexibility and conceptual clarity. These decisions have been based on a survey of user requirements from scientists within a range of research organizations including the Australian Plant Phenomics Facility (APPF) as well as the Australian Phenomics Network (APN), working on collaborative research projects that involve large scale data and distributed teams.

\begin{itemize}
\item There is a top-level domain concept, called $Project$\footnote{The choice of concept names in the domain ontology is actually irrelevant to the proposed architecture - names such as $Project$ and others are chosen as they are general and representative for a large number of scientific disciplines.}, under which other concepts (such as $Investigation$ and $Material$) reside in a hierarchical manner.

\item Access control (authorization) is defined on the $Project$ level rather than on an individual object level, i.e., a given user will have the same access rights for all objects within a given project.
\item Within a $Project$ hierarchy, objects are in a parent-child relationship in a tree structure such that each child can only have one parent. This ensures that access rights are properly propagated from parent to child and there is no chance of confusion.
\item Additionally, inter-object, many-to-many reference relationships can be defined to enhance flexibility of the architecture as it allows cross referencing between objects to be established.
\item Objects cannot be shared across $Project$s. Instead, objects must be copied from one project and pasted into another one. Such a rule simplifies object management with the elimination of possible side-effects caused by sharing object between projects.

\item There cannot be interference between different versions of a given concept and between objects that are instances of different concept versions.
\end{itemize}

\subsection{Ontology-centric Operations}\label{sec:op}
Versioning, dynamic composition of domain concepts and objects, and ontological reasoning are important operations in the \emph{business logic layer} in Figure~\ref{fig:arch}.  Applied together, these techniques contribute to making this architecture extensible. In this subsection, we describe how these operations are performed on domain concepts and objects (we refer to them as \emph{entities} thereafter). The semantics of these operations are further described in Section~\ref{sec:spec}.

\textbf{Versioning.} As stated previously, all entities are semantically described using ontologies. Each entity (concept and object) maintains a sequence of \emph{versions} of ontologies, from the earliest to the latest, as its semantic definition. Each version contains a unique timestamp and an ontology. A version is created whenever the entity is updated. These versioned ontologies provide a complete change history for each and every entity.
Domain entities refer to each other in their definitions, hence the cross references between versions of ontologies from different entities are also maintained.

Versioning information is maintained in ontologies with the use of OWL \emph{annotation properties} \texttt{poddModel:versionInfo} and \texttt{poddModel:mapsToVersion}. \texttt{poddModel:versionInfo} annotates an ontology with its own timestamp. \texttt{poddModel:mapsToVersion} annotates the cross references between OWL classes and individuals with the timestamp of the target of the cross reference. Semantics of these annotation properties are further defined in Section~\ref{sec:versioning}.

\textbf{Ontological reasoning.} Before entities can be ingested or updated, their definitions need to be checked to ensure semantic integrity. As ontologies serve as entity definitions, ontological reasoning is performed for integrity checking. Different aspects of integrity are verified, including type correctness and cardinality of entity attributes and associations between entities.

In our architecture, as described in the previous subsection, a domain concept is represented as an OWL class, and a domain object is represented as an OWL individual. Hence, the problem of integrity checking of the concept/object reduces to the verification of consistency of the corresponding OWL class/individual. Such verification tasks are performed by OWL reasoners automatically.

Verification of OWL classes/individuals cannot be performed in isolation as they refer to each other in their definitions. All relevant ontological definitions need to be brought together to form \emph{transient} ontologies. The consistency of such ontologies is verified before the consistency of the OWL class/individual is verified. This is because an inconsistent ontology makes any inference logically invalid.

\textbf{Dynamic composition.} Transient ontologies are formed by dynamically composing OWL classes and objects in runtime. Dynamic composition is performed in the following scenarios.

\begin{itemize}
\item Before a new object is created, concept ontologies are composed to generate a \emph{template} to drive the rendering of the user input page in HTML. In this case, the latest versions of ontologies of all involved concepts are used for the composition.

\item Similarly, before a new object/concept is saved, the latest versions of relevant concept/object ontologies are composed to verify the consistency of user inputs. The concept may reference other concepts, the corresponding versions (which may not be the latest) of other concepts' ontologies are used in the composition. Effectively, the \emph{closure} of concept ontologies are obtained in the composition.

\item When an existing object is updated, the updated OWL individual ontology is checked against the latest version of the corresponding concept. For other cross referenced objects, the corresponding versions of their ontologies are used in the composition. 
\end{itemize}

From the above discussion it can be seen that reasoning and composition are applied together to maintain entity consistency, and that reasoning tasks are performed on different levels. We refer to these levels as \emph{localities} in Section~\ref{sec:locality}. Initial concept definitions are extracted from base and domain ontologies (described in the next Section). This \emph{bootstrapping} process is further described in Section~\ref{sec:bootstrap}.

The interactions among these operations are illustrated in the example in Figure~\ref{fig:versions} below. As can be seen here, each of the concepts $C$, $D$ and objects $G$, $H$ and $K$ has a sequence of versions of ontologies ($O_i$'s, $P_j$'s, $Q_k$'s, etc.) and each of these ontologies has a corresponding timestamp ($t_i$'s, $u_j$'s and $v_k$'s).

Concepts can refer to particular versions of each other's ontologies, as illustrated in Figure~\ref{fig:versions} between concepts $C$ and $D$. Each version of the ontology of an object maps to the corresponding version of the ontology in its concept's definition. For example, ontology version $v_{2}$ of object $G$ refers to version $u_{2}$ in its defining concept $C$. Ontology version $t_{1}$ in object $H$ refers to version $u_{1}$ in the same concept. Hence, they can evolve independently without interference. Similarly, objects can refer to each other's ontologies as well, and the versions of such cross references are maintained as well. Different versions of ontologies of object $K$ refer to different versions of ontologies of objects $G$ and $H$, respectively.

\begin{figure*}[htb]
\centering
\includegraphics[height=90mm]{versions.pdf}

\vspace{-8pt} \caption{A graphical representation of ontology versions and their relationship.}\label{fig:versions}
\end{figure*}

The above three operations are essential to making the proposed architecture extensible and domain-independent. The semantics of these operations are further specified in Section~\ref{sec:spec}.

\section{Ontology-based Domain Modeling}\label{sec:ont}
Domain concepts and objects are described using ontological terms. These terms are defined in the base and domain-specific ontologies that will be described in this section. The \emph{bootstrapping} of concepts using these system ontologies will be further described in Section~\ref{sec:bootstrap}.

As we emphasized previously, the domain model should be flexible enough to accommodate the rapid changes and the dynamic nature of scientific research. In the following subsection, we present the base ontology that we have implemented that defines general vocabularies that are common to many data-intensive disciplines. In Section~\ref{sec:phe_ont}, we present how the base ontology is extended to support phenomics research.

In our modeling approach, we define ontologies on multiple levels. On the base level, the ontology defines domain-independent concepts and their inter-relationships with each other. Based on this base ontology, other more domain-specific ontologies can be developed hierarchically.

\subsection{The Base Domain Ontology}
Inspired by FuGe and OBI, we create the base domain ontology in OWL to define essential concepts in scientific investigations, their attributes and inter-relationships in an object-oriented fashion. As stated in the previous section, concepts are modeled as OWL classes; relationships between concepts and object attributes are modeled as OWL object- and datatype-property assertions. Concrete objects are modeled as OWL individuals.

We define the following design principles for the base ontology.

\begin{itemize}
\item All essential concepts are modeled as subclasses of an abstract top-level OWL class \emph{PODDConcept} that captures common attributes and relationships.

\item All relationships between concepts are captured by domain properties, which can be further divided into two property hierarchies, one for parent-child relationships and the other for reference relationships. Each of the two hierarchies have an abstract top-level property, called \emph{contains} and \emph{refersTo}, respectively.

\item All parent-child relationships are modeled in a property hierarchy as sub-properties of the abstract property \emph{contains}, and all reference relationships are modeled in another property hierarchy as sub-properties of the abstract property \emph{refersTo}.

\item For each domain concept $C$, one property is defined in each of the above hierarchies with its range defined to be $C$. The domains of such properties are not specified so that they can be used by any applicable domain concept to establish a relationship between them.

\item Class attributes are modeled using OWL restrictions.

\item Essential domain concepts can be sub-classed to provide more specialized and refined information.

\item To ensure that each object can have at most one parent object, the inverse property of \emph{contains}, \emph{isContainedBy}, is defined so that a max cardinality restriction can be added to the top-level concept \emph{PODDConcept} to enforce it.

\item We use annotation properties on OWL class definitions and axioms to describe their non-semantic properties. For example, for each OWL subclass axiom, we assign a $weight$ annotation property value to suggest their relative placement in the browser.
\end{itemize}

For an overview, inter-relationships of some of the concepts in the base ontology are shown in Figure~\ref{fig:ont}. For brevity reasons, only parent-child relationships are shown here. The OWL object properties and cross references between classes are not shown.

\begin{figure*}[htb]
\centering
\includegraphics[width=\textwidth]{top_ont.pdf}
\vspace{-16pt}
\caption{A top-level view of the parent-child relationships between essential concepts.}\label{fig:ont}
\end{figure*}

Some definitions of the top-level constructs are summarized in Figure~\ref{for:concept} in OWL DL syntax.

On the left-hand side we define the OWL class $PODDConcept$, which is the super class of all the domain concepts in our ontologies. It also defines the property $contains$ and its inverse $isContainedBy$. The expression $\leq 1~ isContainedBy$ ensures that each instance of $PODDConcept$ can only be contained by at most one other object, hence maintaining the single-parent structure in the $contains$ hierarchy.

On the right-hand side we define the concept $Project$ to be a subclass of the conjunction of 4 anonymous class expressions, stating that project must have exactly one project plan, at least one investigation, exactly one start date, and at most one publication date (a project may not have been published).

\vspace{-8pt}
\begin{figure}[htb]
\centering\small
\begin{minipage}[t]{.4\textwidth}
\begin{align*}
&PODDConcept \sqsubseteq \top\\
&\top \sqsubseteq \forall contains.PODDConcept\\
&isContainedBy \sqsubseteq (^-contains)\\
&PODDConcept \sqsubseteq~ \leq 1~ isContainedBy
%&\top \sqsubseteq \forall refersTo PODDConcept\\
%&referencedBy\sqsubseteq (^- refersTo)
\end{align*}
\end{minipage}
\begin{minipage}[t]{.4\textwidth}
\begin{align*}
Project \sqsubseteq~ =~ &1~ hasProjectPlan~\sqcap\\
        & \geq~ 1~ hasInvestigation~\sqcap\\
        &\sqsubseteq~ =~ 1~ hasStartDate~\sqcap\\
        &\sqsubseteq~ \leq~ 1~ hasPublicationDate
\end{align*}
\end{minipage}
%\vspace{-4pt}
\caption{Top-level ontology constructs in the base ontology.}\label{for:concept}
\end{figure}

\subsection{Domain-specific Ontologies}\label{sec:phe_ont}
%As stated before, ontologies such as the Gene Ontology~\cite{citeulike:212874} and Plant Ontology~\cite{citeulike:3008167} are widely used in biomedical research to capture domain knowledge about genes, proteins, sequences and organism phenotypes. In our ontology-centric approach, these ontologies can be used to extend our top-level ontology, add annotations on domain objects to enrich their semantic descriptions and enable cross-application reference and integration.

To describe domain knowledge in phenomics, we extend the base ontology and develop a phenomics ontology by defining additional concepts including \emph{Genotype}, \emph{Gene}, \emph{Phenotype} and \emph{Sequence} as subclasses of \emph{PODDConcept}. Additional OWL object- and datatype-properties are also defined to model the attributes and relationships of these concepts, as shown in Figure~\ref{fig:podd_ont}. Note that \emph{Phenotype} is a subclass of \emph{Observation}. To cater for different domains in phenomics, we can further specialize the phenomics ontology to one for the plant domain and one for the mouse domain.

\begin{figure}[htb]
%\vspace{-16pt}
\centering
\includegraphics[width=\textwidth]{podd_ont.pdf}
\vspace{-16pt}
\caption{Essential classes in the domain-specific ontology for phenomics.}\label{fig:podd_ont}
\end{figure}

Some important new OWL classes can be seen in Figure~\ref{fig:phe_ont}. Note that in the phenomics ontology we reuse definitions (e.g., $Observation$) from the base ontology to define new concepts. We also refine definitions (e.g., $Project$ and $Material$) in the base ontology with additional restrictions in the phenomics ontology.

\begin{figure}[htb]
\small
\begin{minipage}[t]{.45\textwidth}
\centering
\begin{align*}
Genotype &\sqsubseteq~ PODDConcept\\
         &\sqsubseteq~ \forall hasGene.Gene\\
         &\sqsubseteq~ \leq~ 1~ hasEcotype\\
         &\sqsubseteq~ \leq~ 1~ hasSubspecies
\end{align*}
\end{minipage}
\begin{minipage}[t]{.4\textwidth}
\small\centering
\begin{align*}
Phenotype & \sqsubseteq Observation\\
Project & \sqsubseteq \forall hasGenotype.Genotype\\
Material & \sqsubseteq \forall hasPhenotype.Phenotype~\sqcap\\
         &\quad~ \forall refersToGenotype.Genotype
\end{align*}
\end{minipage}
\caption{Domain-specific OWL definitions.}\label{fig:phe_ont}
\end{figure}

\subsection{Roles of Domain Ontologies in Object Life Cycle}
The base ontology defines essential concepts in a domain-independent fashion. Domain-specific knowledge can be incorporated by extending the base ontology for discipline-specific systems.

As stated in Section~\ref{sec:intro}, the ontology-based domain model is at the center of the whole life cycle of objects. In this subsection, we briefly describe the roles that the domain ontologies perform at various stages of the object life cycle.

\begin{list}{\labelitemi}{\leftmargin=30pt}
\item[\textbf{Ingestion}] When an object is created, its definition is expressed in ontological terms. Such definitions will be used to (a) guide the rendering of object creation interfaces and (b) validate the attributes and inter-object relationships the user has entered before the object is ingested. When an object is ingested, its definitions are stored as RDF assertions.

\item[\textbf{Retrieval \& update}] When an object is retrieved from the repository, its attributes and inter-object relationships are retrieved from its RDF assertions, which are used to drive on-screen rendering. When any value is updated, it is validated and updated in this object's RDF assertions.

\item[\textbf{Query \& search}] An object's assertions will be stored in an RDF triple store, which can be queried using the SPARQL query language. Similarly, RDF assertions are indexed to provide functionalities such as full-text search and faceted browsing.

\item[\textbf{Publication \& export}] When an object is published or exported, its metadata, in RDF, will be retrieved and exported.
\end{list}

\subsection{Vocabulary Reuse}
Ontologies in the Semantic Web are designed to be shared, reusable and reused. A large number of ontologies and associated concrete RDF data have been developed in the Linked Data project~\cite{citeulike:5008761}. Hence, it is beneficial, and advisable to reuse vocabularies whenever possible.

In our architecture, external vocabularies are used as \emph{annotations} on concepts and objects while local ontologies define their basic semantics. Such an approach maintains semantic integrity of the domain while maximizing knowledge reuse and integration.

\section{Knowledge \& Data Co-evolution: Dynamic Ontology Composition \& Versioning}\label{sec:spec}

As stated previously, high extensibility is an important design goal for our ontology-centric architecture. A prerequisite to this goal is that the knowledge and the data described by the knowledge are able to evolve independently. Moreover, to ensure consistency, it is also necessary to perform integrity consistency checks on certain operations that modify the state of the system, knowledge or data.

The architecture proposed here achieves these two goals through the combination of concepts and objects \emph{versioning} and \emph{dynamic composition}. In this section, we formally define their definitions and show how we employ them in a declarative fashion.

%\subsection{Concepts, objects \& their ontological definitions}
As can be seen in Figures~\ref{fig:ont} - \ref{fig:phe_ont}, the semantics of the domain is described by a number of \emph{concepts}. Concrete \emph{objects}, defined by the concepts, represent real-world metadata entities that encapsulate raw data. The attributes and relationships between concepts and objects are entirely defined by ontologies\footnote{Note that the interpretation of \emph{ontology} in this section is slightly relaxed to include both OWL and RDF definitions.}. The versioning, reasoning and composition operations are defined over these ontologies. In Section~\ref{sec:op} we introduced these operations. In this section, we give formal definitions and present their semantics.

\subsection{Notations}
Here we define the notations that will be used in the rest of this section.
\begin{itemize}
\item Concepts and objects: $A$, $B$, $C$,
%\item Definitions of concepts and objects: $\mathcal{O}$, $\mathcal{P}$, $\mathcal{Q}$,
\item The ontology datatype: $\bbold{O}$, and individual ontologies: $O$, $P$, $Q$, and 
\item The time datatype: $\bbold{T}$, and individual timestamps: $t_1$, $t_2$, $t_3$.
\end{itemize}

\subsection{Ontology Versioning}\label{sec:versioning}
A concept (or object) is defined by a \emph{sequence} of ontologies, ordered by their creation timestamps. Initially the sequence contains only one item, the initial ontology definition for the concept (or object).

Formally, for any given concept $C$, we define its definition $\mathcal{O}_C$ as a non-empty partial function from time $\bbold{T}$ to ontologies $\bbold{O}$.

\begin{align}
\mathcal{O}_C &\colon \bbold{T} \pfun \bbold{O} \mathrm{,~where~~} \#\mathcal{O}_C \neq 0\label{eqn:version}
\end{align}

In the above definition, the expression $\#\mathcal{O}_C$ denotes the cardinality of the function $\mathcal{O}_C$. The condition that the cardinality is not 0 ensures that $C$'s definition is non-empty. By this definition, $\bbold{O}_C$, the definition of $C$, is a set of pairs $(time, ontology)$: $\{(t_1, O_1),(t_2,O_2),\ldots,(t_n,O_n)\}$, such that $t_i$'s are unique from each other. Hence, this definition implicitly defines a temporal sequence of ontologies.

Whenever a concept definition is updated, a new ontology is created and added as the latest version to the sequence $\mathcal{O}_C$. For example, assuming that the current sequence is $\{(t_1, O_1),(t_2,O_2),\ldots,(t_n,O_n)\}$, a modification will add a new ontology $O_{n+1}$ to the sequence, which then becomes $\{(t_1, O_1),(t_2,O_2),\ldots,(t_n,O_n), (t_{n+1},O_{n+1}\}$, and that $t_{i+1} > t_i, \forall 1\leq i\leq n$.

In OWL and RDF, we support such a versioning mechanism through OWL annotation properties. Each version of a concept and object ontology is annotated with a builtin OWL annotation property \texttt{owl:versionInfo}, with the annotation value being the creation time of the ontology. Additionally, each version of an object ontology is also annotated with an annotation property \texttt{poddModel:mapsToVersion}. The value of this annotation is the timestamp of the version of the ontology of the concept that defines this object. We can view these two annotation properties as functions\footnote{For readability reasons we omit namespace names for the annotation properties in the following discussions.}:

\begin{align}
\mathtt{versionInfo}_C &\colon \bbold{O}\pfun \bbold{T} @ \mathtt{versionInfo}_C = (\mathcal{O}_C)^{\sim}\label{eqn:versioninfo}\\
\mathtt{mapsToVersion}_C &\colon \bbold{T}\pfun \bbold{T}\label{eqn:mapsTo}
\end{align}

In Definition~(\ref{eqn:versioninfo}), we define \texttt{versionInfo} as a function from ontologies $\bbold{O}$ to time $\bbold{T}$. It is easy to see that it is the inverse function of $\mathcal{O}$ defined in Definition~(\ref{eqn:version}). %Note that \texttt{versionInfo} is defined as a total function as we require each ontology to have a timestamp.

Definition~(\ref{eqn:mapsTo}) defines \texttt{mapsToVersion} as a function from time $\bbold{T}$ to time $\bbold{T}$. As stated above, for a given object $A$ and a given version of $A$'s ontology $O_A$, the version of the ontology of the corresponding concept $C$, $O_C$, that defines $O_A$ can be obtained by applying the above three functions:

\begin{align}
O_C = \mathcal{O}_C(\mathtt{mapsToVersion}_A(\mathtt{versionInfo}_A(O_A))) 
\end{align}

With the above versioning mechanism defined, we impose the following additional conditions to ensure that concepts and objects can evolve independently.
\begin{itemize}
\item For any given concept or object, a modification results in a new version of the ontology.
\item For objects, each new version of the ontology maps (\texttt{mapsToVersion}) to the latest version of ontology of its defining concept.
\item For each concept $C_1$,  if it references another concept $C_2$ during modification of its definition, the latest version of the ontology in $C_2$ is used.
\end{itemize}

\subsubsection{Auxiliary Functions}
We also define two auxiliary functions \emph{latest} and \emph{earlierThan}\footnote{The symbol $\power$ in the definitions that follow represents power set.}:

\begin{align}
latest&\colon\power(\bbold{T}\cross\bbold{O})\fun \bbold{O}\text{ where } \forall x\in\power(\bbold{T}\cross\bbold{O});O\in\bbold{O}, \label{eqn:latest}\\
      & latest(x) = O\iff\nonumber \exists t\colon\bbold{T} @ (t,O)\in x \land\nonumber\\
      &\t3\quad~~~ \nexists t_1\colon\bbold{T};O_1\colon\bbold{O} |(t_1,O_1)\in x \land t\leq t_1\nonumber\\
earlierThan&\colon\bbold{T}\cross\power(\bbold{T}\cross\bbold{O})\pfun\power(\bbold{T}\cross\bbold{O})\label{eqn:earlier}\\
      &\text{ where } \forall t\colon\bbold{T};x,y\colon \power(\bbold{T}\cross\bbold{O}) @ earlierThan(t,x) = y \iff\nonumber\\
      &\t1 y\subseteq x\land \nonumber\\
      &\t1 \forall (u,p)\colon x | u < t @ (u,p)\in y\land \nonumber\\
      &\t1 \forall (u,p)\colon y @ u < t\nonumber
%      &\t1 \exists S\colon \power x @ S = \{v\colon\bbold{T};p\colon\bbold{O} | v\leq t @ (v,p)\}\nonumber\\
%      &\t1 O = latest(S)\nonumber
\end{align}

Given a set of pairs from time $\bbold{T}$ to ontologies $\bbold{O}$, the function \emph{latest} returns the latest ontology from this set. This function can be applied to an entity (concept or object) $C$ to return the latest ontology definition: $latest(\mathcal{O}_C)$.

The function \emph{earlierThan} returns the maximal subset of a set of pairs from time $\bbold{T}$ to ontologies $\bbold{O}$ such that the timestamp of each of the pairs in this subset is earlier than the given time point.

\subsection{Dynamic Composition \& Reasoning of Ontologies}\label{sec:locality}
As stated in Section~\ref{sec:arch_desc}, we employ ontological inference to validate the integrity of concept and object definitions. Three types of reasoning tasks are performed for this purpose: ontology consistency checking, OWL class satisfiability checking (for concepts), and OWL individual assertion checking (for objects). Each of the latter two tasks is defined against an ontology $O$\footnote{Note that $O$ is a simplification from the more verbose and precise definition involving domain of interpretation $\Delta^\mathcal{I}$, interpretation function $\mathcal{I}$, the TBox and the ABox. Interested readers are referred to~\cite{basicdl03} for details.}.

\begin{align}
&O\nvDash \bot &\text{[Ontology consistency]}\label{eqn:ont}\\ 
&O\models \lnot \mathtt{C}\sqsubseteq \bot &\text{[class satisfiability]}\label{eqn:sat}\\
&O\models \mathtt{C}(\mathtt{a}) &\text{[individual class assertion]}\label{eqn:assert}
\end{align}

An inconsistent ontology can imply a false formula, hence the consistency of an ontology needs to be maintained. Definition (\ref{eqn:ont}) states that a consistent ontology $O$ should not be a model for the \texttt{owl:Nothing} ($\bot$) class. Definition (\ref{eqn:sat}) states that, with respect to an ontology $O$, for the class \texttt{C} to be consistent, the logical expression that \texttt{C} is not a subclass of $\bot$ must follow from $O$. Similarly, definition (\ref{eqn:assert}) states that for an individual \texttt{a} to be an instance of class \texttt{C}, this logic expression must follow from the ontology $O$.

Versioning of ontologies makes it possible for concepts and objects to evolve independently and it is one of the main reasons of the improved extensibility of the ontology-centric architecture. However, independent co-evolution of concepts and objects makes it a more challenging task to maintain semantic integrity of each of these entities, and hence that of the whole data management system.

In the rest of this section, we present in detail how ontologies of different concepts/objects are composed dynamically for the inference task. The formal definitions presented here are inspired by works on distributed description logics~\cite{Borgida:2002:DDL:646748.701674} and tiered logic for agents~\cite{DBLP:conf/icaart/CruzC09}.

First of all we introduce the notion of \emph{localities} for reasoning. Locality is an abstract concept: a locality gives scope for ontology reasoning tasks. There is one \emph{local} locality for each entity (concept or object) $C$. As $C$ uniquely represents the entity, we use $C$ to represent its locality as well. There is also a single and unique \emph{global} locality, denoted $\mathbf{G}$. With these notations, we denote an ontology $O_C$ for entity $C$ in its own locality $O_C^C$, $O^A_C$ in the locality of $A$, and simply $O_C$ in the global locality, omitting the superscript $^\mathbf{G}$.

\subsubsection{Transfer Rules Between Ontology Localities}
The dynamic composition of ontologies requires reasoning on ontologies to be applied for different concepts/objects in their own localities. Moreover, sometimes reasoning needs to be performed on the global level. Ontologies need to be \emph{transferred} semantically for these reasoning tasks. Contexts (localities) and transfer rules for logics have been defined in~\cite{springerlink:10.1007/978-3-642-11819-7_15,buvac95}. Such rules have also been shown to be sound and complete~\cite{buvac95}. Here we adapt these rules for ontologies and show them in Figure~\ref{fig:rules}, where $O$ represents an ontology, $\alpha$, $\beta$ represent the right-hand side boolean expressions in Definitions (\ref{eqn:ont}), (\ref{eqn:sat}) and (\ref{eqn:assert}), and $l$ and $k$ represent localities. The symbol $\models_l$ is to be interpreted as ``shows in the locality $l$'' and $\models$ as ``shows globally''.

\setlength\fboxsep{1pt}
\setlength\fboxrule{0.5pt}
\begin{figure*}[htb]
\centering\small
\fbox{
\begin{minipage}{.98\textwidth}
\hspace{2pt}
\begin{minipage}[t]{.4\textwidth}
\infer[\text{[exit]}]
{O\models \alpha^l}
{O\models_l \alpha}
\end{minipage}
\hspace{72pt}
\begin{minipage}[t]{.3\textwidth}
\infer[\text{[enter]}]
{O\models \alpha^l}
{O\models_l \alpha}
\end{minipage}
\hspace{44pt}
\begin{minipage}[t]{.3\textwidth}
\infer[\text{[K]}]
{O\models (\alpha^l\rightarrow \beta^l)}
{O\models (\alpha\rightarrow \beta)^l}
\end{minipage}
\vspace{16pt}
\linebreak
\begin{minipage}[t]{.4\textwidth}
\infer[\text{[DT]}]
{O\models (\lnot \alpha)^l\iff\lnot (\alpha^l)}
{}
\end{minipage}
\hspace{8pt}
\begin{minipage}[t]{.4\textwidth}
\infer[\text{[Flat]}]
{O\models (\alpha^l)^k \iff \alpha^l}
{}
\end{minipage}
\hspace{16pt}
\begin{minipage}[t]{.4\textwidth}
\infer[\text{[Flat-0]}]
{O\models_l \alpha\iff \alpha^l}
{}
\end{minipage}
\end{minipage}}
\caption{Transfer rules for ontologies between localities.}\label{fig:rules}
\end{figure*}

The (enter) and (exit) rules allow a formula to move up and down between localities. The (K) and (DT) rules ensure that logical connectives are preserved between localities. The (Flat) and (Flat-0) rules preserves truth/falsehood between localities.

\subsubsection{Global Ontology Consistency}
We maintain global consistency by ensuring the intersection of latest ontologies for all concepts is consistent (is not a model for $\bot$, as stated in Definition (\ref{eqn:ont})). This is performed whenever a new concept is added or an existing concept is modified. For any given concept $C$, we \emph{lift} the individual ontology from its own locality to the global locality. %For this reasoning talk, ontologies discussed in this subsection are the latest version for each concept.

For a data management system in this architecture with $n$ concepts, the overall data model $O^\mathbf{G}$ is the intersection of all ontologies in the global locality:

\begin{align}
O^\mathbf{G} = \bigcap_{i=1}^n O_{C_i} \text{ where } O_{C_i}\colon \mathcal{O}
\end{align}

\subsubsection{Concept Satisfiability}
As stated before, each version of the ontology of a concept defines the concept and possibly references other concepts. Hence, the satisfiability of the concept $C$ (at any given time point) is equivalent to the satisfiability of the corresponding OWL class $\mathtt{C}$ in the ontology $O_C^C$.

The ontology of any other concept that is referenced in the definition of $C$ needs to be included before the satisfiability of $C$ is checked, as defined in Definition (\ref{eqn:sat}). For any concept $C$, we define a function $closure_C$ that returns the closure of a version of its ontology $O$:

\begin{align}
closure_C&\colon \bbold{O}\fun\bbold{O} @ \forall (O,t)\colon\mathcal{O}_C, closure_C(O) = O_C^C \cap (\bigcap O_{C_i}^C)\label{eqn:aug_ont}\\
    &\t1 \text{for all concepts $C_i$ referenced in $C$ and}\nonumber\\
    &\t1 O_{C_i}^C = closure_{C_i}(latest(earlierThan(t, \mathcal{O}_{C_i})))\nonumber
\end{align}

Let $O_C$ denote a particular version of ontology for $C$ at time point $t$ (i.e., $(t,O_C)\in\mathcal{O}_C$). Let $O^C$ denote the ontology closure for $O_C$. Therefore,

\begin{align}
O^C &= closure_C(O_C)\\
O^C &\models \lnot \mathtt{C} \sqsubseteq \bot\label{eqn:aug_sat}
\end{align}

In Definition (\ref{eqn:aug_ont}), the ontology $O^C$ is defined to be the intersection of $O_C$ and all ontologies of other concepts \emph{lifted} from their own localities and \emph{lowered} to the locality $C$. Note that we apply the two functions in Definitions (\ref{eqn:latest}) and (\ref{eqn:earlier}) to get the latest ontology from $C_i$ that is earlier than $t$, as it does not make sense to reference an ontology defined in the future. $O^C$ then is used to check the concept satisfiability in Definition (\ref{eqn:aug_sat}).

\subsubsection{Individual Class Assertion}
Similar to concept satisfiability, individual class assertions involve the closure of the object ontology, which includes the ontology of the object itself, the ontology closure of the concept and the ontologies of all the other objects referenced in this object. For a given object $A$, we define a function $closure_A$ that returns the closure of a version of its ontology.

\begin{align}
closure_A&\colon \bbold{O}\fun\bbold{O} @ \forall (O,t)\colon\mathcal{O}_A, closure(O) = O_A^A~ \cap\\
    &\t1 closure_C(\mathcal{O}_C(\mathtt{mapsToVersion}(t)))~ \cap\nonumber\\
    &\t1 (\bigcap O_{A_i}^A) \text{, for all objects $A_i$ referenced in $A$ and }\nonumber\\
    &\t3 O_{A_i}^A = closure_{A_i}(latest(earlierThan(t,\mathcal{O}_{A_i})))\nonumber
\end{align}

For object $A$, let $O_A$ denote a particular version of its ontology at time point $t$ (i.e., $(t,O_A)\in\mathcal{O}_A$). Let $C$ be the concept that defines object $A$. Finally, let \texttt{C} denote the OWL class corresponding to concept $C$ and \texttt{a} denote the OWL individual corresponding to $A$. The closure of the ontology $O_A$ and the class assertion are as follows:

\begin{align}
O^A &= closure_A(O_A)\\
O^A &\models \mathtt{C}(\mathtt{a})
\end{align}

\subsection{Bootstrapping Concept Co-evolution}\label{sec:bootstrap}
In Section~\ref{sec:ont} we described the base ontology and a domain-specific ontology for phenomics. In these ontologies, initial definitions (first version of the ontologies) are defined. These ontologies will be decomposed to individual concept ontologies during the system bootstrapping process. Specifically, all \texttt{rdfs:subClassOf} (in the form of $C\sqsubseteq D$) and \texttt{owl:equivalentClass} (in the form of $C\equiv D$) axioms will be grouped by their left-hand side classes. Property and other common definitions will be grouped into a common ontology. After bootstrapping, the concepts will then evolve independently.

\section{The PODD Data Management System}\label{sec:sys}
\subsection{Implementation}
Based on the ontology-centric architecture presented in Section~\ref{sec:arch} and the base ontology presented in Section~\ref{sec:ont} we have implemented the PODD data management system. An instance of the PODD architecture has been deployed in production for the Australian Plant Phenomics Facility (APPF) and can be found at \url{http://podd.plantphenomics.org.au/}. The source code of the PODD repository software has been published in this repository\footnote{\url{http://podd.plantphenomics.org.au/podd/object/poddObject:696}}, demonstrating the versatility of the architecture.

%\vspace{-8pt}
\begin{figure}[htb]
\centering
\includegraphics[trim = 60mm 30mm 50mm 24mm, clip,height=72mm]{podd_arch.pdf}

\vspace{-8pt} \caption{The architecture and main components of the PODD system.}\label{fig:podd_arch}
\end{figure}

In developing the PODD system, we chose to employ a number of mature technologies, as can be seen in Figure~\ref{fig:podd_arch}. (1) We use Fedora Commons for the storage and retrieval of domain objects. Together with raw data files, the OWL (for concepts) and RDF (for objects) definitions of each concept and object are stored in a versioned datastream \texttt{PODD}, which is used by the PODD system in various tasks such as object creation, rendering, validation, update and visualization. Moreover, Fedora supports distributed storage through plugins, making it possible to increase the repository storage with demand. (2) We incorporate the Sesame~\cite{conf/iswc02/sesame} triple store to support complex query answering using SPARQL. Sesame \emph{contexts} are used to give scope to the RDF triples for each domain object. As described in Section III, access control needs to be enforced on a per project basis, which also needs to be enforced for query answering in the triple store. By identifying triples of individual objects, we are able to control contexts a user can access through query expansion. (3) We use the Solr\footnote{\url{http://lucene.apache.org/solr/}} open-source search engine platform to provide full-text search and faceted browsing capabilities. Similar to the structure of the Sesame triple store, there is a one-to-one correspondence between domain objects in the repository and the Solr \emph{documents}, the logical indexing units. (4) We use the Pellet~\cite{citeulike:2615678} open-source OWL reasoning engine together with the ontology manipulation library OWL API~\cite{citeulike:6578934} for the inference of ontologies. (5) Lastly, we use a MySQL database to store user and access control related information, as it is orthogonal to other domain concepts.

Although the architecture and the system are based on ontologies, the interface is designed to hide ontology-related complexity from the user and present information in an easy to use manner for all repository functions. For example, Figure~\ref{fig:browser} shows the browser view of a plant phenomics project that investigates leaf stomata in the Arabidopis thaliana model plant. In this view, the objects are shown in a tree-like structure by following parent-child assertions of subproperties of \emph{contains} defined in  the base and domain ontologies.

\begin{figure*}[htb]
\centering
\includegraphics[width=120mm]{browser.png}
\vspace{-8pt} 
\caption{The browser view of a plant project in the PODD repository.}\label{fig:browser}
\end{figure*}

\subsection{Recent Development}
To ease the importing of large amounts of data from existing databases into PODD, we have designed the PODD-TAB format, a simple tab-delimited flat file format that captures essential meta information about raw data. The PODD-TAB format is designed as an intermediate format for users without knowledge of ontologies. Together with the PODD-TAB format, we have developed APIs and tools that enables automatic deposit of data from databases in the Australian Phenomics Network.

As we stated previously, we store semantic metadata in an RDF triple store. To support advanced users, we have developed a query interface so that users can create complex SPARQL queries to interrogate the repository. SPARQL queries are also used internally by the repository to retrieve information otherwise hard to obtain. For example, for each object, we use a SPARQL query (Figure~\ref{fig:query}) to retrieve all the objects that refer to it.

\begin{figure}[htb]
\footnotesize\centering
\begin{minipage}{.9\textwidth}
\begin{verbatim}
PREFIX   rdfs:  <http://www.w3.org/2000/01/rdf-schema#>
PREFIX   poddObject:  <http://www.podd.org/object#>
PREFIX   owl:  <http://www.w3.org/2002/07/owl#>
PREFIX   xsd:  <http://www.w3.org/2001/XMLSchema#>
PREFIX   rdf:  <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
PREFIX   poddModel:  <http://www.podd.org/poddModel#>
SELECT DISTINCT ?subject ?type ?label 
WHERE {
    ?subject ?p ?o .
    ?p rdfs:subPropertyOf poddModel:refersTo .
    FILTER (str(?o) = 'http://www.podd.org/object#poddObject:7792')
    ?subject rdf:type ?type .
    ?subject rdfs:label ?label .
    filter (regex(str(?type), 'http://www.podd.org/poddModel#'))
} ORDER BY (?type) (?label)
\end{verbatim}
\end{minipage}
\caption{A SPARQL query to obtain all objects that refer to object \texttt{poddObject:7792}.}\label{fig:query}
\end{figure}

Development has also started on a module to support federated authentication based on the Shibboleth authentication framework\footnote{\url{http://shibboleth.internet2.edu/}}, which allows cross-organization single sign-on and removes the need for users to create a new user account for a Web application. Making use of the Australian Access Federation (AAF)\footnote{\url{http://www.aaf.edu.au/}} infrastructure allows users from AAF-enabled research institutes and universities (more than 60 at the moment) in Australia and New Zealand to access PODD in a seamless manner. The authentication functionality has been implemented and will be integrated in the next major release of PODD.

\subsection{Preliminary Evaluation}
We have started to deploy the PODD system in Australian phenomics research centers including APPF and APN and begun engaging users in the evaluation of the performance, flexibility, usability and scalability of the system. User feedback to date has shown that the system is intuitive and efficient.

It is well known that native RDF triple stores are not yet as efficient as relational database systems, especially in terms of query performance~\cite{10.1109/ICDE.2009.28,Bizer09theberlin}. In the PODD system, we employ a number of approaches to alleviate this problem. Firstly, we give scope to RDF triples of individual objects so a SPARQL query will only be matched against a (potentially very small) subset of the entire triple store, therefore improving query performance. Secondly, we index all RDF datatype property assertions and RDF type assertions in the Solr index to enable faceted searching and filtering. Together with the use of logical operators in search queries, full-text search can be used effectively to perform complex discovery tasks in most cases.

In summary, following the ontology-centric architecture and through the use of mature technologies we have successfully developed PODD, a scalable, extensible data management system. At the same time, data management tasks such as versioning, logical organization of data, authentication \& authorization and discovery are all supported effectively. The PODD system demonstrates the feasibility and practicality of the proposed ontology-centric architecture.

\section{Conclusion}\label{sec:concl}
In this work, our contribution to scientific data management is three-fold: firstly, the proposal of the ontology-centric architecture for developing highly extensible and domain-agnostic data management systems that support the evolution of domain concepts; secondly, the development of a base ontology that defines essential concepts in a domain-independent fashion; and thirdly, the development of a phenomics ontology and the data management system (based on both existing and new technologies) that serves as a validation of the feasibility of the proposed architecture in a virtual laboratory environment.

We have identified a number of future work directions that we would like to pursue:

\begin{itemize}
\item We will investigate the integration with existing domain ontologies such as the Gene Ontology and the Plant Ontology. Potential approaches include using terms defined in these ontologies to annotate metadata objects, and establishing semantic links between concepts/properties through ontology alignment/mapping.

\item We would like to investigate the generalization of the ontology-centric approach and the further abstraction of the PODD base ontology so that the architecture can be applied to other domains including archeology, geological and environmental sciences.

\item Thus far we have focused on essential functionalities of the PODD system. We will continue the development of the PODD system to provide additional functionalities such as data visualization, data mining and finer-grained representation of repository objects (file contents and metadata) in RDF.

\item As introduced in Section~\ref{sec:intro}, a huge and growing volume of semantic metadata have been made publicly available as part of the Linked Data project~\cite{citeulike:5008761}. In the Linked Data project, data is organized by application domain into individual RDF datasets. Thus far datasets about a number of domains, including biomedical research, government, geospatial, media and scientific publications, have been created. As of March 2011\footnote{\url{http://www4.wiwiss.fu-berlin.de/lodcloud/state/}}, the Linked Data \emph{cloud} contains more than 28 billion RDF triples and more than 395 million inter-dataset links. This metadata is rich, of high quality and continuously curated. It would be very beneficial for a data management system to reuse this public knowledge.

In our ontology-centric architecture, domain concepts and objects are described using ontological terms in RDF, and such semantic metadata is stored in an RDF triple store. Such a design encourages knowledge integration through the reuse of existing RDF vocabularies. We will investigate a methodology for adding semantic annotation to concepts and objects in the repository.

We will also investigate approaches to semantically disseminate concepts and objects in our architecture in Linked Data style. As stated above, due to the ontology-centric nature of the architecture, we do not foresee any major difficulties.

\item We will investigate the feasibility of extending the PODD ontology and architecture to support scientific workflow modeling and composition in a virtual laboratory environment, possibly integrating with the widely-used $^{my}$Experiment ecosystem~\cite{DBLP:journals/fgcs/RoureGS09}.

\item Our current authorization framework is based on user roles,  which are hardcoded and hence inflexible. We will investigate authorization frameworks such as Shibboleth, XACML\footnote{\url{http://www.oasis-open.org/committees/xacml/}} and SemanticAuthz~\cite{MarinPerez201140} and their suitability in our framework.

\item Last but not least, we will further enhance privacy preservation in the dissemination of repository contents through adoption of fine-grained privacy models such as~\cite{Sun2011348}.
\end{itemize}

\section*{Acknowledgment}
This work is part of a National eResearch Architecture Taskforce (NeAT) project, supported by the Australian National Data Service (ANDS) through the Education Investment Fund (EIF) Super Science Initiative, and the Australian Research Collaboration Service (ARCS) through the National Collaborative Research Infrastructure Strategy Program.

The authors wish to acknowledge the support of the NCRIS funded Australian Plant Phenomics Facility, including its primary nodes: The High Resolution Plant Phenomics Centre; and The Plant Accelerator. We also wish to acknowledge the support of the NCRIS funded Australian Phenomics Network.

Finally we wish to thank Dr Xavier Sirault and Dr Kai Xu for their assistance in developing the domain models.

\bibliographystyle{elsarticle-num}
\bibliography{all,liyf}

\end{document}

%%
%% End of file `elsarticle-template-num.tex'.

### 2012-03-02 ibis_workshop_2010/slides.tex
initial files from http://projects.arcs.org.au/svn/podd/trunk/doc/publications/

\documentclass[ignorenonframetext,compress]{beamer}

\mode<presentation>

\usepackage{times}
\usefonttheme{structurebold}

\usepackage[english]{babel}
\usepackage{tabularx}
\usepackage{url}
\usepackage{zed-csp}
\usepackage{subfigure}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{color}

\renewcommand*{\thesubfigure}{}

\usetheme{Madrid}

\title[PODD -- Ontology-driven Repository]{PODD: An Ontology-driven Data Repository for Collaborative Phenomics Research}

\author[Yuan-Fang Li]{
    \textbf{Yuan-Fang Li}\\
    \url{liyf@itee.uq.edu.au}
    }

\institute[UQ]{
    The eResearch Lab, School of ITEE\\The University of Queensland\\
    }
\vspace{+80pt}
\date[IBS Workshop 2010]{\footnotesize IBS Phenomics Data and Informatics Workshop - Canberra 2010}

\AtBeginSection[]{\frame{\frametitle{Outline}\tableofcontents[current]}}

\begin{document}

\begin{frame}
    \titlepage
\end{frame}

\part{Main Part}

\section{Introduction}
\begin{frame}{Motivation}
\begin{block}{Challenges in phenomics research data management}
    \begin{itemize}
        \item Data is huge
        \begin{itemize}
            \item APN: estimated 1.8TB + 30\% growth/year
            \item TPA: 0.5TB/week $\sim 25$ TB/year
        \end{itemize}
        \pause
        \item New platforms/processes/technologies emerge
        \pause
        \item Data needs \textbf{\emph{context}}
        \begin{itemize}
            \item Scientific, administrative \& other metadata
        \end{itemize}
    \end{itemize}
\end{block}
\pause
\begin{block}{Repositories for the management of data}
    \begin{itemize}
      \item Not all questions answered
    \end{itemize}
\end{block}
\end{frame}

There are 3 major challenges to data management in phenomics research.

1. The volume of data. Phenomics research generates large amounts of
data. For example, APN, the Australian Phenomics Network, is
estimated to generate 1.8TB of data in the first year with 30\%
growth. That is more than 7TB over 3 years. And APN is only one
research center. We still have APPF, which operates 2 centers. The
large volume makes the storage and distribution much more difficult.

2. Phenomics is a new area. As a result, new imaging platforms,
technologies and processes emerge all the time. A data repository
needs to be constructed in such a way that it can cater for future
platforms and technologies.

3. Data needs context. 7TB of data isn't very useful if it just sits
there and collects dust. To make use of the data, we need to
surround them with context, the metadata, that describe the raw data
in a meaningful way. Metadata can have dimensions, such as
scientific and administrative. They can help researchers to discover
data, integrate data and analyze data is a vital part of a data
repository.

There are quite a few databases for genomics and protein research,
such as Emsembl and UniProt. Unfortunately, not all the requirements
for phenomics data management are met. For example, genomic data is
relatively well defined, which isn't the case for phenomics.
Phenomics data may come off from a variety of imaging and measuring
platforms and the appropriate metadata needs to be captured.

The above challenges motivated us to develop the PODD repository.

\begin{frame}{Clients \& Collaborators}
\begin{block}{Australian Integrated Biological Science Facilities}
    \begin{itemize}
        \item<2-> Australian Plant Phenomics Facility (APPF)
            \begin{itemize}
                \item High-throughput (TPA) \& high-resolution (HRPPC) centers
            \end{itemize}
        \item<3-> Australian Phenomics Network (APN)
            \begin{itemize}
                \item Mouse models, \emph{deep} imaging and measuring platforms
            \end{itemize}
        \item<4-> Atlas of Living Australia (ALA)
            \begin{itemize}
                \item Biodiversity data collection \& management
            \end{itemize}
    \end{itemize}
\end{block}
~\\
\begin{figure}[ht]
\centering
\onslide<2->\subfigure[]{
\includegraphics[width=20mm]{conveyor.png}
}
\onslide<2->\subfigure[]{
\includegraphics[width=20mm]{sample.png}
}
\onslide<3->\subfigure[]{
\includegraphics[width=20mm]{mouse.png}
}
\onslide<4->\subfigure[]{
\includegraphics[width=20mm]{organism.png}
}
\end{figure}
\end{frame}

The PODD repository is designed to be used by a number of Australian
research organizations and projects.

Australian Plant Phenomics Facility has two research centers.

The Plant Accelerator (TPA) at the University of Adelaide is a
high-throughput plant research center. What you see here is a
conveyor belt system. RFID-tagged plants travel in the system and
their images are automatically taken by the imaging stations.

The High Resolution Plant Phenomics Center, the HRPPC, is located
here in Canberra. The center focuses on deep phenotyping of plants
using a number of equipments. Some of them operate on single plants
in controlled environments, others operate on fields.

The Australian Phenomics Network (APN) is a collaborative research
facility established to provides mouse models for human and animal
diseases. For example, APN uses digital slide scanner to perform
histopathology and organ pathology imaging and analysis.

The Atlas of Living Australia is developing a biodiversity data
management system for Australia. Collaboration with ALA is also one
of PODD's goals.

\begin{frame}{Data Management Requirements}
\begin{overprint}
\onslide<1>
\begin{block}{Data capturing}
\begin{tabularx}{\linewidth}{p{35mm}X}
Flow Cytometry & FACS data \\
Histopathology & Zeiss slide images \\
Plant imaging & Lemnatec images, Flourogroscan images, 3D imaging \\
Infrared imaging & FLIR images \\
Chemical measurements & Chlorophyll content, Stomatal conductance \\
Visual observation & Manual reports\\
$\cdots$ & $\cdots$
\end{tabularx}
\end{block}

\onslide<2>
\begin{block}{Data generation}
\begin{tabularx}{\linewidth}{p{30mm}X}
Project & Project proposal, project plan \\
Investigation & Objectives, design \\
Materials & Lines/genotypes, samples, growth conditions \\
Devices & Specs, settings, versions \\
Processes & Workflows, protocols, variations \\
Measurements & Data, images \\
Analysis & Observations, results \\
$\cdots$ & $\cdots$
\end{tabularx}
\end{block}

\onslide<3>
\begin{block}{Data management}
\begin{tabularx}{\linewidth}{p{50mm}l}
Data distribution & ? \\
Data sharing & ? \\
Data publishing & ? \\
Access control & ? \\
Archival \& versioning & ? \\
Structure \& metadata & ? \\
Data discovery \& analysis & ?
\end{tabularx}
\end{block}
\end{overprint}
\end{frame}

One of the requirements of PODD is to support data capturing from
various imaging platforms.

As you can see on this slide. there are quite a number of different
ways of imaging and observing the subject. The challenge here is,
besides capturing the raw image files and observations, capture
sufficient metadata.

Another requirement is data generation. Data and metadata can be
generated on different levels and at different times. All this
information needs to be recorded in the repository.

Last but not least, there is a need for the effective management of
data. Tasks such as distribution, publishing, access control and
versioning belong to this category. As far as we know, there is no
data repository that handles these tasks for phenomics research.

\begin{frame}{PODD: an ontology-driven repository}
\begin{block}{Goals}
    \begin{itemize}
        \item Supports data captured by different platforms
        \item Supports different data formats
        \item Supports effective data management
        \begin{itemize}
            \item Metadata, distribution, access control, discovery, etc.
        \end{itemize}
    \end{itemize}
\end{block}
\pause
\begin{block}{Approach}
    \begin{itemize}
        \item Define domain ontologies -- focus of this talk
        \item Design an appropriate architecture
        \item Develop an ontology-aware repository
    \end{itemize}
\end{block}
\end{frame}

So, the goals of developing the PODD repository is to support the
above data management tasks for the phenomics community.

After our initial study, we realized that the traditional
database-driven approach may not be the best way of doing it. We
will come and revisit this point later.

Ontologies are a formal representation of a particular application
domain. Besides being a controlled vocabulary that define important
concepts and relations, an ontology can define complex constraints
on these concepts and relations, making it a very expressive
conceptualization of the domain.

Compared to database schemas, ontologies are more open and
extensible. For these reasons, we choose to develop PODD in an
ontology-based approach. In this talk I will focus on describing how
the domain ontology we develop is used in the repository in the
lifecycles of the management of domain objects.

\begin{frame}{PODD: The High-level Architecture}
 \vspace{-6mm}
 \begin{figure}[t]
  \begin{center}
  \includegraphics[trim=48mm 20mm 10mm 35mm, clip, width=130mm]{architecture.pdf}
   \end{center}
  \label{fig}
 \end{figure}
\end{frame}

Here you see the high-level architecture of the PODD repository. On
the basis are a few components. Fedora Commons is the underlying
repository that we use to store all the metadata information. iRODS
is a federated data storage software that can be used by Fedora
Commons. ANDS provides an iRODS-based service called Data Fabric,
which we will use in PODD to host the data captured.

Ontologies will be stored in Fedora Commons as well as in a RDF
triple store to enable data discovery using SPARQL queries and full
text search on metadata. We also use databases to store information
about users, access rights and some other things.

The data access layer works with ontologies and objects and is
responsible for serializing and deserializing objects from the
underlying storage.

The various services and the web interface is guarded by a security
layer that controls access based on fine-grained user access rights.

In the following few slides I will quickly introduce a few projects
related to PODD. I'll also briefly introduce the OWL ontology
language.

\section{Related Work}
\begin{frame}{Related Work}
\begin{block}{FuGe -- Functional Genomics Experiment}
    \begin{itemize}
        \item \emph{Material}, \emph{Protocol}, \emph{Data}, etc.
        \item Can be extended to support phenomics
        \item[$\cross$] May be difficult to extend for new concepts
    \end{itemize}
\end{block}
\pause
\begin{block}{OBI -- Ontology for Biomedical Investigations}
    \begin{itemize}
        \item \textit{``An integrated ontology for the description of life-science and clinical
        investigations.''}
        \item Comprehensive: 2,600+ classes, 10,000+ axioms
        \item[$\cross$] Complex, computationally ($\mathcal{SHOIN}(D)$)
    \end{itemize}
\end{block}
\end{frame}

FuGe is a UML and database based conceptual model for functional
genomics experiments. It models important domain concepts such as
material, protocol and data. The model is stored as database
schemas, as a result, FuGe may be difficult to extend.

The OBI ontology - ontology for biomedical investigation is a
comprehensive but complex ontology. It defines domain concepts 
starting from a very high abstract level. As a result, it is 
quite big, containing about 26 hundred classes and 10 thousand
axioms. It is also computationally expensive. For this reason and
that we need to performing reasoning repeatedly in PODD, we chose
not to use OBI as our domain ontology.

\begin{frame}{Related Work}
\begin{block}{Web Ontology Language (OWL)}
\begin{itemize}
    \item Precise, open \& extensible -- exactly what we need!
    \item Provides core vocabularies for expressing complex ontologies -- \emph{data models}
    \item Main language constructs
    \begin{itemize}
        \item Classes, e.g, \texttt{Human}
        \item Predicates , e.g., \texttt{hasParent}
        \item Individuals, e.g., \texttt{Aristotle}
    \end{itemize}
    \item APIs, query engines \& automated reasoners available
\end{itemize}
\end{block}
\end{frame}

The Web Ontology Language is the ontology language with a precise semantics 
and is designed to be open and extensible. For example, OBI is developed
in OWL. It provides language constructs to define domain concepts and relations
and axioms on them.

The three main ingredients in OWL are classes, predicates and individuals. 
For example, Human can be defined as a class; hasParent can be defined as
a predicate and Aristotle can be defined as an individual.

Quite a few open-source APIs, query engines and reasoners are available for OWL.

\section{The PODD Ontology}
\begin{frame}{The Ontology Approach}
\begin{block}{Benefits}
    \begin{itemize}
    \setlength{\itemsep}{6pt}
    \item<1-> Greater extensibility
        \begin{itemize}
        \item New concepts/relations can be easily added/modified
        \end{itemize}
    \item<2-> Better reuse \& integration
        \begin{itemize}
        \item Ontologies are \emph{open}
        \item Other ontologies can be integrated on multiple levels
        \end{itemize}
    \item<3-> Balance between expressivity \& reasoning complexity
        \begin{itemize}
        \item Formal semantics enables automated query \& analysis
        \item Off-the-shelf tools available
        \end{itemize}
    \end{itemize}
\end{block}
\end{frame}

The ontology-based approach has a number of benefits.

Firstly, the repository can be more extensible. In PODD we will store all
object information in ontologies. In such a way new concepts and relations
can be added more easily than if the information is stored in a database.

An ontology-based repository is also better in terms of reuse and integration.
Ontologies are designed to be shared so when you make your ontology available,
you share your model. When you publish your instance data, concrete domain data
is shared. So integration can happen on multiple levels.

The OWL language is defined with a formal semantics. This enables automated
query answering and analysis. OWL is designed with expressivity and reasoning
complexity in mind so that the core reasoning tasks can be made fully automated. 
As shown in a previous slide, although OWL is a relatively young language,
there are off-the-shelf tools available to use.

\begin{frame}{The PODD Ontology}
\begin{block}{Modeling essentials}
    \begin{itemize}
    \item Domain entities
        \begin{itemize}
        \item Abstract concepts
        \item Concrete objects
        \end{itemize}
    \item Domain entities defined using OWL \emph{ontologies}
    \end{itemize}
\end{block}
\pause
\begin{block}{Modeling in OWL}
    \begin{itemize}
    \item Domain concepts -- \emph{OWL classes}
    \item Inter-concept relations -- \emph{OWL predicates} \& \emph{OWL restrictions}
    \item Concrete domain objects -- \emph{OWL Individuals}
    \item Comments, descriptions -- \emph{OWL annotations}
    \end{itemize}
\end{block}
\end{frame}

In the same spirit of FuGe and OBI, the PODD ontology defines 
essential domain concepts so that we can describe concrete domain
objects. These concepts and objects are defined using OWL ontologies
and RDF triples.

\begin{frame}{The PODD Ontology -- An Example}
\begin{example}{The \textbf{\emph{Project}} concept}
    \begin{itemize}
    \item The top-level concept
    \item Constraints on inter-object relations \& attributes -- like everything else
    \end{itemize}
    \pause
    {\small
    \begin{align*}
    Project &\sqsubseteq~ =~ 1~ hasProjectPlan \sqcap \forall hasProjectPlan.ProjectPlan\\
            &\sqsubseteq~ \geq~ 1~ hasInvestigation \sqcap \forall hasInvestigation.Investigation\\
            &\sqsubseteq~ =~ 1~ hasStartDate \sqcap \forall hasStartDate.date\\
            &\sqsubseteq~ \leq~ 1~ hasPublicationDate \sqcap \forall hasPublicationDate.date
    \end{align*}}
\end{example}
\begin{block}{}
    \begin{itemize}
    \item \textcolor{blue}{\textbf{Extensibility}} from inheritance of OWL classes \& predicates
    \end{itemize}
\end{block}
\end{frame}

In this slide I show a simple, incomplete modeling of the top-level concept
in our ontology, the Project. In our model of phenomics, everything starts 
from a research project and all other objects, such as Investigation and
Material are child or descendant objects.

As you will see, we define constraints on attributes and inter-concept relations
of the Project concept. This is how we constrain all of our concepts.

Shown here are 8 constraints. The first line says that any Project instance must
be mapped to exactly 1 other object by the hasProjectPlan object and this 1 object
must be of type ProjectPlan.

\begin{frame}{The PODD Ontology}
\onslide<2->
\begin{columns}[t]
\column{.5\textwidth}
\begin{itemize}
    \item Inspired by FuGe \& OBI
    \item Aim: define all \emph{essential} domain concepts, attributes \& relations
    \item $\sim30$ classes, \\$\sim 80$ predicates, \\$\sim 200$ Axioms
    \item Faster reasoning, querying, etc.
\end{itemize}
\vfill
\onslide<1->
\column{.5\textwidth}
\vspace{-10mm}
 \begin{figure}[t]
  \begin{center}
  \includegraphics[trim=3mm 26mm 2mm 8mm, clip, height=72mm]{hier.pdf}
   \end{center}
 \end{figure}
\end{columns}
\end{frame}

Here you see a list of important domain concepts in the PODD ontology.
In the development of our ontology we took inspiration from FuGe and OBI
but with the aim that the PODD ontology will cover essential domain
concepts and relations without becoming too complex.

As it currently stands, the ontology defines about 30 OWL classes, 80
predicates and more than 200 logical axioms.

With this smaller ontology, all core reasoning tasks can be performed
much faster.

\begin{frame}{PODD Ontology -- Roles}
\begin{block}{Ontologies \emph{drives} repository functions}
\begin{list}{}{}
\item[\textcolor{blue}{\textbf{Presentation}}] Drives the rendering of the web pages
    \begin{itemize}
    \item Object creation, editing, display, etc.
    \end{itemize}
\item[\textcolor{blue}{\textbf{Management}}] Object (de)serialization to/from ontologies
\item[\textcolor{blue}{\textbf{Validation}}] Ontology reasoning performed
    \begin{itemize}
    \item Validation on object type, cardinality, etc.
    \end{itemize}
\item[\textcolor{blue}{\textbf{Discovery}}] Multiple ways of finding information
    \begin{itemize}
    \item Queries using SPARQL
    \item Searches using Lucene
    \end{itemize}
\end{list}
\end{block}
\end{frame}

The PODD ontology is at the center of the repository and it drives quite a lot
of repository functions. All the object metadata is saved as RDF triples
conforming to the ontology definitions. When a web page of creating, editing
or displaying an object is requested, the ontologies are used to determine
what gets displayed and in what way.

The domain objects are serialized into ontologies when they are saved and 
deserialized when they are loaded from the underlying storage.

Before we save and publish an object, the object's triples are validated against
the domain ontology. Objects can be queried and searched for on the RDF triples.

\begin{frame}{The PODD Ontology -- Object creation page rendering}
 \begin{figure}[t]
  \begin{center}
  \includegraphics[width=76mm]{create.png}
   \end{center}
 \end{figure}
\end{frame}

\section{Conclusion}
\subsection*{ }
\begin{frame}{Conclusion}
\begin{block}{To recap}
\begin{itemize}
\item Large amounts of data need to be managed
    \begin{itemize}
    \item There is a need for data archival, storage \& discovery
    \end{itemize}
\item Current approaches lacking/inadequate/inflexible
    \begin{itemize}
    \item Emerging processes, platforms, technologies require extensible infrastructure
    \end{itemize}
\item An ontology-driven approach as the foundation of PODD
    \begin{itemize}
    \item Extensible \& open
    \end{itemize}
\end{itemize}
\end{block}
\end{frame}

This slides shows a screen capture of the project creation page.
Note that most of these fields are generated by reading the domain
ontology and figuring out what fields are needed and what their types,
cardinalites are.

For example, in a previous slide we defined that a project must have 
exactly 1 start date attribute and it must be a date. So here we show 
an input box for this attribute and signal that it is mandatory. Note
that lead institution is also a mandatory attribute while the duration
is not. When this project object is saved, the values are checked against
the domain ontology and any errors or omissions will be reported back
to the user.

\begin{frame}{Conclusion}
\begin{block}{Where we are now}
\begin{itemize}
\item PODD ontology stabilizing
\item Development of basic repository functionality
\item Development of PODD web interface
\end{itemize}
\end{block}
\pause
\begin{block}{What's next}
\begin{itemize}
\item Development of batch data import/export processes
\item Development of object discovery services
\item Integrate with AAF federated authentication services
\item \textcolor{gray}{Exposing data for discovery}
\item \textcolor{gray}{Integrating with other data sources}
\end{itemize}
\end{block}
\end{frame}

\begin{frame}
	\frametitle{}
    \begin{center}
\Huge \textsc{\textcolor{blue}{\textbf{Thank You!}}}
\end{center}
\vspace{+20pt}
\begin{block}{Acknowledgment}
Faith Davies, Gavin Kennedy, Jane Hunter\\
eResearch Lab, Schoo of ITEE, UQ
\end{block}
\end{frame}

\end{document}

### 2012-03-02 icadl2010/icadl_38/podd.tex
initial files from http://projects.arcs.org.au/svn/podd/trunk/doc/publications/

\documentclass{llncs}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{boxedminipage}
\usepackage{oz}

\setcounter{tocdepth}{3}
\def\keywordname{{\textbf{Keyword:}}}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

\title{PODD: An Ontology-driven Data Repository for Collaborative Phenomics Research}

\author{Yuan-Fang Li \and Gavin Kennedy \and Faith Davies \and Jane Hunter}

\institute{School of ITEE, The University of Queensland
\email{\{uqyli4,g.kennedy1,f.davies,j.hunter1\}@uq.edu.au}}

\maketitle

\begin{abstract}
Phenomics, the systematic study of phenotypes, is an emerging field
of research in biology. It complements genomics, the study of
genotypes, and is becoming an increasingly critical tool to
understand phenomena such as plant morphology and human diseases.
Phenomics studies make use of both high- and low-throughput imaging
and measurement devices to capture data, which are subsequently used
for analysis. As a result, high volumes of data are generated on a
regular basis, making storage, management, annotation and
distribution a challenging task. Sufficient contextual information,
the metadata, must also be maintained to facilitate the
dissemination of these data. The challenge is further complicated by
the need to support emerging technologies and processes in phenomics
research. This paper describes our effort in designing and
developing an ontology-driven, open, extensible data repository to
support collaborative phenomics research in Australia.

\keywords{OWL, ontology, repository, phenomics, data management}
\end{abstract}

\section{Introduction}\label{sec:intro}
An organism's phenotype is an observable or quantifiable trait of
the organism as a consequence of its genetic makeup combined with
its developmental stage, environment and disease conditions.
Phenomics is the systematic and comprehensive study of an organism's
phenotype and is determined through a combination of high-throughput
and high-resolution imaging- and measurement-based analysis
platforms. Phenomics research, together with genomics research,
represents a holistic approach to biological
study~\cite{dhoule09,Sauer200458,nevo01}.

Unlike genomics, phenomics research emphasizes physical, observable
traits of the subject under study. Like genomics, vast amounts of
data are produced by imaging and measurement platforms and analysis
tools. The storage, management, analysis and publication of these
data is a challenging problem.

Specifically, there are three key challenges for data management in
phenomics research.
\begin{itemize}
\item The ability to provide a data management service that can
manage large quantities of heterogeneous data in multiple formats
(text, image, video) and not be constrained to a finite set of
imaging and measurement platforms and data formats.

\item The ability to support metadata-related services to
provide context and structure for data within the data management
service to facilitate effective search, query and dissemination.

\item The ability to accommodate evolving and emerging technologies
and processes as phenomics is still a rapidly developing field of
research.
\end{itemize}

The Phenomics Ontology Driven Data (PODD)
repository\footnote{\url{http://www.itee.uq.edu.au/~eresearch/projects/podd/}}
is being developed to meet the above challenges facing the
Australian phenomics research community, aiming at providing
efficient and flexible repository functionalities for large-scale
phenomics data. An important goal of PODD is to provide a mechanism
for maintaining structured and precise \emph{metadata} around the
raw data so that they can be distributed and published in a reusable
fashion.

Differences in research project reporting, organisms under study,
research objectives, research methodologies and imaging and
measurement platforms may result in differences in the \emph{models}
of data. In order to accommodate as wide a variety of biological
research activities as possible, we have constructed the domain
model using OWL~\cite{hoph03a} ontologies, instead of the
traditional UML class diagrams and database schemas. The OWL domain
model is at the core of the PODD repository as it drives the
creation, storage, validation, query and search of data and
metadata. In contrast to traditional data repositories that use
database schemas as the underlying model, the employment of OWL
ontologies as the domain model makes PODD highly extensible.

In this paper, we present our work in addressing the above
challenges and highlight the OWL-based modeling approach we take.
The rest of the paper is organized as follows. In
Section~\ref{sec:overview} we present some related work and give a
brief overview of PODD and relevant technical background.
Section~\ref{sec:design} presents the high-level design of the
repository. In Section~\ref{sec:ont}, we discuss the PODD domain
ontology in more detail and show how the ontology-based modeling
approach is used in the life cycle of domain objects. Finally,
Section~\ref{sec:conclusion} concludes the paper and identifies
future directions.

\section{Overview}\label{sec:overview}
\subsection{Related Work}
In biological research, a large number of databases have been
developed to host a variety of information such as genes
(Ensembl\footnote{\url{http://www.ensembl.org/}}), proteins
(UniProt\footnote{\url{http://www.uniprot.org/}}), scientific
publications
(PubMed\footnote{\url{http://www.ncbi.nlm.nih.gov/pubmed/}}) and
micro-array data
(GEO\footnote{\url{http://www.ncbi.nlm.nih.gov/geo/}}).

These databases are generally characterized by the fact that they
specialize in a particular kind of data (protein sequences,
publications, etc.) and the conceptual domain model is relatively
well understood and stable. As a result, the need for extensibility
and flexibility is not very high.

Phenomics is a fast growing discipline in biology and new
technologies and processes are evolving and emerging rapidly. As a
result, the domain model must be flexible enough to cater for such
changes. Currently there are a number of related domain models
available.

Functional Genomics Experiment Model (FuGe)~\cite{citeulike:1756058}
is an extensible modeling framework for high-throughput functional
genomics experiments, aiming at increasing the consistency and
efficiency of experimental data modeling for the molecular biology
research community. Centered around the concept of experiments, it
encompasses domain concepts such as protocols, samples and data.
FuGe is developed using UML from which XML Schemas and database
definitions are derived. The FuGe model covers not only
biology-specific information such as molecules, data and
investigation, it also defines commonly-used concepts such as audit,
reference and measurement.

Extensions in FuGe are defined using inheritance of UML classes. We
feel that the extensibility we require is not met by FuGe as any
addition of new concepts would require the development of new
database schemas and code. Moreover, the concrete objects reside in
relational databases, making subsequent integration and
dissemination more difficult.

The Ontology for Biomedical Investigations
(OBI)\footnote{\url{http://obi-ontology.org/}} is an on-going effort
of developing an integrative ontology for biological and clinical
investigations. It takes a top-down approach by reusing high-level,
abstract concepts from other ontologies. It includes 2,600+
OWL~\cite{hoph03a} classes and 10,000+ axioms (in the import closure
of the OBI ontology). Although OBI is very comprehensive, its size
and complexity makes reasoning and querying of OBI-based ontologies
and RDF graphs computationally expensive and time consuming.

\subsection{The PODD Repository}
Under the National Collaborative Research Infrastructure Strategy
(NCRIS), the Australian Government has funded two major phenomics
initiatives: The Australian Plant Phenomics Facility (APPF),
specializing in phenotyping crop and model plant species; and the
Australian Phenomics Network (APN), which specializes in the
phenotyping of mouse models. Both facilities have common
requirements to gather and annotate data from both high- and
low-throughput phenotyping devices. The scale of measurement can be
from the micro or cellular level, through the level of a single
organism, and up to (in the case of the APPF) the macro or field
level.

An organism's phenotype, observable and quantifiable traits, is
often the product of the organism's genetic makeup, its development
stage, disease conditions and its environment. Any measurement made
against an organism needs to be recorded in the context of these
other data. The opportunity exists to create a repository to record
the data, its contextual data (metadata) and data classifiers in the
form of ontological or structured vocabulary terms. The structured
nature of this repository would support manual and autonomous data
discovery as well as provide the infrastructure for data based
collaborations with domestic and international research
institutions. Currently there are no such integrated systems
available to the two facilities.

The National eResearch Architecture Taskforce (NeAT) Australia
initiated the PODD project to fill this gap. In PODD, we have
engaged in the design and development of the Phenomics Ontology
Driven Data (PODD) repository. The goal of PODD is to capture,
manage, annotate and distribute the data generated by phenotyping
platforms. It supports both Australian and international biological
research communities by providing repository and data publication
services.

\subsection{The OWL Ontology Language}
The Web Ontology Language (OWL)~\cite{hoph03a} is one of the
cornerstone languages in the Semantic Web technology stack. Based on
description logics~\cite{DBLP:conf/dlog/McGuinness03}, OWL DL (a sub
species of OWL) defines a precise and unambiguous semantics, which
is carefully crafted so that it is very expressive yet core
reasoning tasks can be fully automated.

Information in OWL, as in RDF~\cite{rdfprimer04}, is modeled in
triples: $\langle subject, predicate,\allowbreak object\rangle$,
where $subject$ is the entity of interest, $predicate$ represents a
particular characteristic/property of the entity and $object$ is the
actual value of that property. \emph{Class}es are first-class
citizens in OWL. They represent abstract concepts in a particular
domain. Concrete objects are represented by OWL \emph{individual}s.
OWL \emph{predicate}s are used to relate OWL entities (classes,
individuals, predicates) to their attributes or other entities.

\vspace{-16pt}
\begin{figure}[htb]
\scriptsize
\hspace{-28pt}
\begin{minipage}[t]{0.6\textwidth}
\begin{syntax}
\mathcal{C} &::= &C                              & Class name\\
            & |  &\top                           & Top class\\
            & |  &\bot                           & Bottom class\\
            & |  &\mathcal{C} \sqcup \mathcal{C} & Class union\\
            & |  &\mathcal{C} \sqcap \mathcal{C} & Class intersection\\
            & |  &\lnot \mathcal{C}              & Class negation\\
            & |  &\forall P.\mathcal{C}          & Universal quantification\\
            & |  &\exists P.\mathcal{C}          & Existential quantification\\
            & |  &P\colon o                      & Value restriction\\
            & |  &\geq n~ P                      & At least number restriction\\
            & |  &\leq n~ P                      & At most number restriction\\
            & |  &\{a_1, \cdots, a_n\}           & Enumeration
\end{syntax}
\vspace{-16pt}
\caption{OWL expressions}\label{fig:owl-des}
\end{minipage}
%
\hspace{-12pt}
\begin{minipage}[t]{0.45\textwidth}
\begin{syntax}
\mathcal{AX} &::=& C \sqsubseteq \mathcal{C}                & Class subsumption\\
             & | & C = \mathcal{C}                          & Class equivalence\\
             & | & C \sqcap \mathcal{C} = \bot              & Class disjointness\\
             & | & P \sqsubseteq P                          & Property subsumption\\
             & | & P = P                                    & Property equivalence\\
             & | & \geq 1~ P\sqsubseteq \mathcal{C}         & Property domain\\
             & | & \top \sqsubseteq \forall P.\mathcal{C}   & Property range\\
             & | & \top \sqsubseteq \leq 1~ P               & Functional property\\
             & | & P = ({}^- P)                             & Inverse property
\end{syntax}
\vspace{+20pt} \centering\caption{OWL class and predicate
axioms}\label{fig:owl-axm}
\end{minipage}
\end{figure}

\vspace{-12pt}
\emph{Class descriptions} can be used to construct (anonymous)
complex class expressions from existing ones, as shown in
Figure~\ref{fig:owl-des}. $\mathcal{C}$ represents (possible
complex) class expressions; $C$ is a class name; $P$ stands for a
predicate; $n$ is a natural number and $a_i$'s are individuals.
OWL uses \emph{axioms} to place restrictions on OWL classes and
predicates. These axioms include class subsumption, equivalence,
disjointness; predicate domain, range, etc. Figure~\ref{fig:owl-axm}
shows some of the axioms.

The OWL language has been widely used in life sciences and
biotechnology~\cite{journals/bib/RuttenbergRSM09,citeulike:1882392,citeulike:212874}
as a modeling language for its expressivity and extensibility. There
is also growing tool support for tasks such as reasoning, querying
and visualization, making it a viable option for the modeling and
representation of domain concepts and objects in phenomics.

\section{High-level Design of PODD}\label{sec:design}
PODD is intended to be an open platform that allows any user to
access data that is either published or explicitly shared with them
by the data owners. Moreover, PODD has been envisioned to provide
data management services for a wide variety of research projects and
phenotyping platforms.

The key design considerations of PODD include:
\begin{itemize}
\item Data storage and management. It is estimated that several TB
of data will be generated by PODD clients per year. Hence, the
ability to efficiently manage large volumes of data is crucial.

\item Repository reusability. Data generated by a wide range of
projects and phenotyping platforms will be stored in the repository.
Hence, the domain model needs to be flexible enough to cater for the
administrative, methodological and technical differences across
projects and platforms.

\item Data persistence and identification. In order to support the
dissemination of scientific findings, data in the repository needs
to be publicly accessible after being published. Hence, a persistent
naming scheme is required.
\end{itemize}

In the development of PODD we employ a number of core technologies
to meet the above requirements.

\begin{itemize}
\item We use Fedora Commons\footnote{\url{http://www.fedora-commons.org/}}, a digital
repository for the management, storage and retrieval of domain
objects.

\item We use iRODS~\cite{irods07}, a distributed, grid-based storage
software system, for the actual storage solution of domain objects
across a virtual data fabric.

\item We incorporate the Sesame\footnote{\url{http://www.openrdf.org/}}
triple store for the storage and query of RDF triples (ontology
definitions of concrete objects).

\item We use the Lucene\footnote{\url{http://lucene.apache.org/}}
open-source search engine for the full-text index and search of repository
contents, including values in the RDF triple store.
\end{itemize}

The high-level architecture of the PODD repository can be seen in
Figure~\ref{fig:arch}.

\begin{figure}[htb]
\centering
\includegraphics[trim = 45mm 50mm 45mm 40mm, clip,height=70mm]{architecture.pdf}
\vspace{-16pt} \caption{The high-level architecture of the PODD
repository.}\label{fig:arch}
\end{figure}

\section{Ontology-based Domain Modeling}\label{sec:ont}
One of the important design decisions to be made early in the
development process is the domain model. Domain modeling aims at
providing solutions for two important tasks. Firstly, efficient and
flexible data organization; and secondly, data contextualization in
the form of \emph{metadata}, so as to provide meaning for the raw
data (documents, publications, image files, etc.) to facilitate
search, query, dissemination, and so on.

As we emphasized previously, the domain model should be flexible
enough to accommodate the rapid changes and advancement of phenomics
research. Inspired by FuGe and OBI, we created our own PODD ontology
in OWL to define essential domain concepts and relations.
%To
%improve interoperability, we maintain mappings between terms in the
%PODD and OBI ontologies through annotations.

\subsection{The PODD Ontology}
As PODD is designed to support the data management tasks for
phenomics research, a number of essential domain concepts and the
relationships between these concepts need to be modeled. Domain
concepts will be modeled as OWL classes; relationships between
concepts and object attributes will be modeled as OWL object- and
datatype-predicates. Concrete objects will be modeled as OWL
individuals.

\subsubsection{The Domain Concepts} In our modeling, the top-level
concept is \textbf{\emph{Project}}, which is an administrative
concept and contains essential meta information about the research
project, such as the administering organization, principal
investigator, project membership, project status, etc.

A number of concepts may be associated with the project concept.

\begin{description}
\item[\emph{Project Plan}] describing the
current project plan at the core metadata element level.

\item[\emph{Platform}] describing any single technical
measurement platform used in the project. Technical measurement
platform means any platform for which parameters and parameter
values may be captured.

\item[\emph{Genotype}] describing the genotype of the
materials used in the investigations. Multiple genotypes are
described here and then can become fields in the instances of the
Material object.

\item[\emph{Investigation}] describing a planned process
within a project that executes some form of study design and
produces a coherent set of results. It can be considered equivalent
to an experiment.
\end{description}

The \textbf{\emph{Investigation}} concept is of central importance.
It captures the data and metadata of experiments under a project. A
number of concepts are defined to assist in the modeling of
investigations.

\begin{description}
\item[\emph{Experimental Design}] describing experimental design
components, e.g. plant layouts, sampling strategies, etc.

\item[\emph{Growth Condition}] describing growth conditions, such as
growth chambers used, environmental settings, etc.

\item[\emph{Process}] representing a planned component of an
investigation. It is a description of a series of steps taken to
achieve the objective of the investigation.

\item[\emph{Protocol}] describing a step within a process that is a
consistent whole. e.g. sterilize seeds, plant seedlings, image the
plants.

\item[\emph{Material}] describing the materials used in the
investigation. Materials can be either inputs or outputs. They can
be chemicals, substrates, whole organisms or samples taken from a
whole organism. The meaning of the Material is usually derived from
its position in the model (as well as core metadata elements such as
Type).

\item[\emph{Event}] capturing ad-hoc events and actions that occur
against an individual material. In most instances the events and
their timing are described in the process/protocol. An Event object
can be utilized to either record fixed events in a form that allows
for investigative analysis, or to record one off observations (e.g., the
plant under observation died).

\item[\emph{Measurement}] describing a single measurement against a
single material. e.g. an image of a plant is a single measurement.
Measurement objects can capture measurement variables (e.g. shutter
speed, lighting, etc).

\item[\emph{Analysis}] An Analysis object is a variation on a
protocol object, in that it describes a step in a process. It is
currently linked to the Prject, Investigation objects as well as to
specific Measurement objects since analyses can be performed on
single measurements and also on the entire investigation or project
based on multiple inputs.
\end{description}

\subsubsection{Inter-concept Relationships}
The structures and workflows of phenomics research activities are
captured using inter-concept relationships, which are defined using
OWL predicates.

Different research projects will utilize different measurement
platforms and have significantly different approaches and project
designs.. As a result, different projects may have different
structures. In OWL, there are a number of ways of defining the same
relationship. In order to achieve high modeling flexibility and
accommodate as many scenarios as possible, we have made the
following design decisions:

\begin{itemize}
\item Use OWL restrictions to define inter-concept relationships.
OWL restrictions impose constraints on the OWL classes they are
defined in.

\item Only define domain, or range, but not both, for predicates, so that the
predicates can be used by different concepts.
\end{itemize}

For each of the concepts described in the previous subsection except
for \emph{Project}, we define an OWL object-predicate with the range
being the concept. For example, for concept \emph{Analysis}, we
define a predicate \emph{hasAnalysis} and define its range to be
\emph{Analysis}.

\subsubsection{Object Attributes}
Attributes are inherent properties of an object, such as the start
date of a project, the timestamp of an event, etc. In our modele, we
use OWL datatype-predicates to model object attributes, similar to
the modeling of inter-object relationships.

Figure~\ref{fig:project_owl} shows the partial definition of the OWL
class $Project$, in OWL DL syntax~\cite{hoph03a}. Restriction
\ref{for:hasPlan} states that any $Project$ instance must have
exactly one $ProjectPlan$ (through the predicate $hasProjectPlan$,
the range of which is $ProjectPlan$). The other 3 restrictions are
similarly defined.

\vspace{-20pt}
\begin{figure}[htb]
\centering
\begin{align}
Project &\sqsubseteq~ =~ 1~ hasProjectPlan\label{for:hasPlan}\\
        &\sqsubseteq~ \geq~ 1~ hasInvestigation\\
        &\sqsubseteq~ =~ 1~ hasStartDate\\
        &\sqsubseteq~ \leq~ 1~ hasPublicationDate
\end{align}
\vspace{-16pt} \caption{Partial OWL Definition for the Project
concept.}\label{fig:project_owl}
\end{figure}

\vspace{-20pt}
\subsection{Ontology-based Model in Object Life Cycle}\label{sec:rav}
Concrete objects, instantiations of various concepts such as
\emph{Project} and \emph{Investigation}, are stored in PODD and can
subsequently be retrieved for different purposes. As stated in
Section~\ref{sec:intro}, the ontology-based domain model is at the
center of the whole life cycle of objects. In this subsection, we
briefly present the roles the ontology-based model perform at
various stages of the object life cycle.

\begin{description}
\item[Ingestion] When an object is created, the user specifies which
type of object she intends to create and the repository will pull up
all the ontological definitions for that type (from the OWL class
corresponding to that type and its super classes). Such definitions
will be used to (a) guide the rendering of object creation
interfaces and (b) validate the attributes and inter-object
relationships the user has entered before the object is ingested.

\item[Retrieval \& update] When an object is retrieved from the repository,
its attributes and inter-object relations are retrieved from its
ontology definition, which is used to drive the on-screen rendering.
When any value is updated, it is validated and updated in object's
ontology definition.

\item[Query \& search] An object's ontology definitions will be stored
in an RDF~\cite{rdfprimer04} triple store, which can be queried
using the SPARQL~\cite{sparql} query language. Similarly, ontology
definitions can be indexed and searched by search engines such as
Lucene.
\end{description}

In summary, ontology-based domain modeling enables us to build very
expressive and extensible phenomics domain models. Ample tool
support is also available to perform ontology-based tasks such as
validation, querying and searching.

\section{Conclusion}\label{sec:conclusion}
Phenomics is an emergent discipline that is poised to have a
significant impact upon industrial-scale biological research.
Phenomics presents a number of data management challenges, such as
managing high volumes of data, integrating highly heterogeneous
datasets and ensuring the data will exist in perpetuity.

To meet the data management needs of the Australian phenomics
research community, the PODD repository is being developed to enable
efficient storage, retrieval, contextualization, query, discovery
and publication of large amounts of data.

Central to the design and development of PODD is the use of OWL
ontologies as the domain model. Based on description logics and with
an emphasis on the Web, the OWL language features a precise
semantics, high expressivity and high extensibility. It also has
mature and growing tool support. As a result, the OWL language has
been widely used in bioinformatics and life sciences to mark up
genetic, molecular and disease information.

In the PODD model, core domain concepts are defined as OWL classes.
Their attributes and relationships with other domain concepts are
defined as OWL class restrictions. Concrete domain objects are then
initialized, conforming to the ontologies defined for their
concepts. Such a modeling approach has a number of benefits.

\begin{itemize}
\item Firstly, concepts are defined with unambiguous syntax and
precise semantics, enabling automated validation and rendering.

\item Secondly, the extensible nature of OWL language ensures that
new concepts can be easily added.

\item Thirdly, as OWL and RDF are both open standards,
interoperability between repositories is expected to be high.
\end{itemize}

In this paper, we introduce the high-level architecture of the PODD
repository and the main technologies used in developing it. We focus
on the ontology-based domain modeling approach, present the PODD
domain model and discuss its role in the life cycle of concrete
domain objects.

The development of the PODD repository will focus on enhancing
ontology-based modeling, representation and processing of phenomics
data. Ontology-based annotation services, automated data integration
and data visualization will be part of the future directions.

\section*{Acknowledgement}
The authors wish to acknowledge the support of the National
eResearch Architecture Taskforce (NeAT) and the Integrated
Biological Sciences Steering Committee (IBSSC). The authors wish to
thank Dr Xavier Sirault and Dr. Kai Xu for the discussion on the
development of the PODD ontology.

\newcommand{\gobble}[1]{}
\begin{thebibliography}{10}

\bibitem{citeulike:212874}
M.~Ashburner, C.~A. Ball, J.~A. Blake, et~al.
\newblock {Gene Ontology: Tool for the Unification of Biology}.
\newblock {\em Nat Genet}, 25(1):25--29, May 2000.

\bibitem{hoph03a}
Ian Horrocks, Peter~F. Patel-Schneider, and Frank van Harmelen.
\newblock {From $\mathcal{SHIQ}$ and {RDF} to {OWL}: The Making of a Web
  Ontology Language}.
\newblock {\em Journal of Web Semantics}, 1(1):7--26, 2003.

\bibitem{dhoule09}
David Houle.
\newblock {Numbering the Hairs on Our Heads: The Shared Challenge and Promise
  of Phenomics}.
\newblock {\em {Proceedings of the National Academy of Sciences}}, October
  2009.

\bibitem{citeulike:1756058}
Andrew~R. Jones, Michael Miller, Ruedi Aebersold, et~al.
\newblock {The Functional Genomics Experiment model (FuGE): an Extensible
  Framework for Standards in Functional Genomics}.
\newblock {\em Nature Biotechnology}, 25(10):1127--1133, October 2007.

\bibitem{rdfprimer04}
F.~Manola and E.~Miller (editors).
\newblock {{RDF Primer}}.
\newblock \url{http://www.w3.org/TR/rdf-primer/}, February 2004.

\bibitem{DBLP:conf/dlog/McGuinness03}
Deborah~L. McGuinness.
\newblock Configuration.
\newblock In Franz Baader, Diego Calvanese, Deborah McGuinness, Daniele Nardi,
  and Peter~F. Patel-Schneider, editors, {\em Description Logic Handbook},
  pages 388--405. Cambridge University Press, 2003.

\bibitem{nevo01}
Eviatar Nevo.
\newblock Evolution of genome-phenome diversity under environmental stress.
\newblock {\em {Proceedings of the National Academy of Sciences}},
  98(11):6233--6240, March 2001.

\bibitem{sparql}
Eric Prud'hommeaux and Andy Seaborne.
\newblock {SPARQL Query Language for RDF}.
\newblock \url{http://www.w3.org/TR/2006/CR-rdf-sparql-query-20060406/}, April
  2006.

\bibitem{irods07}
A.~Rajasekar, R.~Moore, and F.~Vernon.
\newblock {iRODS: A Distributed Data Management Cyberinfrastructure for
  Observatories}.
\newblock In {\em American Geophysical Union, Fall Meeting 2007}, December
  2007.

\bibitem{journals/bib/RuttenbergRSM09}
Alan Ruttenberg, Jonathan Rees, Matthias Samwald, and M.~Scott
Marshall.
\newblock {Life Sciences on the Semantic Web: the Neurocommons and Beyond}.
\newblock {\em Briefings in Bioinformatics}, 10(2):193--204, 2009.

\bibitem{Sauer200458}
Uwe Sauer.
\newblock {High-throughput Phenomics: Experimental Methods for Mapping
  Fluxomes}.
\newblock {\em Current Opinion in Biotechnology}, 15(1):58--63, 2004.

\bibitem{citeulike:1882392}
Barry Smith, Michael Ashburner, Cornelius Rosse, et~al.
\newblock {The OBO Foundry: Coordinated Evolution of Ontologies to Support
  Biomedical Data Integration}.
\newblock {\em Nature Biotechnology}, 25(11):1251--1255, November 2007.

\end{thebibliography}
\bibliographystyle{plain}
\end{document}

### 2012-03-02 icadl2010/podd.tex
initial files from http://projects.arcs.org.au/svn/podd/trunk/doc/publications/

--- 2012-03-02 icadl2010/icadl_38/podd.tex
@@
 
-\newcommand{\gobble}[1]{}
-\begin{thebibliography}{10}
-
-\bibitem{citeulike:212874}
-M.~Ashburner, C.~A. Ball, J.~A. Blake, et~al.
-\newblock {Gene Ontology: Tool for the Unification of Biology}.
-\newblock {\em Nat Genet}, 25(1):25--29, May 2000.
-
-\bibitem{hoph03a}
-Ian Horrocks, Peter~F. Patel-Schneider, and Frank van Harmelen.
-\newblock {From $\mathcal{SHIQ}$ and {RDF} to {OWL}: The Making of a Web
-  Ontology Language}.
-\newblock {\em Journal of Web Semantics}, 1(1):7--26, 2003.
-
-\bibitem{dhoule09}
-David Houle.
-\newblock {Numbering the Hairs on Our Heads: The Shared Challenge and Promise
-  of Phenomics}.
-\newblock {\em {Proceedings of the National Academy of Sciences}}, October
-  2009.
-
-\bibitem{citeulike:1756058}
-Andrew~R. Jones, Michael Miller, Ruedi Aebersold, et~al.
-\newblock {The Functional Genomics Experiment model (FuGE): an Extensible
-  Framework for Standards in Functional Genomics}.
-\newblock {\em Nature Biotechnology}, 25(10):1127--1133, October 2007.
-
-\bibitem{rdfprimer04}
-F.~Manola and E.~Miller (editors).
-\newblock {{RDF Primer}}.
-\newblock \url{http://www.w3.org/TR/rdf-primer/}, February 2004.
-
-\bibitem{DBLP:conf/dlog/McGuinness03}
-Deborah~L. McGuinness.
-\newblock Configuration.
-\newblock In Franz Baader, Diego Calvanese, Deborah McGuinness, Daniele Nardi,
-  and Peter~F. Patel-Schneider, editors, {\em Description Logic Handbook},
-  pages 388--405. Cambridge University Press, 2003.
-
-\bibitem{nevo01}
-Eviatar Nevo.
-\newblock Evolution of genome-phenome diversity under environmental stress.
-\newblock {\em {Proceedings of the National Academy of Sciences}},
-  98(11):6233--6240, March 2001.
-
-\bibitem{sparql}
-Eric Prud'hommeaux and Andy Seaborne.
-\newblock {SPARQL Query Language for RDF}.
-\newblock \url{http://www.w3.org/TR/2006/CR-rdf-sparql-query-20060406/}, April
-  2006.
-
-\bibitem{irods07}
-A.~Rajasekar, R.~Moore, and F.~Vernon.
-\newblock {iRODS: A Distributed Data Management Cyberinfrastructure for
-  Observatories}.
-\newblock In {\em American Geophysical Union, Fall Meeting 2007}, December
-  2007.
-
-\bibitem{journals/bib/RuttenbergRSM09}
-Alan Ruttenberg, Jonathan Rees, Matthias Samwald, and M.~Scott
-Marshall.
-\newblock {Life Sciences on the Semantic Web: the Neurocommons and Beyond}.
-\newblock {\em Briefings in Bioinformatics}, 10(2):193--204, 2009.
-
-\bibitem{Sauer200458}
-Uwe Sauer.
-\newblock {High-throughput Phenomics: Experimental Methods for Mapping
-  Fluxomes}.
-\newblock {\em Current Opinion in Biotechnology}, 15(1):58--63, 2004.
-
-\bibitem{citeulike:1882392}
-Barry Smith, Michael Ashburner, Cornelius Rosse, et~al.
-\newblock {The OBO Foundry: Coordinated Evolution of Ontologies to Support
-  Biomedical Data Integration}.
-\newblock {\em Nature Biotechnology}, 25(11):1251--1255, November 2007.
-
-\end{thebibliography}
+\bibliography{all}
 \bibliographystyle{plain}

### 2012-03-02 icadl2010/slides/slides.tex
initial files from http://projects.arcs.org.au/svn/podd/trunk/doc/publications/

--- 2012-03-02 ibis_workshop_2010/slides.tex
@@
 
-\title[PODD -- Ontology-driven Repository]{PODD: An Ontology-driven Data Repository for Collaborative Phenomics Research}
+\title[PODD -- An Ontology-driven Repository]{PODD: An Ontology-driven Data Repository for Collaborative Phenomics Research}
 
@@
     }
-\vspace{+80pt}
-\date[IBS Workshop 2010]{\footnotesize IBS Phenomics Data and Informatics Workshop - Canberra 2010}
+
+\vspace{+380pt}
+\date[ICADL 2010]{\footnotesize $12^{th}$ International Conference on Asian Digital Libraries -- Gold Coast, Australia}
 
@@
         \begin{itemize}
-            \item APN: estimated 1.8TB + 30\% growth/year
+%            \item APN: estimated 1.8TB + 30\% growth/year
             \item TPA: 0.5TB/week $\sim 25$ TB/year
@@
         \pause
-        \item New platforms/processes/technologies emerge
+        \item New platforms/processes/technologies quickly emerge
         \pause
@@
 1. The volume of data. Phenomics research generates large amounts of
-data. For example, APN, the Australian Phenomics Network, is
-estimated to generate 1.8TB of data in the first year with 30\%
-growth. That is more than 7TB over 3 years. And APN is only one
-research center. We still have APPF, which operates 2 centers. The
-large volume makes the storage and distribution much more difficult.
+data. For example, The Plant Accelerator, one of the research centers
+in Australian Plant Phenomics Facility, is
+estimated to generate 25TB of data per year. And TPA is only one
+research center. Under APPF there is still the High-resolution Plant
+Phenomics Center. The Australian Phenomics Network is yet another center
+that works on mouse phenomics research and generates large amounts of data.
+The large volume makes the storage and distribution a challenging task.
 
@@
 technologies and processes emerge all the time. A data repository
-needs to be constructed in such a way that it can cater for future
-platforms and technologies.
+needs to be constructed in such a way that it can capture data and
+metadata off new platforms and technologies.
 
-3. Data needs context. 7TB of data isn't very useful if it just sits
-there and collects dust. To make use of the data, we need to
-surround them with context, the metadata, that describe the raw data
-in a meaningful way. Metadata can have dimensions, such as
-scientific and administrative. They can help researchers to discover
-data, integrate data and analyze data is a vital part of a data
-repository.
+3. Data needs context. 25TB of data isn't very useful if it just 
+sits there and collects dust. To make use of the data, we need to
+surround them with metadata that describe the raw data in a 
+meaningful way. There are different kinds of metadata, such as
+scientific and administrative. They can help researchers to 
+discover data, integrate data and analyze data is a vital part 
+of a data repository.
 
-There are quite a few databases for genomics and protein research,
+There are quite a few databases for genomics and proteomics research,
 such as Emsembl and UniProt. Unfortunately, not all the requirements
 for phenomics data management are met. For example, genomic data is
-relatively well defined, which isn't the case for phenomics.
-Phenomics data may come off from a variety of imaging and measuring
-platforms and the appropriate metadata needs to be captured.
+relatively well understood and structured, which isn't the case for 
+phenomics. Phenomics data may come off from a variety of imaging and 
+measurement platforms and the appropriate metadata needs to be captured.
 
@@
             \begin{itemize}
-                \item Biodiversity data collection \& management
+                \item Biodiversity information portal
             \end{itemize}
@@
 
-Australian Plant Phenomics Facility has two research centers.
+Australian Plant Phenomics Facility operates two research centers.
 
@@
 The High Resolution Plant Phenomics Center, the HRPPC, is located
-here in Canberra. The center focuses on deep phenotyping of plants
-using a number of equipments. Some of them operate on single plants
-in controlled environments, others operate on fields.
+in Canberra. The center focuses on deep phenotyping of plants
+using a variety of equipments. Some of them operate on single 
+plants in controlled environments, others operate in fields.
 
@@
 facility established to provides mouse models for human and animal
-diseases. For example, APN uses digital slide scanner to perform
+diseases. For example, APN uses digital slide scanners to perform
 histopathology and organ pathology imaging and analysis.
 
-The Atlas of Living Australia is developing a biodiversity data
-management system for Australia. Collaboration with ALA is also one
-of PODD's goals.
+The Atlas of Living Australia is developing a biodiversity 
+information portal for Australia. Collaboration with ALA is 
+also one of PODD's goals.
 
@@
 Chemical measurements & Chlorophyll content, Stomatal conductance \\
-Visual observation & Manual reports\\
+Visual observation & Manual reports (plant, mouse phenotypes)\\
 $\cdots$ & $\cdots$
@@
 \onslide<2>
-\begin{block}{Data generation}
+\begin{block}{Metadata capturing}
 \begin{tabularx}{\linewidth}{p{30mm}X}
@@
 \onslide<3>
-\begin{block}{Data management}
-\begin{tabularx}{\linewidth}{p{50mm}l}
-Data distribution & ? \\
-Data sharing & ? \\
-Data publishing & ? \\
-Access control & ? \\
-Archival \& versioning & ? \\
-Structure \& metadata & ? \\
-Data discovery \& analysis & ?
-\end{tabularx}
+\begin{block}{Data management tasks}
+\begin{itemize}
+\item Data distribution \& sharing
+\item Data publishing
+\item Access control
+\item Archival \& versioning
+%\item Structure \& metadata
+\item Data discovery \& analysis
+\item Data integration
+\end{itemize}
 \end{block}
@@
 One of the requirements of PODD is to support data capturing from
-various imaging platforms.
+various platforms.
 
 As you can see on this slide. there are quite a number of different
-ways of imaging and observing the subject. The challenge here is,
+ways of imaging and measuring a subject. The challenge here is,
 besides capturing the raw image files and observations, capture
@@
 
-Another requirement is data generation. Data and metadata can be
+Another requirement is managing the metadata. Metadata can be
 generated on different levels and at different times. All this
-information needs to be recorded in the repository.
+information needs to be recorded in the repository together with
+the data.
 
-Last but not least, there is a need for the effective management of
-data. Tasks such as distribution, publishing, access control and
-versioning belong to this category. As far as we know, there is no
-data repository that handles these tasks for phenomics research.
+The above requirements span across the whole range of data management
+tasks, including distribution, publishing, access control and versioning. 
+In the phenomics domain, the development of a repository that supports 
+these tasks need to take diversity and change into consideration.
 
@@
     \begin{itemize}
-        \item Supports data captured by different platforms
-        \item Supports different data formats
-        \item Supports effective data management
+        \item Acquisition and storage of large volumes of data
         \begin{itemize}
-            \item Metadata, distribution, access control, discovery, etc.
+        	%\item Different platforms \& different formats
+        	\item Distribution, access control, versioning, etc.
+	    \end{itemize}
+        \item Data conxtualization
+        \begin{itemize}
+			\item Logical organization
+        	\item Provenance tracking
+			\item Discovery \& integration
+        \end{itemize}
+        \item Prepare for change
+        \begin{itemize}
+        	\item Changes in domain model
         \end{itemize}
@@
     \begin{itemize}
-        \item Define domain ontologies -- focus of this talk
-        \item Design an appropriate architecture
-        \item Develop an ontology-aware repository
+        \item An ontology-driven approach
+        \item \textbf{Ontologies as the domain model}
+        \item Benefits: flexibility \& extensibility
     \end{itemize}
@@
 domain. Besides being a controlled vocabulary that define important
-concepts and relations, an ontology can define complex constraints
+concepts and relations, an ontology can also define complex constraints
 on these concepts and relations, making it a very expressive
@@
 lifecycles of the management of domain objects.
-
-\begin{frame}{PODD: The High-level Architecture}
- \vspace{-6mm}
- \begin{figure}[t]
-  \begin{center}
-  \includegraphics[trim=48mm 20mm 10mm 35mm, clip, width=130mm]{architecture.pdf}
-   \end{center}
-  \label{fig}
- \end{figure}
-\end{frame}
-
-Here you see the high-level architecture of the PODD repository. On
-the basis are a few components. Fedora Commons is the underlying
-repository that we use to store all the metadata information. iRODS
-is a federated data storage software that can be used by Fedora
-Commons. ANDS provides an iRODS-based service called Data Fabric,
-which we will use in PODD to host the data captured.
-
-Ontologies will be stored in Fedora Commons as well as in a RDF
-triple store to enable data discovery using SPARQL queries and full
-text search on metadata. We also use databases to store information
-about users, access rights and some other things.
-
-The data access layer works with ontologies and objects and is
-responsible for serializing and deserializing objects from the
-underlying storage.
-
-The various services and the web interface is guarded by a security
-layer that controls access based on fine-grained user access rights.
 
@@
         \item Can be extended to support phenomics
-        \item[$\cross$] May be difficult to extend for new concepts
+        \item[$\cross$] Defined in UML \& mapped to database schemas -- difficult to extend for new concepts
     \end{itemize}
@@
 genomics experiments. It models important domain concepts such as
-material, protocol and data. The model is stored as database
+material, protocol and data and how they relate to each other. 
+The model is defined as UML class diagrams and mapped to database 
 schemas, as a result, FuGe may be difficult to extend.
@@
 comprehensive but complex ontology. It defines domain concepts 
-starting from a very high abstract level. As a result, it is 
-quite big, containing about 26 hundred classes and 10 thousand
-axioms. It is also computationally expensive. For this reason and
-that we need to performing reasoning repeatedly in PODD, we chose
-not to use OBI as our domain ontology.
+starting from a very high abstract level. As a result, it is quite 
+big, containing about 26 hundred classes and 10 thousand logic
+axioms. It is also computationally expensive. We need to performing 
+reasoning repeatedly in PODD, we chose not to use OBI as our domain 
+ontology.
 
@@
     \item Precise, open \& extensible -- exactly what we need!
-    \item Provides core vocabularies for expressing complex ontologies -- \emph{data models}
-    \item Main language constructs
+    \item Provides core language constructs \& vocabularies for expressing complex ontologies -- \emph{data models}
+    \item APIs, query engines \& automated reasoners available
+\end{itemize}
+\end{block}
+\begin{block}{Fedora Commons}
 \begin{itemize}
-        \item Classes, e.g, \texttt{Human}
-        \item Predicates , e.g., \texttt{hasParent}
-        \item Individuals, e.g., \texttt{Aristotle}
-    \end{itemize}
-    \item APIs, query engines \& automated reasoners available
+	\item Mature open-source digital repository software
+	\item Modular \& extensible
+	\item Widely used 
 \end{itemize}
@@
 
+Fedora Commons is an open-source digital contents management system. It's designed
+to be modular and extensible so that it can be used with a variety of plugins
+and configurations. It's a very mature software and has been widely used 
+to implement institution-wide repository systems.
+
 \section{The PODD Ontology}
-\begin{frame}{The Ontology Approach}
-\begin{block}{Benefits}
-    \begin{itemize}
-    \setlength{\itemsep}{6pt}
-    \item<1-> Greater extensibility
-        \begin{itemize}
-        \item New concepts/relations can be easily added/modified
-        \end{itemize}
-    \item<2-> Better reuse \& integration
-        \begin{itemize}
-        \item Ontologies are \emph{open}
-        \item Other ontologies can be integrated on multiple levels
-        \end{itemize}
-    \item<3-> Balance between expressivity \& reasoning complexity
-        \begin{itemize}
-        \item Formal semantics enables automated query \& analysis
-        \item Off-the-shelf tools available
-        \end{itemize}
-    \end{itemize}
-\end{block}
-\end{frame}
-
-The ontology-based approach has a number of benefits.
-
-Firstly, the repository can be more extensible. In PODD we will store all
-object information in ontologies. In such a way new concepts and relations
-can be added more easily than if the information is stored in a database.
-
-An ontology-based repository is also better in terms of reuse and integration.
-Ontologies are designed to be shared so when you make your ontology available,
-you share your model. When you publish your instance data, concrete domain data
-is shared. So integration can happen on multiple levels.
-
-The OWL language is defined with a formal semantics. This enables automated
-query answering and analysis. OWL is designed with expressivity and reasoning
-complexity in mind so that the core reasoning tasks can be made fully automated. 
-As shown in a previous slide, although OWL is a relatively young language,
-there are off-the-shelf tools available to use.
-
 \begin{frame}{The PODD Ontology}
 \begin{block}{Modeling essentials}
-    \begin{itemize}
-    \item Domain entities
-        \begin{itemize}
-        \item Abstract concepts
-        \item Concrete objects
-        \end{itemize}
-    \item Domain entities defined using OWL \emph{ontologies}
-    \end{itemize}
-\end{block}
-\pause
-\begin{block}{Modeling in OWL}
+%    \begin{itemize}
+%    \item Domain entities
+%        \begin{itemize}
+%        \item Abstract concepts
+%        \item Concrete objects
+%        \end{itemize}
+%    \item Domain entities defined using OWL \emph{ontologies}
+%    \end{itemize}
+%\end{block}
+%\pause
+%\begin{block}{Modeling in OWL}
     \begin{itemize}
@@
     \item Comments, descriptions -- \emph{OWL annotations}
+    \end{itemize}
+\end{block}
+\pause
+\begin{block}{Benefits}
+	\begin{itemize}
+	\item Extensibility through inheritance
+	\item Reuse \& integration through \textbf{ontology mapping} \& \textbf{ontology annotation}
+		\begin{itemize}
+		\item Gene Ontology, Plant Ontology, etc.
+		\end{itemize}
 	\end{itemize}
@@
 
+%\begin{frame}{The Ontology-driven Approach}
+%\begin{block}{Benefits}
+%    \begin{itemize}
+%    \setlength{\itemsep}{6pt}
+%    \item<1-> Greater extensibility
+%        \begin{itemize}
+%        \item New concepts/relations can be easily added/modified
+%        \end{itemize}
+%    \item<2-> Better reuse \& integration
+%        \begin{itemize}
+%        \item Ontologies are \emph{open}
+%        \item Other ontologies can be integrated
+%        \begin{itemize}
+%        	\item Class, property \& individual \textbf{mapping}
+%			\item Ontological \textbf{annotations}
+%        \end{itemize}
+%        \end{itemize}
+%    \item<3-> Balance between expressivity \& reasoning complexity
+%        \begin{itemize}
+%        \item Formal semantics enables automated query \& analysis
+%        %\item Off-the-shelf tools available
+%        \end{itemize}
+%    \end{itemize}
+%\end{block}
+%\end{frame}
+
+The ontology-based approach has a number of benefits.
+
+Firstly, the repository can be more extensible. In PODD abstract
+domain concepts are stored in the repository as well. After we've 
+defined the domain ontology, concepts and properties can be extended 
+to create new ones. In such a way new concepts and relations
+can be added more easily than if the information is stored in a database.
+
+Note here that concept and object definitions are all versioned so
+we can freely modify an concept without affecting existing objects.
+
+An ontology-based repository is also better in terms of reuse and integration.
+Ontologies are designed to be shared so when you make your ontology available,
+you share your model. When you publish your instance data, concrete domain data
+is shared. So integration can happen on multiple levels.
+
+\begin{frame}{The PODD Ontology -- Overview}
+\vspace{-12pt}
+ \begin{figure}[t]
+  \begin{center}
+  \includegraphics[trim=48mm 30mm 38mm 20mm, clip, width=112mm]{model.pdf}
+   \end{center}
+  \label{fig:ont}
+ \end{figure}
+\end{frame}
+
+In this slide we show how some high-level concepts are logically
+connected in the PODD domain model. Project is the top-level concept
+with a bunch of attributes. It is also connected to some other
+concepts like ProjectPlan, Platform and Investigation. Each of these
+concepts also should have a number of core metadata attributes.
+
+The concepts here are organized in a tree-like structure. What is not
+shown on this slide is cross references between concepts. They are
+useful when a particular object is referred to by some other objects.
+
 \begin{frame}{The PODD Ontology -- An Example}
@@
     \item The top-level concept
-    \item Constraints on inter-object relations \& attributes -- like everything else
+    \item Constraints on inter-object relations \& attributes
     \end{itemize}
@@
             &\sqsubseteq~ \geq~ 1~ hasInvestigation \sqcap \forall hasInvestigation.Investigation\\
-            &\sqsubseteq~ =~ 1~ hasStartDate \sqcap \forall hasStartDate.date\\
+            &\sqsubseteq~ =~ 1~ hasStartDate \sqcap \forall hasStartDate.\textit{xsd:date}\\
             &\sqsubseteq~ \leq~ 1~ hasPublicationDate \sqcap \forall hasPublicationDate.date
@@
 
-In this slide I show a simple, incomplete modeling of the top-level concept
+In this slide I show an incomplete modeling of the top-level concept
 in our ontology, the Project. In our model of phenomics, everything starts 
@@
 
-As you will see, we define constraints on attributes and inter-concept relations
-of the Project concept. This is how we constrain all of our concepts.
+As you see, we define constraints on attributes and inter-concept relations
+of the Project concept. This is how we model all of our concepts.
 
@@
 
-\begin{frame}{The PODD Ontology}
-\onslide<2->
-\begin{columns}[t]
-\column{.5\textwidth}
-\begin{itemize}
-    \item Inspired by FuGe \& OBI
-    \item Aim: define all \emph{essential} domain concepts, attributes \& relations
-    \item $\sim30$ classes, \\$\sim 80$ predicates, \\$\sim 200$ Axioms
-    \item Faster reasoning, querying, etc.
-\end{itemize}
-\vfill
-\onslide<1->
-\column{.5\textwidth}
-\vspace{-10mm}
- \begin{figure}[t]
-  \begin{center}
-  \includegraphics[trim=3mm 26mm 2mm 8mm, clip, height=72mm]{hier.pdf}
-   \end{center}
- \end{figure}
-\end{columns}
-\end{frame}
+%\begin{frame}{The PODD Ontology}
+%\onslide<2->
+%\begin{columns}[t]
+%\column{.5\textwidth}
+%\begin{itemize}
+%    \item Inspired by FuGe \& OBI
+%    \item Aim: define all \emph{essential} domain concepts, attributes \& relations
+%    \item $\sim30$ classes, \\$\sim 80$ predicates, \\$\sim 200$ Axioms
+%    \item Faster reasoning, querying, etc.
+%\end{itemize}
+%\vfill
+%\onslide<1->
+%\column{.5\textwidth}
+%\vspace{-10mm}
+% \begin{figure}[t]
+%  \begin{center}
+%  \includegraphics[trim=3mm 26mm 2mm 8mm, clip, height=72mm]{hier.pdf}
+%   \end{center}
+% \end{figure}
+%\end{columns}
+%\end{frame}
 
@@
 \begin{frame}{PODD Ontology -- Roles}
-\begin{block}{Ontologies \emph{drives} repository functions}
+\begin{block}{Ontologies \textbf{\emph{drive}} repository functions}
 \begin{list}{}{}
-\item[\textcolor{blue}{\textbf{Presentation}}] Drives the rendering of the web pages
+\item[\textcolor{blue}{\textbf{Presentation}}]~\\% Drives the rendering of the web pages
     \begin{itemize}
@@
     \end{itemize}
-\item[\textcolor{blue}{\textbf{Management}}] Object (de)serialization to/from ontologies
-\item[\textcolor{blue}{\textbf{Validation}}] Ontology reasoning performed
+\item[\textcolor{blue}{\textbf{Storage}}] ~\\
 	\begin{itemize}
-    \item Validation on object type, cardinality, etc.
+	\item Object (de)serialization to/from ontologies
 	\end{itemize}
-\item[\textcolor{blue}{\textbf{Discovery}}] Multiple ways of finding information
+\item[\textcolor{blue}{\textbf{Validation}}] ~\\%Ontology reasoning performed
+    \begin{itemize}
+    \item Validation based on concept constraints
+    \end{itemize}
+\item[\textcolor{blue}{\textbf{Discovery}}] ~\\%Multiple ways of finding information
     \begin{itemize}
     \item Queries using SPARQL
-    \item Searches using Lucene
+    \item Full-text search
     \end{itemize}
@@
 
-\begin{frame}{The PODD Ontology -- Object creation page rendering}
+\begin{frame}{The PODD Repository: The High-level Architecture}
+\vspace{-10pt}
  \begin{figure}[t]
   \begin{center}
-  \includegraphics[width=76mm]{create.png}
+  \includegraphics[trim=48mm 30mm 38mm 20mm, clip, width=96mm]{architecture.pdf}
    \end{center}
+  \label{fig:arch}
  \end{figure}
 \end{frame}
+
+Here you see the high-level architecture of the PODD repository. 
+We take a layered approach in designing and developing the system.
+
+The data access layer at the bottom is responsible for 
+serializing and deserializing concepts and objects in 
+the underlying storage. Objects and their metadata are
+stored in Fedora Commons, a very widely-used digital
+contents repository software system. iRODS is a federated 
+data storage software that can be plugged into Fedora
+Commons as a storage module. We use iRODS to store the
+Fedora datastreams.
+
+Object definitions in RDF will be stored in a Sesame triple 
+store and a Lucene index so that we can query and search for
+information. We also use databases to store information
+about users, access rights and some other things.
+
+The business logic layer is responsible for managing the 
+lifecycles and interactions of objects and concepts. OWL
+reasoning is a central component of this layer as we use
+ontologies in all sorts of data management tasks.
+
+On top of that is the security layer, which handles authentication
+and authorization. We have a fine-grained authorization module
+where access rights are granted on a per-user and per-project basis.
+
+At the top of the stack is the interface layer. PODD is a web-based
+system so all functionality can be accessed through a browser as
+well as through API calls.
 
@@
     \begin{itemize}
-    \item Emerging processes, platforms, technologies require extensible infrastructure
+    \item Emerging processes, platforms, technologies require a extensible conceptual framework
     \end{itemize}
-\item An ontology-driven approach as the foundation of PODD
+\item An ontology-driven architecture as the foundation of PODD
     \begin{itemize}
+    \item Ontologies as the domain model
     \item Extensible \& open
@@
 
-This slides shows a screen capture of the project creation page.
-Note that most of these fields are generated by reading the domain
-ontology and figuring out what fields are needed and what their types,
-cardinalites are.
-
-For example, in a previous slide we defined that a project must have 
-exactly 1 start date attribute and it must be a date. So here we show 
-an input box for this attribute and signal that it is mandatory. Note
-that lead institution is also a mandatory attribute while the duration
-is not. When this project object is saved, the values are checked against
-the domain ontology and any errors or omissions will be reported back
-to the user.
+As a result, in PODD we propose an ontology-driven architecture
+for developing extensible data repositories. In this framework
+databases take a back seat and ontologies are used to describe
+the domain. We believe it is more extensible, flexible and open.
 
@@
 \begin{itemize}
-\item PODD ontology stabilizing
+\item PODD ontology for phenomics research
 \item Development of basic repository functionality
@@
 \item Development of object discovery services
-\item Integrate with AAF federated authentication services
+\item Integration with Shibboleth authentication 
 \item \textcolor{gray}{Exposing data for discovery}
@@
 \end{frame}
+
+So far we have developed an ontology for plant and
+mouse phenomics research. Centered around this ontology
+we have implemented basic repository functionalities
+including object storage, versioning, validation, 
+publishing, searching and presentation generation.
+
+In future we will implement a number of tasks that are
+important for phenomics research. These include data
+import and export, object discovery and more flexible
+authentication with Shibboleth.
+
+We will also be looking at how we can expose data in
+the PODD repository to external world and how they 
+can be integrated with existing datasets.
 
@@
 \begin{block}{Acknowledgment}
-Faith Davies, Gavin Kennedy, Jane Hunter\\
-eResearch Lab, Schoo of ITEE, UQ
+Faith Davies, Gavin Kennedy, Jane Hunter @
+eResearch Lab, School of ITEE, UQ
 \end{block}

### 2012-03-02 iswc2011_demo/podd.tex
initial files from http://projects.arcs.org.au/svn/podd/trunk/doc/publications/

\documentclass{llncs}

\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{zed-cm}
\usepackage{booktabs}
\usepackage{color}
\usepackage{alltt}
\usepackage{boxedminipage}
\usepackage{graphicx}

\title{PODD -- An End-to-end Semantic Data Management System}

\author{Yuan-Fang Li$^\dagger$, Gavin Kennedy$^{\S,\sharp}$, Faith Ngoran$^\S$, Philip Wu$^\P$, Jane Hunter$^\S$}

\institute{
$^\dagger$Clayton School of IT, Monash University\\
$^\S$School of ITEE, The University of Queensland\\
$^\sharp$High Resolution Plant Phenomics Centre, Canberra\\
$^\P$The John Curtin School of Medical Research, Australian National University%\\
%Contact: \email{\texttt{yuanfang.li@monash.edu}}
}

\begin{document}
\maketitle

\begin{abstract}
The massive amounts of data and the rapid evolution of data models contribute to making scientific data management a very challenging task. In this paper we present PODD, an ontology-based data management system. In PODD, ontologies are used to completely capture the definition and relationship of domain concepts and object. The open and semantic nature of ontologies makes PODD extensible and amenable for integration and reuse.
\end{abstract}

\section{Introduction}\label{sec:intro}
Data management has become an indispensable part of modern scientific research~\cite{1107503}. Among other factors, there are three core challenges for the effective data management in scientific research. (1) The ability to provide a data management system (DMS) that can manage large quantities of heterogeneous data. (2) The ability to support metadata-related services to provide context and structure for data within the DMS to facilitate effective search, query and dissemination. (3) The ability to accommodate evolving and emerging knowledge, technologies and processes.

Database-based DMS have traditionally been applied successfully to manage
research data. In such approaches, usually the domain model needs to stay relatively stable as database extension and migration is
often an error-prone and laborious task. Consequently, this approach is
not suitable for domains where data and model evolution is the norm
rather than the exception.

Ontology language OWL is both open and semantically precise. It has been successfully used widely a number of domains for knowledge representation and data integration~\cite{journals/bib/RuttenbergRSM09}. In this paper, we present our work in designing PODD (Phenomics Ontology Driven Data Management)~\cite{Li2011}, a semantic, domain-independent DMS system that takes advantage of the intrinsic characteristics of OWL. The
ontology-based domain model is at the core of PODD as it defines the
logical hierarchy of data in scientific experiments.
Data integrity is therefore maintained via ontology reasoning.

\section{The PODD System Achitecture \& Main Functionality}\label{sec:arch}
A most distinguishing characteristic of PODD is its flexibility, which is exhibited in two ways. \emph{Firstly}, the system can be readily applied to another domain with just a change of the underlying domain model (ontologies). \emph{Secondly}, existing concepts and objects can evolve independently without affecting the definition or integrity of each other.

Such flexibility is achieved by defining domain entities using OWL ontologies, in which: (1)
OWL classes represent domain concepts; (2) OWL properties model concept
attributes and their relationships; (3) OWL restrictions specify
constraints on concepts; (4) OWL individuals define concrete
domain objects where attributes and relationships are defined using OWL
assertions, and finally (5) OWL annotations express auxiliary information. Raw data files are attached to concrete domain objects.

\vspace{-12pt}
\begin{figure}[htb]
\centering
\includegraphics[trim = 64mm 15mm 52mm 28mm, clip,width=64mm]{podd_arch.pdf}
\vspace{-16pt} \caption{The high-level architecture and components in PODD.}\label{fig:arch}
\end{figure}

\vspace{-8pt}
The high-level design of PODD takes a modular and layered approach, as can be seen in Figure~\ref{fig:arch}. The foundation \textbf{data access layer} is responsible for low-level tasks (creation, modification and deletion of concepts and objects), consisting of an underlying repository system (Fedora Commons), an RDF triple store (OpenRDF Sesame), an in-house database
that stores essential information and a full-text search engine (Apache Lucene). The \textbf{business logic layer} in the middle is responsible for managing
concepts and objects, such as versioning, object conversion and
integrity validation. The \textbf{security layer} controls access
(authentication and authorization) to concepts and objects and guards
all operations on them. Secure channels (HTTPS) and Shibboleth\footnote{\url{http://shibboleth.internet2.edu/}} integration have also beeb developed for more secure and flexible access to the system.
At the top of the stack is the \textbf{interface layer}, where the data
management system can be accessed using a number of interfaces such as
a Web browser or API calls.

Data can be deposited in PODD through the Web interface and Web Services. Information stored in PODD can be retrieved in a number of ways: (1) manual browsing (shown in Figure~\ref{fig:browser}), (2) SPARQL queries, (3) Lucene full-text search, (4) Web Services through REST-style APIs and (5) automated harvesting through RDFa and OAI-PMH annotations.

\section{Ontology-based Domain Modeling}\label{sec:ont}
In our modeling approach, ontologies are developed on multiple levels. The base ontology defines domain-independent concepts and their inter-relationships with each other. Based on this ontology, other more domain-specific ontologies are developed hierarchically. To describe knowledge in phenomics, for example, we extend the base ontology
by defining additional concepts including \emph{Genotype}, \emph{Gene},
\emph{Phenotype} and \emph{Sequence} as subclasses of
existing ones. Additional OWL object and datatype properties are
also defined to model the attributes and relationships of these
concepts, as shown in Figure~\ref{fig:podd_high}. These ontologies play a vital role in all major data management life cycles, including ingestion, retrieval, update and publication.

\vspace{-20pt}
\begin{figure}[htb]
\centering
\includegraphics[trim = 2mm 0mm 2mm 0mm, clip,width=\textwidth]{ont_podd.pdf}
\vspace{-16pt} \caption{Extended Domain Ontology for
phenomics.}\label{fig:podd_high}
\end{figure}

\vspace{-32pt}
\section{PODD Deployment}\label{sec:podd}
PODD has been successfully deployed at two different Australian research institutes to satisfy their data management needs. An instance of PODD has been deployed in production for the Australian Plant Phenomics Facility (APPF)\footnote{\url{http://podd.plantphenomics.org.au/}}. Another instance has been deployed in production for the Australian Phenomics Network (APN)\footnote{\url{http://podd.australianphenomics.org.au/}} for mouse phenomics research. The source code of the PODD repository software has been published in this repository\footnote{\url{http://podd.plantphenomics.org.au/podd/object/poddObject:696}}. Such deployments demonstrates the versatility of the system.

\vspace{-8pt}
\begin{figure}[htb]
\centering
\includegraphics[trim = 5mm 0mm 5mm 0mm, clip,width=\textwidth]{browser.png}
\vspace{-8pt} \caption{The browser view of a plant project in a PODD
DMS.}\label{fig:browser}
\end{figure}

%\vspace{-24pt}
\section{Conclusion}\label{sec:concl}
In this demonstration, we present the PODD data management system. Modelling domain entities entirely in OWL ontologies, PODD takes advantage of the intrinsic semantic rigor and openness of the OWL language. Therefore, PODD is characterized by great flexibility and extensibility. It is also amenable to semantic integration and interoperability.

\section*{Acknowledgment}
This work is part of a National eResearch Architecture Taskforce (NeAT) project, supported by the Australian National Data Service (ANDS) through the Education Investment Fund (EIF) Super Science Initiative, and the Australian Research Collaboration Service (ARCS) through the National Collaborative Research Infrastructure Strategy Program.

\bibliography{all,liyf}
\bibliographystyle{abbrv}
\end{document}

### 2012-03-02 jws/cover.tex
initial files from http://projects.arcs.org.au/svn/podd/trunk/doc/publications/

\documentclass{letter}

%\usepackage[top=40mm, bottom=40mm, left=30mm, right=16mm]{geometry}
\usepackage[top=40mm]{geometry}

\signature{Yuan-Fang Li}
\address{School of ITEE\\The University of Queensland}

\begin{document}

\begin{letter}{Editors, Special Issue on Semantic Web Dynamics\\Journal of Web Semantics}

\opening{Dear Editors:}

I am writing in response to the call for papers of the Special Issue
on Semantic Web Dynamics in the Journal of Web Semantics. With this
letter I submit a manuscript titled ``An Ontology-driven
Architecture for Extensible Scientific Data Management Systems'' for
your consideration.

This paper describes our work in designing an ontology-driven
architecture for developing extensible data management systems
to support dynamic domain models. We also describe our
efforts in developing the PODD system based on this
architecture as a demonstration of the practicability of the
proposed architecture.

Please note that this manuscript is based on an extended
abstract accepted at the 12th International Conference on
Asian Digital Libraries (ICADL 2010), titled ``PODD: An
Ontology-driven Data Repository for Collaborative
Phenomics Research''. Compared to the extended abstract,
the current manuscript contains substantial new contents,
including:

\begin{itemize}
\item A more comprehensive survey of relevant works.

\item The proposal of the domain-independent
ontology-driven architecture for developing
data management systems.

\item A more detailed description of the base ontology
and its roles in the data management system.

\item The separation of the domain-independent
base ontology and the phenomics-specific ontology
and the description of how they are integrated in
the PODD system.

\item A more detailed description of the PODD system.
\end{itemize}

Thank you in advance for reviewing our paper. Please
do not hesitate to contact me should you have any
queries.

\closing{Sincerely,}

\end{letter}
\end{document}

### 2012-03-02 jws/podd.tex
initial files from http://projects.arcs.org.au/svn/podd/trunk/doc/publications/

\documentclass{elsarticle}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{boxedminipage}

\setcounter{tocdepth}{3}
\def\keywordname{{\textbf{Keyword:}}}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\journal{Journal of Web Semantics}

\begin{document}

\begin{frontmatter}

\title{An Ontology-driven Architecture for Extensible Scientific Data Management Systems}

\author{Yuan-Fang Li\corref{cor1}}
\ead{uqyli4@uq.edu.au}
\author{Gavin Kennedy}
\ead{g.kennedy1@uq.edu.au}
\author{Faith Davies}
\ead{f.davies@uq.edu.au}
\author{Jane Hunter}
\ead{j.hunter1@uq.edu.au}

\cortext[cor1]{Author for correspondence}

\address{School of ITEE, The University of Queensland\\
GP South, Staff House Road, St Lucia 4072, Australia}

\begin{abstract}
Data management has become a critical challenge faced by a wide
array of scientific disciplines in which the provision of sound data
management is pivotal to the success and impact of research
projects. The huge and rapidly growing amounts of data to be managed
and the fact that the \emph{models} of data evolve over time
contribute to making data management an increasingly complex
undertaking that warrants a rethinking of its design. We believe
that a number of intrinsic characteristics of Semantic Web ontology
languages OWL and RDFS, such as semantic rigor and the extensible
nature, make them an ideal conceptual platform on which effective
data management systems can be developed. In this paper, we present
our efforts in designing an open and extensible architecture with an
ontology at its core. In this architecture, the behaviors of domain
concepts and objects are captured entirely by ontological entities,
around which all data management tasks are carried out. Moreover,
the open and semantic nature of ontology languages also makes such
systems amenable to greater data reuse and interoperability. To the
best of our knowledge, this is the first proposal of an
ontology-driven architecture for extensible data management systems.
An ideal domain for applying these principles is phenomics, the
systematic study of phenotypes of organisms. Phenomics research
generates high volumes of heterogeneous data and makes use of
emerging imaging and measurement technologies and processes, thus
making it an ideal testbed for data management systems. In this
context, we describe the development of PODD, a data management
system for phenomics research, as a step towards validating the
practicality of the ontology-driven architecture.
\end{abstract}

\begin{keyword}
Ontology-driven architecture \sep OWL \sep data management systems \sep PODD
\end{keyword}

\end{frontmatter}

\section{Introduction and Background}\label{sec:intro}
Data management is the practice of managing (digital) data
and resources, encompassing a wide range of
activities including acquisition, storage, retrieval, discovery,
access control, publication, integration, curation and archival.
For many data-intensive scientific disciplines such as life sciences and
bioinformatics, sound data management informs and enables
research and it has become an indispensable component~\cite{1107503}.

The need for effective data management is, in a large part, due to
the fact that huge amounts of digital data are being generated by
modern instruments.
% some stats/citations?
Furthermore, the fast evolution of technologies/processes and
discovery of new scientific knowledge require flexibility in
handling dynamic data and models in data management systems.

Database systems have traditionally been used successfully
to manage research data~\cite{brm2007}. Database schemas are
used as domain models to capture the attributes and relationships
of domain concepts. One implication of the above approach is
that domain models need to stay relatively stable as database
extension or migration is often an error-prone and laborious task
in practice. Hence, we believe that this approach is not suitable
for domains where data and model evolution is the
norm rather than the exception.

Semantic Web ontology languages such as RDF
Schema~\cite{rdfschema04} and OWL~\cite{hoph03a} possess
expressive, rigorously-defined
semantics and non-ambiguous syntaxes. Moreover, they have been
designed to be open and extensible to support knowledge and data
exchange on the Web scale~\cite{linkeddata,aue07dbpedia}.
These intrinsic characteristics make them an ideal conceptual
platform on which a flexible data management system that supports
dynamic data and models can be built.

Ontology language OWL has been widely used in a number of domains,
notably in life sciences and
biotechnology~\cite{journals/bib/RuttenbergRSM09,citeulike:1882392,citeulike:212874}
as a modeling language for its expressivity and extensibility. There
is also growing tool support for tasks such as reasoning, querying
and visualization, making it a viable option for the modeling and
representation of scientific domain concepts.

Moreover, with the rapid progression of Semantic Web-based
data integration through the community-driven Linked Data
project~\cite{citeulike:5008761},
it is advantageous for data management systems to support
Semantic Web languages and standards natively to benefit
from the fast-increasing, integrated open datasets.

In this paper, we present our work in designing
a domain-independent, ontology-driven architecture for scientific
data management systems. In our architecture, we support data and model
dynamics through ontology-based domain modeling. In this architecture,
the ontology is at the core of the system where the behaviors of
abstract domain concepts and concrete domain objects are entirely
defined with ontological vocabularies. Logical structure of
data is therefore maintained and enforced via ontological
definitions and reasoning and not via database schemas and associated
constraints.

We also describe the modeling of the base ontology in
this architecture and show how the domain can be
conceptualized in an ontology in an extensible and
flexible way.

Based on such a design, we have developed the Phenomics Ontology
Driven Data (PODD) data management system~\cite{podd_icadl} that
supports the effective management of large-scale, evolving data for
phenomics research. An example of concept change is the concept
\emph{GrowthCondition} in phenomics, which captures the
environmental conditions in which an organism grows. For different
disciplines and projects, it can be specialized to
\emph{CabinetCondition} to capture conditions in controlled
environments and \emph{FieldCondition} to capture conditions in the
field.

The ontology-based domain model is at the core of PODD as it drives
the creation, storage, validation, query and search of data and
metadata. In contrast to traditional data management systems that
use database schemas as the underlying model, the employment of OWL
ontologies as the domain model makes PODD highly extensible.

In biology, an organism's phenotype is its observable or
quantifiable trait, such as the height of a plant or the size of a
leaf, as a consequence of its genetic makeup combined with its
developmental stage, environment and disease conditions. Phenomics
is the systematic and comprehensive study of an organism's phenotype
and is determined through a combination of high-throughput and
high-resolution imaging- and measurement-based analysis platforms.
Phenomics research, together with genomics research, represents a
holistic and promising new approach to biological
study~\cite{dhoule09,Sauer200458,nevo01,plant_furbank}.

Unlike genomics, phenomics research emphasizes the study of
physical, observable traits of organisms. Like genomics, vast
amounts of data are produced by imaging and measurement platforms
and analysis tools. The effective storage, management, analysis and
discovery of these data is a challenging problem. Specifically,
there are three key challenges for data management in phenomics
research.
\begin{itemize}
\item The ability to provide a data management service that can
manage large quantities of heterogeneous data in multiple formats
(text, image, video) and not be constrained to a finite set of
imaging and measurement platforms and data formats.

\item The ability to support metadata-related services to
provide context and structure for data within the data management
service to facilitate effective search, query and dissemination.

\item The ability to accommodate evolving and emerging technologies
and processes as phenomics is still a rapidly developing field of
research.
\end{itemize}

PODD has been developed to meet the above challenges facing
the Australian phenomics research community by providing
efficient and flexible repository functionalities for large-scale
phenomics data, and providing a mechanism
for maintaining structured and precise \emph{metadata} around the
raw data so that they can be stored, distributed and published
in a reusable fashion. We would like to emphasize that although
the PODD system is geared
towards phenomics research, the ontology-driven architecture
we propose in this system is actually domain-independent and can be
applied in any scientific discipline where research output can
be conceptually organized in a structured manner.
%We will describe the PODD system and show how the ontology-based
%architecture is realized through leveraging mature open-source
%technologies. In the implementation of the PODD system, we have
%made some careful design decisions, which will be documented
%in the paper as well.

%In this paper, we present our work in addressing the above
%challenges and highlight the OWL-based modeling approach we take.
The rest of the paper is organized as follows. In
Section~\ref{sec:overview} we present related work and give a
brief overview of the motivation and goals of the PODD
project. Section~\ref{sec:design} presents in detail the
ontology-based architecture for data management systems.
In Section~\ref{sec:ont}, in a bioinformatics setting,
we discuss the PODD domain ontology in more detail and
show how the ontology-based modeling
approach is used in the life cycle of domain objects.
In Section~\ref{sec:podd}, we describe the implementation
of the PODD data management system. Finally,
Section~\ref{sec:conclusion} concludes the paper and identifies
future directions.

\section{Related Work, Motivation \& Goals}\label{sec:overview}
Over the years attempts have been made to develop content repository
systems and architectures to meet institutional and personal
data management needs. In this section, we introduce a number of
such systems and architectures. With a survey of related work,
we present the motivation behind the ontology-driven architecture
and the goals we wish to achieve with the PODD data management system.

\subsection{Data and Resource Management Systems}
A number of open-source content repository specifications and
software systems have been developed.

Fedora Commons\footnote{\url{http://www.fedora-commons.org/}} is an open-source
digital resource management system based on the principles of modularity,
interoperability and extensibility. In Fedora Commons, abstract concepts are defined
as \emph{models}, on which inter-relationships and behaviors
can be further defined. Data in Fedora Commons repositories are
organized into \emph{objects}, which have \emph{datastreams} that stores
either metadata or data. Fedora Commons makes heavy use of Semantic Web
technologies through the use of common RDF vocabularies and the integration
with the Mulgara triple store\footnote{\url{http://www.mulgara.org/}},
which can be used for metadata storage and query through SPARQL.
Fedora Commons has been used as a backend to implement document
management systems, digital libraries and institutional repositories.

Apache Jackrabbit\footnote{\url{http://jackrabbit.apache.org/}} is an
open-source implementation of the Content Repository for Java Technology
(JCR) API\footnote{\url{http://jcp.org/en/jsr/detail?id=283}}. In JCR,
data is stored in a tree of \emph{nodes}, which can hold \emph{properties}
of arbitrary values, which is conceptually similar to Fedora Commons.
Types can be defined on nodes to place certain restrictions on them.

Fedora Commons and JCR both support fairly basic mechanisms for
defining object relationships. Hence, they are usually used as the
underlying repository solution on which complex data and document
management systems are built. These systems include Biodiversity
Heritage Library\footnote{\url{http://www.biodiversitylibrary.org/}}
and Fez\footnote{\url{http://fez.library.uq.edu.au/}}, among others.
As stated previously, these systems use database schemas as their
domain models.

Data management systems have also been developed to support a number
of scientific disciplines including high-energy physics, bioinformatics
and Earth observation.

Bioinformatics Resource Manager (BRM)~\cite{brm2007} is one example of
client-server style data management software for bioinformatics
research. The client software is installed on users' computers to access
(microarray and proteomic) resources stored on BRM server in a
PostgreSQL relational database.
The BRM server supports data acquisition from external sources such as
NCBI~\cite{DBLP:journals/nar/WheelerBBBCCCDEFGHKKKLMMOPSSSSSSSSTTWY06}
and UniProt~\cite{AmosBairoch01012005}. It also supports annotation
using public datasets and connectivity to analytics tools. Data in
BRM is stored under the \emph{Project} concept and is mostly flat, i.e.,
it does not support hierarchical domain concepts such as investigation
and publication.

Besides data management systems, grid-based middleware
systems have also been developed to provide distributed
storage solutions. Such systems include the Storage
Resource Broker (SRB)~\cite{783165} and the CERN Data Grid~\cite{652836}
and other systems that make use of Globus\footnote{\url{http://www.globus.org/}}
middleware. These systems store data in a distributed environment
and usually support authentication, replication, redundancy, etc.
However, they are mostly only concerned about data storage and
replication and hence do not provide full-fledged data management
capabilities. Interested readers can see~\cite{565296}
for a detailed survey of grid resource management systems.

\subsection{Domain Modeling in Scientific Research}
A number of specifications and ontologies have been proposed to
model scientific research activities. In 2004, Council for the
Central Laboratory of the Research Councils (CCLRC) of UK developed
a CCLRC Scientific Metadata Model~\cite{Spallation_metadatamodel}
that models scientific activities in free text. An OWL ontology,
EXPO, was developed~\cite{citeulike:3735746} to capture metadata
about scientific experiments. EXPO was developed in a top-down
approach by extending concepts in the Suggested Upper Merged
Ontology (SUMO)\footnote{\url{http://www.ontologyportal.org/}}.
Although very comprehensive, these models are very verbose and not
very suitable as a model for developing data management systems.

In biological and particularly 'omics research,
a large number of databases have been
developed to host a variety of information such as genes
(Ensembl\footnote{\url{http://www.ensembl.org/}}), proteins
(UniProt\footnote{\url{http://www.uniprot.org/}}), scientific
publications (PubMed\footnote{\url{http://www.ncbi.nlm.nih.gov/pubmed/}}) and
microarray data (GEO\footnote{\url{http://www.ncbi.nlm.nih.gov/geo/}}).
These databases are generally characterized by the fact that they
specialize in a particular kind of data (protein sequences,
publications, etc.) and their conceptual domain models,
such as gene and gene products~\cite{citeulike:212874}
and microarray experiments~\cite{citeulike:151946},
are well understood.

For a scientific data management system to be effective, models of
domain concepts need to be integrated with models of scientific
activities and workflows. However, models of biological and clinical
investigations are less well understood.

The Ontology for Biomedical Investigations
(OBI)\footnote{\url{http://purl.obolibrary.org/obo/obi}} is an
ongoing effort of developing an integrative ontology for biological
and clinical investigations. It takes a top-down approach by reusing
high-level, abstract concepts from other ontologies. It includes
2,600+ OWL~\cite{hoph03a} classes and 10,000+ axioms (in the import
closure of the OBI ontology). Although OBI is very comprehensive,
its size and complexity makes reasoning and querying of OBI-based
ontologies and RDF graphs computationally expensive and time
consuming, making it impractical as a domain model for a data
management system where such reasoning may need to be performed
repeatedly.

Functional Genomics Experiment Model (FuGe)~\cite{citeulike:1756058}
is an extensible modeling framework for high-throughput functional
genomics experiments, aiming at increasing the consistency and
efficiency of experimental data modeling for the molecular biology
research community. Centered around the concept of experiments, it
encompasses domain concepts such as protocols, samples and data.
FuGe is developed using UML from which XML Schemas and database
definitions are derived. The FuGe model covers not only
biology-specific information such as molecules, data and
investigation, it also defines commonly used concepts such as audit,
reference and measurement. Extensions in FuGe are defined using
inheritance of UML classes.

We feel that the extensibility we require is not met by FuGe as any
addition of new concepts would require the development of new
database schemas and code. Moreover, the concrete objects reside in
relational databases, making subsequent integration and
dissemination more difficult.

\subsection{The PODD Data Management System -- Motivation \& Goals}
Phenomics is a fast-growing, data-intensive discipline
with new technologies and processes rapidly emerging
and evolving. As a result, its domain model and data
management systems must also be able to evolve to handle
the complexity, change and scale.

The PODD system is being developed on behalf of two major phenomics
initiatives: the Australian Plant Phenomics Facility (APPF),
specializing in phenotyping crop and model plant species; and the
Australian Phenomics Network (APN), specializing in the
phenotyping of mouse models. Both facilities have common
requirements to gather and annotate data from both high- and
low-throughput phenotyping devices. The scale of measurement can be
from the micro or cellular level, through the level of a single
organism, and up to (in the case of the APPF) the macro or field
level. Imaging, measurement and analysis of organisms on such a large
scale will produce an enormous amount of data.
For example, it has been estimated that more than
60 TB of image data will be generated by APPF research centers
every year, which need to be managed effectively.

Phenomics research makes use of a large variety of imaging and
measurement platforms. For example, in mouse histopathology and
organ pathology research, the Zeiss ``Mirax Scan'' scanner is used
to scan microscope slides. In clinical pathology, a Flow Cytometer
is used to capture laser diffraction images of blood samples. In
plant research, the Lemnatec Scanalyzer is used to capture RGB
images of plants in growth cabinets. The Fluorogroscan system is
used in quenching analysis: the partitioning of light energy used in
photosynthesis on model plants such as Arabidopsis. Other devices,
such as Infrared Thermography Camera to capture leaf temperature and
SPAD Meter to measure the chlorophyll content of plant leaves, are
also used. New devices and instruments will also be used when they
become available. Moreover, existing instruments may be upgraded so
that they can capture more information. The PODD domain model needs
to be flexible to accommodate these changes.

Because an organism's phenotype is often the product of the organism's
genetic makeup, combined with its development stage, disease conditions
and its environment, any measurement made against an organism needs to
be recorded in the context of these other \emph{metadata}. Consequently
the opportunity exists to create a repository to record the data, its
contextual data (metadata) and data classifiers in the form of
ontological or structured vocabulary terms. The structured nature of
this repository would support manual and autonomous data discovery as
well as provide the infrastructure for data based collaborations with
domestic and international research institutions. Currently there are
no such integrated systems available to the two facilities. The
National eResearch Architecture Taskforce (NeAT) Australia initiated
the PODD project to fill this gap. In PODD, we have engaged in the
design and development of the Phenomics Ontology Driven Data (PODD)
repository. The goals of PODD are to capture, manage, annotate and
distribute the data generated by mouse and plant phenomics research
activities. It supports both Australian and international biological
research communities by providing repository and data publication
services.

\section{The Architecture of the Ontology-driven Data Management System}\label{sec:design}

\subsection{Requirements of Data Management Systems}
For any scientific data management systems, a number of requirements
need to be satisfied.

\begin{description}
\item [Data storage and management]
Research activities in data-intensive disciplines such as
'omics often generate huge amounts of data. The ability to efficiently acquire, store and
manage large volumes of data is essential.

\item [Data contextualization]
Sufficient contextual information needs to be maintained for
more effective organization, understanding and discovery of raw data.
Contextual information includes both conceptual
domain models, such as how research activities are organized
and carried out; and metadata such as provenance information.

\item [Data security]
There are many dimensions to data security, including access control,
versioning and backup. An effective data management system needs to
ensure data security through the use of authentication and
authorization and sound versioning and backup solutions.

\item [Data identification and longevity] In order to support the
dissemination of scientific findings, data in the repository needs
to be publicly accessible after being published. Hence, a persistent
and unique naming scheme is required. Moreover, valuable scientific
data also need to be stored in perpetuity.

\item [Data reuse and integration]
Contextual information helps to make sense of raw data.
Moreover, it also needs to be made discoverable,
through means such as full-text search, faceted browsing
and complex query answering, to allow raw data to be integrated and reused.

\item [Model extensibility]
A data management system may need to manage a wide
variety of data, which may be generated by different software
and captured by different platforms. An expressive and extensible
domain model is therefore essential to cater for modification,
addition and deletion of domain concepts. The data management
system also needs to be designed to minimize service disruption
when such a model change occurs.

\end{description}

\subsection{The Ontology-driven Architecture For Data Management Systems}
The most distinguishing characteristic of the ontology-driven architecture
is the central role ontologies play. In this architecture,
raw data are not stored in a flat structure but are attached in
domain objects organized in a logical, hierarchical
system, defined according to the domain model that
represents the structure of research activities.

Current document management systems such as
Fez\footnote{\url{http://fez.library.uq.edu.au/}} typically have a
relatively static domain model and hardwire it as relational schemas
and foreign-key constraints in a custom relational database
independent from the underlying repository system. Consequently, the
information pertinent to the model of each concrete object is stored
in this custom database as well. As stated in the previous section,
this approach is unsuitable for dynamic environments where
conceptual changes are common.

To effectively support a dynamic conceptual framework, the domain
model in the proposed architecture is defined using OWL ontologies,
in which: OWL classes represent domain concepts; OWL properties
define concept attributes and their relationships; OWL restrictions
specify constraints on concepts and finally; OWL individuals define
concrete domain objects where attributes and relationships are
defined using OWL assertions. Raw data files are attached to
concrete domain objects.

Such a conceptual architecture alleviates the problem of imposing hard
relational constraints in a database which is difficult to extend/change.
It is worth noting that referential integrity is not sacrificed in achieving
flexibility: ontological reasoning involving relevant concepts and objects
are performed before object modification to ensure that all constraints
are satisfied.

Another drawback of existing systems is that there can be only one
domain model. When a concept needs to be updated, all the existing
objects need to be updated accordingly, which may be undesirable,
inappropriate and time-consuming. This is, unfortunately, unavoidable
as long as the domain model is defined using database schemas. In our
proposed architecture,
as concept and object definitions are stored in the repository,
such changes can be versioned so that existing instance objects
can remain legitimate when integrity validation is performed as they
can still refer to the previous conceptual definitions.

Similarly, the data and metadata (including ontological definitions)
of each object can be modified, and the modifications should be
versioned so that they can be rolled back.

The major differences between the ontology-driven architecture
and that of current systems can be summarized in Table~\ref{tab:diff} below.

\begin{table}[htdp]
\footnotesize
\centering
\caption{Summary of major conceptual differences between the ontology-driven
architecture and current systems.}\label{tab:diff}
\begin{center}
\begin{tabular}{p{50pt}|p{120pt}|p{120pt}}
\textbf{Aspect}  & \textbf{PODD} & \textbf{Current systems}\\
\hline
Domain modeling & OWL Ontologies & Database schemas\\
\hline
Model organization & In repository, using OWL classes, properties \& individuals & In custom databases, hardwired\\
\hline
Object organization & Ontological definition an integral part, stored in the repository & Stored in databases\\
\hline
Model change & Creates new version of OWL definitions, old objects unaffected & Database \& schema migration, old objects need to be updated\\
\hline
Object change & Creates new version of OWL definitions & Table updates\\
\hline
Referential integrity & OWL reasoning & Database integrity check
\end{tabular}
\end{center}
\end{table}

In developing the ontology-driven architecture,
the following design decisions have been made to balance expressivity,
flexibility and conceptual clarity.

\begin{itemize}
\item There is a top-level domain concept, called \emph{Project}
\footnote{The choice of concept names in the domain ontology is actually
irrelevant to the proposed architecture. Names such as \emph{Project}
and others are chosen as they are general and representative enough.},
under which other concepts (such as \emph{Investigation} and
\emph{Material}) reside in a hierarchical manner.

\item Access control (authorization) is defined on the \emph{Project}
level but not on an individual object level, i.e., a given user will have
the same access rights for all objects within a given project.

\item Within a \emph{Project} hierarchy, objects are in
a parent-child relationship in a tree structure such that
each child can only have one parent. This ensures that
access rights are properly propagated from parent to
child and there is no chance of confusion.

\item Additionally, inter-object, many-to-many reference
relationships can be defined to enhance flexibility of the
architecture as it allows arbitrary links between objects
to be established.

\item Objects cannot be shared across \emph{Projects}. Instead,
objects must be copied from one project and pasted into another one.
Such a rule simplifies object management with the elimination of
possible side-effects caused by sharing object between projects.

\item There should be no interference between different
versions of a given concept and between objects that are instances of
different concept versions.
\end{itemize}

\begin{figure}[htb]
\centering
\includegraphics[trim = 60mm 30mm 50mm 36mm, clip,height=75mm]{architecture.pdf}
\vspace{-16pt} \caption{A high-level depiction of components in the ontology-driven
architecture.}\label{fig:arch}
\end{figure}

The high-level design of ontology-driven architecture takes a modular
and layered approach, as can be seen in Figure~\ref{fig:arch}.
At the foundation is the \textbf{data access layer},
consisting of an underlying repository system, an RDF triple store, an
in-house database that stores essential information and a full-text search
engine. This layer is responsible for low-level tasks when the creation,
modification and deletion of concepts and objects occur.
The \textbf{business logic layer} in the middle is responsible for
managing concepts and objects, such as versioning, object conversion
and integrity validation. The \textbf{security layer} controls access
(authentication and authorization) to concepts and objects and guards
all operations on them. In this architecture, authorization is based
on user attributes, which have two dimensions. Firstly, each user has
a system-wide role, such as registered user or system administrator, which
is used to determine access rights across the system.
Secondly, a project-wide role, such as project administrator and
project observer, can be assigned to a user so that he can have
project-specific access rights. At the top of the stack is
the \textbf{interface layer}, where the data management
system can be accessed using a number of interfaces such as a Web
browser or API calls.

%In this section, we presented the PODD conceptual and system
%architecture that can effectively support concept dynamics
%through an ontology-driven design. In the next section, we present
%details of the PODD domain ontology in a bioinformatics setting.

\section{Ontology-based Domain Modeling}\label{sec:ont}
%One of the important design decisions to be made early in the
%development process is the domain model. Domain modeling aims at
%providing solutions for two important tasks. Firstly, efficient and
%flexible data organization; and secondly, data contextualization in
%the form of \emph{metadata}, so as to provide meaning for the raw
%data (documents, publications, image files, etc.) to facilitate
%search, query, dissemination, and so on.

As we emphasized previously, the domain model should be flexible
enough to accommodate the rapid changes and dynamic nature of
scientific research. In this section, we present the base
ontology and the roles it plays in the ontology-driven
architecture. It should be noted that the architecture
proposed here is domain-independent and it be applied
to any scientific discipline that
shares a similar high-level domain model.

%Section~\ref{sec:podd_ont} describes the PODD ontology in detail,
%introducing essential domain concepts and the how inter-relationships
%and attributes are modeled. Section~\ref{sec:rav}
%describes the important roles the domain ontology performs at the
%various stages of object lifecycle.

\subsection{The Base Domain Ontology}\label{sec:podd_ont}
Inspired by FuGe and OBI, we create the base domain ontology
in OWL to define essential domain concepts, their attributes
and inter-relationships in an object-oriented fashion.
As stated in the previous section,
domain concepts will be modeled as OWL classes; relationships
between concepts and object attributes will be modeled as OWL object and
datatype properties. Concrete objects will be modeled as OWL
individuals.

For an overview, attributes and inter-relationships of some
of the domain concepts and the core metadata of all concepts
in this ontology are shown in Figure~\ref{fig:model}.

\begin{figure}[htb]
\centering
\includegraphics[trim = 54mm 20mm 64mm 30mm, clip,height=92mm]{model.pdf}
\vspace{-16pt} \caption{The attributes and relationships of some
domain concepts.}\label{fig:model}
\end{figure}

First of all, we set out a few design principles of
the domain ontology.
\begin{itemize}
\item All essential domain concepts are modeled as subclasses of
an abstract top-level OWL class \textbf{\emph{PODDConcept}} that
captures common attributes and relationships.

\item All relationships between domain concepts are captured by
\emph{domain properties}, which can be further divided into two
\emph{property hierarchies}, one for parent-child relationships
and the other for reference relationships. Each of the two
hierarchies have an abstract top-level property, called
\textbf{\emph{contains}} and \textbf{\emph{refersTo}}, respectively.

\item All parent-child relationships are
modeled in a property hierarchy as subproperties of the abstract
property \textbf{\emph{contains}}, and all reference relationships
are modeled in another property hierarchy as subproperties of the
abstract property \textbf{\emph{refersTo}}.

\item For each domain concept $T$, one property is defined in each
of the above hierarchies with its range defined to be $T$.
The domains of such properties are not specified so that
they can be used by any applicable domain concept to establish
a relationship between them.

\item Domain concepts (except \textbf{\emph{PODDConcept}})
are mutually disjoint with each other.

\item Essential domain concepts can be subclassed to provide more
specialized and refined information.
\end{itemize}

Moreover, to ensure that each object can have at most one parent object,
the inverse property of \textbf{\emph{contains}},
\textbf{\emph{isContainedBy}}, is defined so that a max cardinality
restriction can be added to the top-level concept \textbf{\emph{PODDConcept}}
to enforce it.

The definitions of the high-level constructs can be summarized
in Figure~\ref{fig:top}, in OWL DL syntax~\cite{hoph03a}.

\vspace{-8pt}
\begin{figure}[htb]
\small\centering
\begin{align*}
&PODDConcept \sqsubseteq \top & isContainedBy = (^-contains)\\
&\top \sqsubseteq \forall contains. PODDConcept
&\top \sqsubseteq \forall refersTo. PODDConcept\\
&PODDConcept \sqsubseteq \leq 1~ isContainedBy
\end{align*}

\vspace{-8pt}
\caption{Top-level ontology constructs in the PODD ontology.}\label{fig:top}
\end{figure}

\subsubsection{The Domain Concepts}
In the base domain ontology, essential domain classes include the following:

\begin{description}
\item[\emph{Project}] is the top-level
concept (under \textbf{\emph{PODDConcept}}) representing
an administrative concept that contains essential meta
information about the research project, such as the
administering organization, principal
investigator, project membership, project status, etc.

\item[\emph{ProjectPlan}] describing the research
plan of the project at the core metadata element level.

\item[\emph{Platform}] describing any single technical
measurement platform used in the project. Technical measurement
platform means any platform for which parameters and parameter
values may be captured.

%\item[\emph{Genotype}] describing the genotype of the
%materials used in the investigations. Multiple genotypes are
%described here and then can become fields in the instances of the
%Material object.

\item[\emph{Investigation}] describing a planned process
within a project that executes some form of study design and
produces a coherent set of results. It can be considered equivalent
to an experiment.

The \textbf{\emph{Investigation}} concept is of central importance.
It captures the data and metadata of experiments under a project. A
number of concepts are defined to assist in the modeling of
investigations.

\item[\emph{Design}] describing experimental design
components, e.g. plant layouts, sampling strategies, etc.

%\item[\emph{Growth Condition}] describing growth conditions, such as
%growth chambers used, environmental settings, etc.

\item[\emph{Process}] representing a planned component of an
investigation. It is a description of a series of steps taken to
achieve the objective of the investigation.

\item[\emph{Protocol}] describing a step within a process that is a
consistent whole. e.g. sterilize seeds, plant seedlings, image the
plants.

\item[\emph{Material}] describing the materials used in the
investigation. Materials can be either inputs or outputs. They can
be chemicals, substrates, whole organisms or samples taken from a
whole organism. The meaning of the \emph{Material} is usually
dependent on the domain and hence can be specified by domain-specific
properties.

\item[\emph{Event}] capturing ad-hoc events and actions that occur
against an individual material. In most instances the events and
their timing are described in the process/protocol. An \emph{Event} object
can be utilized to either record fixed events in a form that allows
for investigative analysis, or to record one off observations (e.g., the
plant under observation died).

\item[\emph{Measurement}] describing a single measurement against a
single material. e.g. an image of a plant is a single measurement.
\emph{Measurement} objects can capture measurement variables (e.g. shutter
speed, lighting, etc).

\item[\emph{Analysis}] is a metadata concepts used to describe
outputs derived from measurements. An analysis can be performed
against a single material object, a set of investigations,
or at the level of the whole project.
\end{description}

\subsubsection{Inter-concept Relationships}
The structures and workflows of phenomics research activities are
captured using OWL object properties.

%Different research projects will utilize different measurement
%platforms and have significantly different approaches and project
%designs. As a result, different projects may have different
%structures. In OWL, there are a number of ways of defining the same
%relationship. In order to achieve high modeling flexibility and
%accommodate as many scenarios as possible, we have made the
%following design decisions:

Following the design principles presented in the beginning of this
section, domain properties such as \textbf{\emph{hasAnalysis}}
and \textbf{\emph{refersToPlatform}} can be defined as follows
in Figure~\ref{fig:props}.

\vspace{-8pt}
\begin{figure}[htb]
\small\centering
\begin{align*}
&hasAnalysis\sqsubseteq contains & refersToPlatform \sqsubseteq refersTo\\
&\top\sqsubseteq \forall hasAnalysis.Analysis & \top \sqsubseteq \forall refersToPlatform.Platform
\end{align*}

\vspace{-8pt}
\caption{Example definitions of domain properties}\label{fig:props}
\end{figure}

%For each of the concepts described in the previous subsection except
%for \emph{Project}, we define an OWL object-predicate with the range
%being the concept. For example, for concept \emph{Analysis}, we
%define a predicate \emph{hasAnalysis} and define its range to be
%\emph{Analysis}.

\subsubsection{Object Attributes}
Attributes are intrinsic properties of an object, such as the status
of a project, the start date of an investigation and the timestamp
of an event. In our model, we use OWL properties to model object
attributes, similar to the modeling of inter-object relationships.

When the possible values of a particular attribute can be enumerated,
such as project status (\emph{active}, \emph{inactive} and \emph{completed}),
an enumerated OWL class is used to represent all the values.
When an attribute represents a grouping of some values, such
as accessions, where an accession has a source and a number,
an OWL class is also defined to represent the grouping.
In this case, auxiliary OWL properties are defined to project
out specific values in the grouping. In all other cases, attributes
are modeled using datatypes.

Figure~\ref{fig:project_owl} shows the partial definition of the OWL
class $Project$. Restriction
(\ref{for:hasPlan}), for example, states that any $Project$ instance must have
exactly one $ProjectPlan$ (through the predicate $hasProjectPlan$,
the range of which is $ProjectPlan$). The other 3 restrictions are
similarly defined.

\vspace{-8pt}
\begin{figure}[htb]
\centering
\small
\begin{align}
Project &\sqsubseteq~ =~ 1~ hasProjectPlan\label{for:hasPlan}\\
        &\sqsubseteq~ \geq~ 1~ hasInvestigation\\
        &\sqsubseteq~ =~ 1~ hasStartDate\\
        &\sqsubseteq~ \leq~ 1~ hasPublicationDate
\end{align}
\vspace{-16pt} \caption{Partial OWL Definition for the Project
concept.}\label{fig:project_owl}
\end{figure}

\subsection{Roles of Domain Ontologies in Object Life Cycle}\label{sec:rav}
The base ontology defines essential concepts independent of the domain.
Domain-specific information can be then captured by extending the
base ontology in individual systems.

Concrete objects which are instantiations of various concepts such as
\textbf{\emph{Project}} and \textbf{\emph{Investigation}}, are stored
in the repository and can
subsequently be retrieved for different purposes. As stated in
Section~\ref{sec:intro}, the ontology-based domain model is at the
center of the whole life cycle of objects. In this subsection, we
briefly describe the roles the domain ontologies perform at
various stages of the object life cycle.

\begin{description}
\item[Ingestion] When an object is created, the user specifies which
type of object she intends to create and the repository will pull up
all the ontological definitions for that type (from the OWL class
corresponding to that type and its super classes). Such definitions
will be used to (a) guide the rendering of object creation
interfaces and (b) validate the attributes and inter-object
relationships the user has entered before the object is ingested.
When the object is ingested, its metadata definitions are stored
as RDF assertions.

\item[Retrieval \& update] When an object is retrieved from the repository,
its attributes and inter-object relations are retrieved from its
RDF assertions, which are used to drive the on-screen rendering.
When any value is updated, it is validated and updated in this
object's RDF assertions.

\item[Query \& search] An object's assertions will be stored
in an RDF~\cite{rdfprimer04} triple store, which can be queried
using the SPARQL~\cite{sparql} query language. Similarly, ontology
definitions are indexed to provide functionalities such as full-text
search and faceted browsing.

\item[Publication \& export] When an object is published or exported,
its metadata, in RDF, will be retrieved and exported.
\end{description}

\subsection{Integration with Existing Domain Ontologies}
As stated before, ontologies such as Gene
Ontology~\cite{citeulike:212874} and Plant
Ontology~\cite{citeulike:3008167} are widely used in biomedical
research to capture information such as genes, proteins, sequences
and organism phenotypes. In our ontology-driven approach, these
ontologies will be used to add annotations on domain objects to
enrich their semantic descriptions and enable cross-application
reference and integration.

In summary, ontology-based domain modeling enables us to build very
expressive and extensible conceptual models which can be extended
to accommodate individual domains. Ample tool
support is also available to perform ontology-based tasks such as
validation, query answering and searching.
%In the next section,
%we describe our effort in building the PODD data management system
%based on the ontology-driven architecture and discuss in more detail
%the roles the ontology plays.

\section{The PODD Data Management System}\label{sec:podd}

Based on the ontology-driven architecture presented in Section~\ref{sec:design}
and the domain ontology presented in Section~\ref{sec:ont},
we have developed the PODD data management system to meet the
data management challenges faced by the Australian phenomics
research community.

To describe domain knowledge in phenomics, we extend the base
ontology described in Section~\ref{sec:ont} by defining the
following concepts: \emph{Genotype}, \emph{Gene}, \emph{Phenotype}
and \emph{Sequence} as subclasses of \emph{PODDConcept}. Additional
OWL object and datatype properties are also defined to model the
attributes and relationships of these concepts, such as shown in
Figure~\ref{fig:phe_ont}. Note that the last two definitions
integrate the new definitions with those in the base ontology.

\vspace{-8pt}
\begin{figure}[htb]
\small
\centering
\begin{minipage}[t]{.55\linewidth}
\centering
\begin{align*}
Genotype &\sqsubseteq~ PODDConcept\label{for:hasPlan}\\
         &\sqsubseteq~ \forall hasGene.Gene\\
         &\sqsubseteq~ \leq~ 1~ hasEcotype\\
         &\sqsubseteq~ \leq~ 1~ hasSubspecies\\
         &\cdots\\
Phenotype & \sqsubseteq PODDConcept\\
          & \sqsubseteq \forall hasMeasurement.Measurement\\
         &\cdots
\end{align*}
\end{minipage}
\begin{minipage}[t]{.4\linewidth}
\centering
\begin{align*}
Gene & \sqsubseteq~ PODDConcept\\
     & \sqsubseteq~ \forall hasSequence.Sequence\\
     & \sqsubseteq~ \leq~ 1~ hasAlias\\
     & \sqsubseteq~ \leq~ 1~ hasChromosome\\
     & \cdots
\end{align*}
\end{minipage}
\begin{align*}
Project & \sqsubseteq \forall hasGenotype.Genotype\\
Material & \sqsubseteq \forall hasPhenotype.Phenotype\\
         & \sqsubseteq \forall refersToGenotype.Genotype
\end{align*}

\vspace{-8pt}
\caption{Domain-specific OWL defintions.}\label{fig:phe_ont}
\end{figure}

The above definitions show how the new OWL classes are
integrated with existing ones through the use of OWL
object properties. Conceptually, it states that phenotype
objects are attached to material objects, which in turn
can reference genotype objects, which are defined as child
objects of projects.

As can be seen above, the domain model is dynamic in that
it accommodates the addition of new classes and properties.
As a result, the other components of the PODD system is also
designed to be amenable to the dynamic nature of the architecture.
%The Web interface is dynamically generated by reading the domain
%ontologies and translating the various ontological constructs
%(OWL restrictions, for example)
%into familiar Web elements. Such a principled approach
%makes the user interface layer both extensible and easy to use.
%
The PODD system is developed in Java for its cross-platform
availability and the huge ecosystem around it. A more detailed
architecture diagram for PODD can be seen in
Figure~\ref{fig:podd_arch}. Below we summarize a number of core
technologies employed in the development of the PODD data management
system and describe briefly how we use them to achieve extensibility
and scability.

\begin{figure}[htb]
\centering
\includegraphics[trim = 60mm 30mm 50mm 28mm, clip,height=75mm]{podd_arch.pdf}
\vspace{-8pt} \caption{The architecture and components of
PODD.}\label{fig:podd_arch}
\end{figure}

\begin{itemize}
\item We use Fedora Commons, a digital repository for the
storage and retrieval of domain objects.

As introduced in Section~\ref{sec:overview}, Fedora Commons is a mature,
widely-used digital repository software. In PODD, we store domain concepts as
Fedora models and concrete domain objects as Fedora objects. The OWL (for
concepts) and RDF (for objects) definition of each concept and object
is stored in a special versioned datastream \texttt{PODD}, which is used
by the system in various tasks such as object creation, rendering, validation
and update. Moreover, raw data files are also stored in Fedora Commons
as versioned datastreams.

It is to be noted that the structure of domain objects is not maintained
in Fedora Commons but rather in PODD system on top of it. In other words,
Fedora Commons is not aware of the semantics of concepts and objects.

\item We use iRODS~\cite{irods07}, a distributed, grid-based storage
software system, as the storage module for Fedora Commons to provide
a distributed storage solution. Transparent to Fedora Commons,
iRODS manages the distribution and replication across a virtual,
geographically-distributed data fabric.

iRODS provides two important
benefits: (1) data transfer can enjoy principle of locality by
replicating files at strategically placed data storage nodes in the fabric;
and (2) storage size is theoretically unlimited as the data fabric can
grow as demand grows.

\item We incorporate the Sesame\footnote{\url{http://www.openrdf.org/}}
triple store to support complex query answering with SPARQL. It also
provides metadata to the full-text index and search component.

For each object, its RDF definitions are stored in a
Sesame \emph{context} that uniquely scopes and identifies
the triples. The use of contexts is necessitated by
two facts. Firstly, it cannot be guaranteed that triples are
unique across all objects. For example, \emph{Accession}
objects are defined as RDF blank nodes with an accession
source and an accession number. Hence, it is possible
that two objects contain the \emph{same} accession object,
which cannot be distinguished in the triple store. Hence,
contexts are necessary to identify all triples of a given
object. Secondly, as described in Section~\ref{sec:design},
access control needs to be enforced on a project level.
Similarly, it also needs to be enforced on query answering
in the triple store. By identifying triples of individual
objects, we are able to control contexts a user can access
through query expansion.

\item We use the Solr\footnote{\url{http://lucene.apache.org/solr/}}
open-source search engine platform to provide full-text search
and faceted browsing capabilities of repository
contents, including values in the RDF triple store.

Similar to the structure of the Sesame triple store,
there is a one-to-one correspondence between domain objects
in the repository and the Solr \emph{documents}, which
represent logical indexing units. Moreover, through
search query expansion, access control can be applied to
search results.

\item We use MySQL database to store user and access control
related information, as it is orthogonal to other domain
concepts.

\end{itemize}

Figure~\ref{fig:browser} shows the browser view of a plant phenomics
project that investigates Arabidopsis. Note that in this view, the
objects are shown in a tree-like structure by following the property
assertions of subproperties of \emph{contains} defined in the base
and domain ontologies.

\begin{figure}[htb]
\centering
\includegraphics[trim = 0mm 2mm 5mm 0mm, clip,width=120mm]{browser.png}
\vspace{-4pt} \caption{The browser view of a plant phenomics project
in PODD.}\label{fig:browser}
\end{figure}

Following the ontology-driven architecture and through the use
of mature technologies, we have successfully developed PODD, a
scalable, extensible data management system. At the same time,
data management tasks such as versioning, logical organization
of data, authentication \& authorization, and discovery, can
all be supported effectively. In summary, the PODD system
demonstrates the feasibility and practicality of the proposed
ontology-driven architecture.

\section{Conclusion}\label{sec:conclusion}
Sound data management practice is a challenge faced by
many data-intensive scientific disciplines. A number of
root causes contribute to this challenge.
Firstly, scientific advancement
usually connotes that the conceptual data model is being
continually evolved, requiring the data management system
to be sufficiently adaptive to accommodate future changes.
Secondly, huge amounts of data are being generated off a
wide variety of instruments and software, requiring the data
management system to be highly scalable for efficient data
processing. Moreover, an important requirement of data management
systems is to organize data in a logical way to facilitate
tasks such as curation, integration, discovery and dissemination.
Hence, the management of metadata is of central importance.

Traditional data management systems are typically developed
around a relational database in which database schemas
define the domain model and constraints on
abstract concepts. A change in schema design normally
requires database migration, which is an error-prone
process. As a result, such systems are not best
suited in a dynamic environment where model evolution is
the norm rather than the exception.

In recent years a number of repository systems such as
Fedora Commons and Apache Jackrabbit have been developed.
These systems impose minimal modeling constraints and hence
are more flexible. As a consequence, however, rich logical
structures of data cannot be easily expressed and maintained
in such systems.

Given their intrinsic characteristics of semantic rigor and
open nature, we believe ontology languages such as OWL and
RDFS are an ideal conceptual foundation on which effective
data management systems can be built. In this paper, we
propose an ontology-driven architecture for developing data
management systems that are able to handle dynamic data and models.

In our architecture, an ontology defines the behaviors of and
relationships between domain objects using OWL vocabularies. Such
definitions play a central role in all data management tasks
including data acquisition, validation, presentation, discovery and
integration.

We present the base ontology which encodes essential domain
knowledge to describe the structure of the data through the use of
OWL restrictions. The base ontology is designed in a way to strike a
balance between richness in modeling capabilities and the ease of
realization of data management requirements such as flexible
authentication and authorization. Moreover, the base ontology can be
naturally specialized to provide domain-specific definitions to
cater for the different needs of individual disciplines.

Phenomics is an emergent research discipline that is poised to have
a significant impact upon industrial-scale biological research.
Phenomics presents a number of data management challenges, such as
the management of evolving technologies and large volumes of data
and the integration of highly heterogeneous datasets. Consequently
phenomics represents an ideal domain for illustrating our
ontology-driven approach to research data management.

To validate the feasibility of the ontology-driven architecture
and to meet the data management needs of the Australian phenomics
research community, we developed the PODD repository to enable
efficient storage, retrieval, contextualization, query, discovery
and publication of large amounts of data.

Through the employment of the ontology-driven architecture,
the domain ontology and underlying repository system (Fedora
Commons), the PODD repository is highly adaptive that
the addition of new concepts and the modification
of existing concepts do not affect data already present
in the system. It is also able to perform effective data
management tasks over a large and growing amount of data.

In summary, our contribution in this work is three-fold: firstly,
the proposal of the ontology-driven architecture for developing data
management systems; secondly, the development of a base ontology
that defines essential domain knowledge; and thirdly, the
development of the PODD data management system in the validation of
the practicality of the proposed approach. To the best of our
knowledge, this is the first proposal of an ontology-driven
architecture for scientific data management systems.

We have identified a number of future work directions that
we would like to pursue. Firstly, we will investigate into the
integration with existing domain ontologies such as the
Gene Ontology~\cite{citeulike:212874} and the Plant
Ontology~\cite{citeulike:3008167}. One possibility
would be to use terms defined in these ontologies to
annotate metadata objects~\cite{1419711}. Secondly, we would like to
investigate the generalization of the ontology-driven
approach so that it can be applied to other areas such
as workflow management systems.
Thirdly, we will continue the development of the PODD system
to provide additional functionalities such as
data visualization, automated data integration and
Linked Data-style data discovery and publication.

\section*{Acknowledgement}
The authors wish to acknowledge the support of the National
eResearch Architecture Taskforce (NeAT) and the Integrated
Biological Sciences Steering Committee (IBSSC) Australia. The authors wish to
thank Dr Xavier Sirault, Dr. Kai Xu and Mr. Philip Wu for
the discussion on the development of the domain ontology
and the PODD system.

\bibliographystyle{elsarticle-num}
\newcommand{\gobble}[1]{}
\begin{thebibliography}{10}
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL
}\fi \expandafter\ifx\csname href\endcsname\relax
  \def\href#1#2{#2} \def\path#1{#1}\fi

\bibitem{1107503}
J.~Gray, D.~T. Liu, M.~Nieto-Santisteban, A.~Szalay, D.~J. DeWitt,
G.~Heber,
  Scientific data management in the coming decade, SIGMOD Rec. 34~(4) (2005)
  34--41.
\newblock \href {http://dx.doi.org/http://doi.acm.org/10.1145/1107499.1107503}
  {\path{doi:http://doi.acm.org/10.1145/1107499.1107503}}.

\bibitem{brm2007}
A.~Shah, M.~Singhal, K.~Klicker, E.~Stephan, H.~Wiley, K.~Waters,
Enabling
  high-throughput data management for systems biology: The bioinformatics
  resource manager, Bioinformatics 23~(7) (2007) 906--909.

\bibitem{rdfschema04}
{D. Brickley and R.V. Guha (editors)}, {Resource Description
Framework (RDF)
  Schema Specification 1.0}, \url {http://www.w3.org/TR/rdf-schema/} (Feb.
  2004).

\bibitem{hoph03a}
I.~Horrocks, P.~F. Patel-Schneider, F.~van Harmelen, {From
$\mathcal{SHIQ}$ and
  {RDF} to {OWL}: The Making of a Web Ontology Language}, Journal of Web
  Semantics 1~(1) (2003) 7--26.

\bibitem{linkeddata}
T.~Berners-Lee, {Linked Data},
  \url{http://www.w3.org/DesignIssues/LinkedData.html} (2007).

\bibitem{aue07dbpedia}
S.~Auer, C.~Bizer, G.~Kobilarov, J.~Lehmann, R.~Cyganiak, Z.~Ives,
  \href{http://dx.doi.org/10.1007/978-3-540-76298-0_52}{{DBpedia: A Nucleus for
  a Web of Open Data}}, in: Proceedings of 6th International Semantic Web
  Conference, 2nd Asian Semantic Web Conference (ISWC+ASWC 2007), 2008, pp.
  722--735.
\newline\urlprefix\url{http://dx.doi.org/10.1007/978-3-540-76298-0_52}

\bibitem{journals/bib/RuttenbergRSM09}
A.~Ruttenberg, J.~Rees, M.~Samwald, M.~S. Marshall, {Life Sciences
on the
  Semantic Web: the Neurocommons and Beyond}, Briefings in Bioinformatics
  10~(2) (2009) 193--204.
\newblock \href {http://dx.doi.org/10.1093/bib/bbp004}
  {\path{doi:10.1093/bib/bbp004}}.

\bibitem{citeulike:1882392}
B.~Smith, M.~Ashburner, C.~Rosse, et~al., {The OBO Foundry:
Coordinated
  Evolution of Ontologies to Support Biomedical Data Integration}, Nature
  Biotechnology 25~(11) (2007) 1251--1255.
\newblock \href {http://dx.doi.org/10.1038/nbt1346}
  {\path{doi:10.1038/nbt1346}}.

\bibitem{citeulike:212874}
M.~Ashburner, C.~A. Ball, J.~A. Blake, et~al.,
  \href{http://dx.doi.org/10.1038/75556}{{Gene Ontology: Tool for the
  Unification of Biology}}, Nat Genet 25~(1) (2000) 25--29.
\newblock \href {http://dx.doi.org/10.1038/75556} {\path{doi:10.1038/75556}}.
\newline\urlprefix\url{http://dx.doi.org/10.1038/75556}

\bibitem{citeulike:5008761}
C.~Bizer, T.~Heath, T.~Berners-Lee, Linked data - the story so far,
  International Journal on Semantic Web and Information Systems (IJSWIS) 5~(3)
  (2009) 1--22.

\bibitem{podd_icadl}
Y.-F. Li, G.~Kennedy, F.~Davies, J.~Hunter, {PODD: An
Ontology-driven Data
  Repository for Collaborative Phenomics Research}, in: Proceedings of 12th
  International Conference on Asian Digital Libraries (ICADL 2010),
  Springer-Verlag, 2010, to appear.

\bibitem{dhoule09}
D.~Houle, {Numbering the Hairs on Our Heads: The Shared Challenge
and Promise
  of Phenomics}, {Proceedings of the National Academy of Sciences} 107 (2009)
  1793--1799.

\bibitem{Sauer200458}
U.~Sauer, {High-throughput Phenomics: Experimental Methods for
Mapping
  Fluxomes}, Current Opinion in Biotechnology 15~(1) (2004) 58--63.
\newblock \href
  {http://dx.doi.org/http://dx.doi.org/10.1016/j.copbio.2003.11.001}
  {\path{doi:http://dx.doi.org/10.1016/j.copbio.2003.11.001}}.

\bibitem{nevo01}
E.~Nevo, Evolution of genome-phenome diversity under environmental
stress,
  {Proceedings of the National Academy of Sciences} 98~(11) (2001) 6233--6240.

\bibitem{plant_furbank}
R.~T. Furbank, Plant phenomics: from gene to form and function,
Functional
  Plant Biology 36~(10) (2009) 5--6.

\bibitem{DBLP:journals/nar/WheelerBBBCCCDEFGHKKKLMMOPSSSSSSSSTTWY06}
D.~L. Wheeler, et~al., Database resources of the national center for
  biotechnology information, Nucleic Acids Research 34~(Database-Issue) (2006)
  173--180.

\bibitem{AmosBairoch01012005}
A.~Bairoch, et~al., {The Universal Protein Resource (UniProt)},
Nucl. Acids
  Res. 33~(suppl\_1) (2005) D154--159.
\newblock \href {http://dx.doi.org/10.1093/nar/gki070}
  {\path{doi:10.1093/nar/gki070}}.

\bibitem{783165}
C.~Baru, R.~Moore, A.~Rajasekar, M.~Wan, {The SDSC storage resource
broker},
  in: CASCON '98: Proceedings of the 1998 conference of the Centre for Advanced
  Studies on Collaborative research, IBM Press, 1998, p.~5.

\bibitem{652836}
W.~Hoschek, F.~J. Ja\'{e}n-Mart\'{\i}nez, A.~Samar, H.~Stockinger,
  K.~Stockinger, Data management in an international data grid project, in:
  GRID '00: Proceedings of the First IEEE/ACM International Workshop on Grid
  Computing, Springer-Verlag, London, UK, 2000, pp. 77--90.

\bibitem{565296}
K.~Krauter, R.~Buyya, M.~Maheswaran, A taxonomy and survey of grid
resource
  management systems for distributed computing, Softw. Pract. Exper. 32~(2)
  (2002) 135--164.
\newblock \href {http://dx.doi.org/http://dx.doi.org/10.1002/spe.432}
  {\path{doi:http://dx.doi.org/10.1002/spe.432}}.

\bibitem{Spallation_metadatamodel}
S.~Sufi, B.~Mathews,
\href{http://epubs.cclrc.ac.uk/bitstream/485/}{{CCLRC
  Scientific Metadata Model: Version 2}} (2004).
\newline\urlprefix\url{http://epubs.cclrc.ac.uk/bitstream/485/}

\bibitem{citeulike:3735746}
L.~N. Soldatova, R.~D. King,
\href{http://dx.doi.org/10.1098/rsif.2006.0134}{An
  ontology of scientific experiments.}, Journal of the Royal Society, Interface
  3~(11) (2006) 795--803.
\newblock \href {http://dx.doi.org/10.1098/rsif.2006.0134}
  {\path{doi:10.1098/rsif.2006.0134}}.
\newline\urlprefix\url{http://dx.doi.org/10.1098/rsif.2006.0134}

\bibitem{citeulike:151946}
A.~Brazma, et~al.,
\href{http://dx.doi.org/10.1038/ng1201-365}{{Minimum
  information about a microarray experiment (MIAME) -- toward standards for
  microarray data}}, Nat Genet 29~(4) (2001) 365--371.
\newblock \href {http://dx.doi.org/10.1038/ng1201-365}
  {\path{doi:10.1038/ng1201-365}}.
\newline\urlprefix\url{http://dx.doi.org/10.1038/ng1201-365}

\bibitem{citeulike:1756058}
A.~R. Jones, M.~Miller, R.~Aebersold, et~al., {The Functional
Genomics
  Experiment model (FuGE): an Extensible Framework for Standards in Functional
  Genomics}, Nature Biotechnology 25~(10) (2007) 1127--1133.
\newblock \href {http://dx.doi.org/10.1038/nbt1347}
  {\path{doi:10.1038/nbt1347}}.

\bibitem{rdfprimer04}
F.~Manola, E.~M. (editors), {{RDF Primer}},
  \url{http://www.w3.org/TR/rdf-primer/} (Feb. 2004).

\bibitem{sparql}
E.~Prud'hommeaux, A.~Seaborne, {SPARQL Query Language for RDF},
  \url{http://www.w3.org/TR/2006/CR-rdf-sparql-query-20060406/} (Apr. 2006).

\bibitem{irods07}
A.~Rajasekar, R.~Moore, F.~Vernon, {iRODS: A Distributed Data
Management
  Cyberinfrastructure for Observatories}, in: American Geophysical Union, Fall
  Meeting 2007, 2007.

\bibitem{citeulike:3008167}
S.~Avraham, C.-W. Tung, K.~Ilic, P.~Jaiswal, E.~A. Kellogg,
S.~Mccouch,
  A.~Pujar, L.~Reiser, S.~Y. Rhee, M.~M. Sachs, M.~Schaeffer, L.~Stein,
  P.~Stevens, L.~Vincent, F.~Zapata, D.~Ware,
  \href{http://dx.doi.org/10.1093/nar/gkm908}{The plant ontology database: a
  community resource for plant structure and developmental stages controlled
  vocabulary and annotations}, Nucl. Acids Res. 36~(suppl\_1) (2008) D449--454.
\newblock \href {http://dx.doi.org/10.1093/nar/gkm908}
  {\path{doi:10.1093/nar/gkm908}}.
\newline\urlprefix\url{http://dx.doi.org/10.1093/nar/gkm908}

\bibitem{1419711}
R.~Schroeter, J.~Hunter, A.~Newman, Annotating relationships between
multiple
  mixed-media digital objects by extending annotea, in: ESWC '07: Proceedings
  of the 4th European conference on The Semantic Web, Springer-Verlag, Berlin,
  Heidelberg, 2007, pp. 533--548.
\newblock \href
  {http://dx.doi.org/http://dx.doi.org/10.1007/978-3-540-72667-8\_38}
  {\path{doi:http://dx.doi.org/10.1007/978-3-540-72667-8\_38}}.

\end{thebibliography}

\end{document}

### 2012-03-02 seres2010/podd.tex
initial files from http://projects.arcs.org.au/svn/podd/trunk/doc/publications/

\documentclass{llncs}

\usepackage{url}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{zed-cm}
\usepackage{booktabs}
\usepackage{color}
\usepackage{alltt}
\usepackage{boxedminipage}
\usepackage{graphicx}

\title{Towards A Semantic \& Domain-agnostic Scientific Data Management System}

\author{Yuan-Fang Li, Gavin Kennedy, Faith Davies, Jane Hunter}

\institute{
School of ITEE\\
The University Of Queensland\\
Brisbane, Queensland 4072, Australia\\
\email{\texttt{\{}uqyli4,g.kennedy1,f.davies,j.hunter\texttt{\}}@uq.edu.au}
}

\begin{document}
\maketitle

\begin{abstract}
Data management has become a critical challenge faced by a wide array
of scientific disciplines in which the provision of sound data
management is pivotal to the achievements and impact of research
projects. Massive and rapidly expanding amounts of experimental data combined with
evolving domain models contribute to making data management
an increasingly challenging task that warrants a rethinking of its
design. In this paper we present PODD, an ontology-centric data management system architecture for scientific experimental data that is extensible and domain independent.
In this architecture, the behaviors of domain concepts and objects are
specified entirely by ontological entities, around which all data
management tasks are carried out. The open and semantic nature of ontology
languages also makes PODD amenable to greater data reuse and
interoperability. To evaluate this architecture, we have developed a data 
management system and applied it to the challenge of managing phenomics data.
\end{abstract}

\section{Introduction}\label{sec:intro}
Data management is the practice of managing (digital) data and
resources, encompassing a wide range of activities including
acquisition, storage, retrieval, discovery, access control,
publication and archival. For many
data-intensive scientific disciplines such as life sciences and
bioinformatics, sound data management informs and enables research and
has become an indispensable component~\cite{1107503}.

The need for effective data management is, in a large part, due to the
fact that massive amounts of digital data are being generated by modern
instruments. Furthermore, the fast evolution of technologies/processes
and discovery of new scientific knowledge require flexibility in
handling dynamic data and models in data management systems. Among
others, there are three core challenges for effective data management
in scientific research.

\begin{itemize}
\item The ability to provide a data management service that can manage
large quantities of heterogeneous data in multiple formats (text, image,
and video) and not be constrained to a finite set of experimental,
imaging and measurement platforms or data formats.

\item The ability to support metadata-related services to provide context
and structure for data within the data management service to facilitate
effective search, query and dissemination.

\item The ability to accommodate evolving and emerging knowledge,
technologies and processes.
\end{itemize}

Database systems have traditionally been used successfully to manage
research data~\cite{brm2007} in which database schemas are used as
domain models to capture attributes and relationships of domain
concepts. One implication of the above approach is that domain models
need to stay relatively stable as database extension and migration is
often an error-prone and laborious task. Consequently, this approach is
not suitable for domains where data and model evolution is the norm
rather than the exception.

Ontology language OWL possesses expressive, rigorously-defined
semantics and non-ambiguous syntaxes. it has been designed to be open
and extensible and to support knowledge and data exchange on the
Web~\cite{aue07dbpedia,journals/bib/RuttenbergRSM09,citeulike:1882392}.
These intrinsic characteristics make them an ideal conceptual platform
on which a flexible scientific data management system can be built.

In this paper, we present our work in designing PODD (Phenomics Ontology 
Driven Data Management), a semantic,
domain-agnostic architecture for systems managing data generated
from scientific experiments that employes ontologies as domain models. The
ontology-based domain model is at the core of PODD as it defines the
behavior of entities in scientific experiments.
Logical structure of data is therefore maintained and enforced via
ontological definitions and reasoning, and not via database schemas and
associated constraints.

In~\cite{podd_icadl} we described an early version of the PODD data
repository to meet the above challenges facing the Australian phenomics
research community. We would like to emphasize that although the PODD system presented
in~\cite{podd_icadl} is geared towards phenomics research, the
ontology-centric architecture we propose in this paper is actually
domain-independent and can be applied in any scientific discipline
where research activities and output can be conceptually organized in a structured
manner.

The rest of the paper is organized as follows. In
Section~\ref{sec:overview} we present related work and give a brief
overview of the motivation and goals of the PODD project.
Section~\ref{sec:arch} presents the ontology-based architecture for
data management systems. In Section~\ref{sec:ont}, we discuss the PODD
ontologies in more detail and show how the ontology-based modeling
approach is used in the life cycle of repository concepts and objects.
In Section~\ref{sec:podd}, we describe the PODD data management
system we developed based on the ontology-driven architecture. Finally,
Section~\ref{sec:concl} concludes the paper and identifies future
directions.

\section{Overview}\label{sec:overview}
%Over the years attempts have been made to develop semantical content
%repository systems and architectures to meet institutional and personal
%data management needs.
In this section, we survey a number of related systems and
architectures. Following the survey, we present the motivation behind
the ontology-centric architecture and the goals we wish to achieve with
the PODD data management system.

\subsection{Related Work}
A number of ontology repositories and search engines have been
developed. Repositories such as NCBO
Bioportal\footnote{\url{http://bioportal.bioontology.org/}} and
Cupboard\footnote{\url{http://kmi-web06.open.ac.uk:8081/cupboard}}
publish ontologies and usually support functionalities including
full-text \& faceted search, hierarchical browsing, visualization and
cross references. Ontology search engines such as
Swoogle\footnote{\url{http://swoogle.umbc.edu/}} and
Watson\footnote{\url{http://watson.kmi.open.ac.uk/}} index and store
large numbers of ontologies and make them searchable.

There are also prior works in developing content repository systems.
Fedora Commons\footnote{\url{http://www.fedora-commons.org/}} 
is a widely used open-source, general-purpose digital resource 
management system based on the principles
of modularity, interoperability and extensibility. In Fedora Commons,
abstract concepts are defined as models, on which inter-relationships
and behaviors can be further defined. Data in Fedora Commons
repositories are organized into objects, which have datastreams that
stores either metadata or data. PhenomicDB~\cite{citeulike:1036090} 
is a multi-organism phenotype-genotype 
database for a number of model organisms. It contains data from a number of 
primary databases including FlyBase, Phenobank, OMIM and NCBI Gene.
More recently, an ontology-based approach has been taken in
VIVO~\cite{vivo_websci} to model, organize and integrate research
activities and researcher profile in an institutional setting.

The Ontology for Biomedical Investigations
(OBI)\footnote{\url{http://purl.obolibrary.org/obo/obi}} is an ongoing
effort aimed at developing an integrative ontology for biological and
clinical investigations. It takes a top-down approach by reusing
high-level, abstract concepts from other ontologies. It includes 2,600+
OWL classes and 10,000+ logical axioms (in the import closure of the OBI
ontology). OBI is very comprehensive and is suitable as an annotation
vocabulary for structured data. However, its size and complexity 
($\mathcal{SHOIN}(D)$) makes reasoning and querying of OBI-based 
ontologies and RDF graphs computationally expensive and time 
consuming\footnote{On a MacBook Pro with 2GB memory and an Intel 
Core 2 Duo 2.4 GHz processor, classifying the OBI ontology (version 
``2009-11-06'') takes more than 6 minutes using Pellet in Prot\'eg\'e. 
Such performance is clearly inadequate for a data management system.}, 
making it impractical as a domain model for a data management system 
where such reasoning may need to be performed repeatedly.

Functional Genomics Experiment Model (FuGe)~\cite{citeulike:1756058} is
an extensible modeling framework for high-throughput functional
genomics experiments, aiming at increasing the consistency and
efficiency of experimental data modeling for the molecular biology
research community. Centered around the concept of experiments, it
encompasses domain concepts such as protocols, samples and data. FuGe
is developed using UML from which XML Schemas and database definitions
are derived. The FuGe model covers not only biology-specific
information such as molecules, data and investigation; it also defines
commonly used concepts such as audit, reference and measurement.
Extensions in FuGe are defined using inheritance of UML classes.

We feel that the extensibility we require is not met by FuGe as any
addition of new concepts would require amendment of database schemas
and code. Moreover, the concrete objects reside in relational
databases, making subsequent integration and dissemination more
difficult.

\subsection{Motivation \& Goals}
Phenomics is a fast-growing, data-intensive discipline with new
technologies and processes rapidly emerging and evolving. As a result,
its domain model and data management systems must also be able to
evolve to handle the complexity, dynamics and scale of the data.

In phenomics, data is usually captured and measured by both high- and
low-throughput phenotyping devices. The scale of measurement can be
from the micro or cellular level, through the level of a single
organism, and up to the macro or field level. Imaging, measurement and
analysis of organisms on such a large scale will produce an enormous
amount of data.

Phenomics research makes use of a large variety of imaging and
measurement platforms. For example, in mouse histopathology and organ
pathology research, the Zeiss ``Mirax Scan'' scanner is used to scan
microscope slides. In clinical pathology, a Flow Cytometer is used to
capture laser diffraction images of blood samples. In plant research,
the Lemnatec Scanalyzer is used to capture RGB images of plants in
growth cabinets. The Fluorogroscan system is used in quenching
analysis: the partitioning of light energy used in photosynthesis on
model plants such as Arabidopsis. Other devices, such as the Infrared
Thermography Camera are used to capture leaf temperature and the SPAD
Meter is used to measure the chlorophyll content of plant leaves. New
devices and instruments will also be employed as they become available.
Moreover, existing instruments may be upgraded so that they can capture
more information. The PODD domain model needs to be flexible to
accommodate these continual changes in the formats, resolution and
source of the data.

Because an organism's phenotype is often the product of the organism's
genetic makeup, its development stage, disease conditions and its
environment, any measurement made against an organism needs to be
recorded in the context of these other metadata. Consequently the
opportunity exists to create a repository to record the data, the
contextual data (metadata) and data classifiers in the form of
ontological or structured vocabulary terms. The structured nature of
this repository will support both manual and autonomous data discovery
as well as provide the infrastructure for data based collaborations
with domestic and international research institutions. Currently there
are no such integrated systems available. The goals of PODD are to
capture, manage, annotate and distribute the data generated by mouse
and plant phenomics research activities.

\section{The Architecture of the Ontology-Centric Data Management
System}\label{sec:arch}

The most distinguishing characteristic of PODD is the central role that
ontologies play. In this architecture, raw data is not stored in a flat
structure but is attached to domain objects organized in a logical,
hierarchical system, defined according to the domain model that
represents the structure of research activities.

Current content management systems typically have a
relatively static domain model and hardwire it as relational schemas
and foreign key constraints in a custom relational database independent
from the underlying repository system. Consequently, the information
pertinent to each concrete object is stored in this custom database as
well. As stated in the previous section, this approach is unsuitable
for dynamic environments where conceptual changes are common.

To effectively support a dynamic conceptual framework, the domain model
in the proposed architecture is defined using OWL ontologies, in which:
OWL classes represent domain concepts; OWL properties define concept
attributes and their relationships; OWL restrictions specify
constraints on concepts and finally; OWL individuals define concrete
domain objects where attributes and relationships are defined using OWL
assertions. Raw data files are attached to concrete domain objects.

Such a conceptual architecture alleviates the problem of imposing hard
relational constraints in a database which is difficult to
extend/change.

Another drawback of existing systems is that there can be only one
domain model. When a concept needs to be updated, all the existing
objects defined by that concept need to be updated accordingly, which
may be undesirable, inappropriate and time-consuming. This is,
unfortunately, unavoidable as long as the domain model is defined using
database schemas. In our proposed architecture, as concept and object
definitions are stored in the repository, such changes can be versioned
so that existing instance objects can remain legitimate when integrity
validation is performed as they can still refer to the previous
conceptual definitions.

\begin{figure}[htb]
\centering
\includegraphics[trim = 60mm 30mm 50mm 38mm, clip,height=55mm]{architecture.pdf}
\vspace{-16pt} \caption{A high-level depiction of components in the
ontology-driven architecture.}\label{fig:arch}
\end{figure}

The high-level design of ontology-centric architecture takes a modular
and layered approach, as can be seen in Figure~\ref{fig:arch}. At the
foundation is the \textbf{data access layer}, consisting of an
underlying repository system, an RDF triple store, an in-house database
that stores essential information and a full-text search engine. This
layer is responsible for low-level tasks when the creation,
modification and deletion of concepts and objects occur. The
\textbf{business logic layer} in the middle is responsible for managing
concepts and objects, such as versioning, object conversion and
integrity validation. The \textbf{security layer} controls access
(authentication and authorization) to concepts and objects and guards
all operations on them.
%In this architecture, authorization is based on user attributes, which have two dimensions.
%Firstly, each user has a system-wide role, such as registered user or
%system administrator, which is used to determine access rights across
%the system. Secondly, a project-wide role, such as project
%administrator and project observer, can be assigned to a user so that
%he can have project-specific access rights.
At the top of the stack is the \textbf{interface layer}, where the data
management system can be accessed using a number of interfaces such as
a Web browser or API calls.

In developing the ontology-centric architecture, the following design
decisions have been made to balance expressivity, flexibility and
conceptual clarity. These decisions have also been based on a survey of
user requirements from scientists within a range of research
organizations including the Australian Plant Phenomics Facility (APPF)
as well as the Institute of Molecular Biology (IMB), Queensland Brain
Institute (QBI) and Australian Institute of Bioengineering \&
Nanotechnology (AIBN), working on collaborative research projects that
involve large scale data and distributed teams:

\begin{itemize}
\item There is a top-level domain concept, called Project , under which
other concepts (such as Investigation and Material) reside in a
hierarchical manner.

\item Access control (authorization) is defined on the Project level but
not on an individual object level, i.e., a given user will have the
same access rights for all objects within a given project.

\item Within a Project hierarchy, objects are in a parent-child
relationship in a tree structure such that each child can only have one
parent. This ensures that access rights are properly propagated from
parent to child and there is no chance of confusion.

\item Additionally, inter-object, many-to-many reference relationships
can be defined to enhance flexibility of the architecture as it allows
arbitrary links between objects to be established.

\item Objects cannot be shared across Projects. Instead, objects must be
copied from one project and pasted into another one. Such a rule
simplifies object management with the elimination of possible
side-effects caused by sharing object between projects.

\item There should be no interference between different versions of a
given concept and between objects that are instances of different
concept versions.
\end{itemize}

\section{Ontology-based Domain Modeling}\label{sec:ont}
As we emphasized previously, the domain model should be flexible enough
to accommodate the rapid changes and dynamic nature of scientific
research. In this section, we present the base ontology and the roles
it plays in the ontology-centric architecture. It should be noted that
the architecture proposed here is domain-independent and it can be
applied to any scientific discipline that shares a similar high-level
domain model.

Note that concepts in the ontologies presented here are models of entities
(activities and objects) in scientific investigations: they define the logical
structure of investigations - objects in an investigation and logic relationship 
between these objects. 
In a sense, the ontologies serve as a data model of the investigations in 
the scientific domain for which the data management system is developed.
In other words, the ontologies are used as a model for the data management 
system implementation.

\subsection{The Base Domain Ontology}
Inspired by FuGe~\cite{citeulike:1756058} and
OBI\footnote{\url{http://purl.obolibrary.org/obo/obi}}, we created the
base domain ontology in OWL to define essential domain concepts, their
attributes and inter-relationships in an object-oriented fashion. As
stated in the previous section, domain concepts will be modeled as OWL
classes; relationships between concepts and object attributes will be
modeled as OWL object and datatype properties. Concrete objects will be
modeled as OWL individuals.

\vspace{-16pt}
\begin{figure}[htb]
\centering
\includegraphics[trim = 5mm 30mm 5mm 46mm, clip,height=50mm]{ont_high.pdf}
\vspace{-16pt} \caption{Main concepts in the base
ontology in the parent-child relationship \emph{contains}.}\label{fig:ont_high}
\end{figure}

\vspace{-16pt}
For an overview, inter-relationships of some of the domain concepts in
this ontology are shown in Figure~\ref{fig:ont_high}. It is worthing 
pointing out that concepts in this figure are shown in the logical 
hierarchy but not the inheritance hierarchy: it describes the structure 
of a scientific investigation and
how different activities/objects in it are related to each other.
For brevity reasons, OWL object properties and cross
references between classes are not shown. We also defined the following
design principles for the domain ontology.

\begin{itemize}
\item All essential domain concepts are modeled as sub-classes of an
abstract top-level OWL class \emph{PODDConcept} that captures common
attributes and relationships.

\item All relationships between domain concepts are captured by domain
properties, which can be further divided into two property hierarchies,
one for parent-child relationships and the other for reference
relation-ships. Each of the two hierarchies have an abstract top-level
property, called \emph{contains} and \emph{refersTo}, respectively.

\item All parent-child relationships are modeled in a property hierarchy
as sub-properties of the abstract property \emph{contains}, and all
reference relationships are modeled in another property hierarchy as
sub-properties of the abstract property \emph{refersTo}.

\item For each domain concept \emph{C}, one property is defined in each of the
above hierarchies with its range defined to be \emph{C}. The domains of
such properties are not specified so that they can be used by any
applicable domain concept to establish a relationship between them.

\item Class attributes are modeled using OWL restrictions.

\item Essential domain concepts can be sub-classed to provide more
specialized and refined information.

\item To ensure that each object can have at most one parent object, the
inverse property of \emph{contains}, \emph{isContainedBy}, is defined
so that a max cardinality restriction can be added to the top-level
concept \emph{PODDConcept} to enforce it.
\end{itemize}

The definitions of some top-level constructs are summarized in
Figure~\ref{fig:top}, in OWL DL syntax~\cite{hoph03a}.

\vspace{-16pt}
{\small
\begin{figure}
\begin{align*}
& PODDConcept \sqsubseteq \top && \top \sqsubseteq \forall contains.PODDConcept\\
& isContainedBy \sqsubseteq (^- contains) && PODDConcept \sqsubseteq \leq 1~ isContainedBy\\
& \top \sqsubseteq refersTo.PODDConcept &&
\end{align*}

\vspace{-8pt}
\caption{Top-level ontology constructs in the PODD
ontology}\label{fig:top}
\end{figure}}

\vspace{-12pt}
In our model, we use OWL properties to model object attributes. When
the possible values of a particular attribute can be enumerated, such
as project status (\emph{active}, \emph{inactive} and
\emph{completed}), an enumerated OWL class is used to represent all the
values. When an attribute represents a grouping of some values, such as
accessions, where an accession has a source and a number, an OWL class
is also defined to represent the grouping. In this case, auxiliary OWL
properties are defined to project out specific values in the grouping.
In all other cases, attributes are modeled using datatypes.

Figure~\ref{fig:project_owl} shows the partial definition of the OWL
class $Project$. Restriction (\ref{for:hasPlan}), for example, states
that any $Project$ instance must have exactly one $ProjectPlan$
(through the predicate $hasProjectPlan$, the range of which is
$ProjectPlan$). The other 3 restrictions are similarly defined.

\vspace{-16pt}
\begin{figure}[htb]
\centering \small
\begin{align}
Project \sqsubseteq~ &=~ 1~ hasProjectPlan~ \sqcap \label{for:hasPlan}\\
        \sqsubseteq~ &\geq~ 1~ hasInvestigation~ \sqcap \\
        \sqsubseteq~ &=~ 1~ hasStartDate~ \sqcap\\
        \sqsubseteq~ &\leq~ 1~ hasPublicationDate
\end{align}
\vspace{-16pt} \caption{Partial OWL Definition for the Project
concept.}\label{fig:project_owl}
\end{figure}

\vspace{-24pt}
\subsection{Roles of Domain Ontologies in Object Life Cycle}
The base ontology defines essential concepts independent of the domain.
Domain-specific knowledge can be incorporated by extending the base
ontology for discipline-specific systems.

As stated in Section~\ref{sec:intro}, the ontology-based domain model
is at the center of the whole life cycle of objects. In this
subsection, we briefly describe the roles that the domain ontologies
perform at various stages of the object life cycle.

\begin{description}
\item[Ingestion]
When an object is created, its definition is expressed in
ontological terms. Such definitions will be used to (a) guide the
rendering of object creation interfaces and (b) validate the attributes
and inter-object relationships the user has entered before the object
is ingested. When an object is ingested, its definitions are stored as
RDF assertions.

\item[Retrieval \& update] When an object is retrieved from the
repository, its attributes and inter-object relations are retrieved
from its RDF assertions, which are used to drive the on-screen
rendering. When any value is updated, it is validated and updated in
this object's RDF assertions.

\item[Query \& search] An object's assertions will be stored in an RDF
triple store, which can be queried using SPARQL. Similarly, ontological
definitions are indexed to provide functionalities such as full-text
search and faceted browsing.

\item[Publication \& export] When an object is published or exported, its
metadata, in RDF, will be retrieved and exported.
\end{description}

\section{The PODD Data Management System}\label{sec:podd}
Based on the ontology-centric architecture presented in
Section~\ref{sec:arch} and the base ontology presented in
Section~\ref{sec:ont} we implemented the PODD data management system -
with the aim being to meet the data management challenges faced by the
Australian phenomics research community.

To describe domain knowledge in phenomics, we extend the base ontology
by defining additional concepts including \emph{Genotype}, \emph{Gene},
\emph{Phenotype} and \emph{Sequence} as subclasses of
\emph{PODDConcept}. Additional OWL object and datatype properties are
also defined to model the attributes and relationships of these
concepts, as shown in Figure~\ref{fig:podd_high}. Note that
\emph{Phenotype} is a subclass of \emph{Observation}.

\begin{figure}[htb]
\centering
\includegraphics[trim = 5mm 30mm 5mm 46mm, clip,height=50mm]{ont_podd.pdf}
\vspace{-16pt} \caption{Extended Domain Ontology for
phenomics.}\label{fig:podd_high}
\end{figure}

Figure~\ref{fig:phe_ont} shows some new definitions in the domain
ontology. Note that the last two definitions
integrate the new definitions with those in the base ontology.

Also note that the concepts defined in the PODD ontologies do not
necessarily represent real-world entities/reality. For example, the OWL class
\emph{Gene} does not intend to be a class that describes genes
in general. Rather, it is used to describe genes that are observed/involved
in scientific investigations.

\vspace{-16pt}
{\small
\begin{figure}[htb]
\centering
\begin{minipage}[t]{.55\linewidth}
\centering
\begin{align*}
Genotype &\sqsubseteq~ PODDConcept\label{for:hasPlan}\\
         &\sqsubseteq~ \forall hasGene.Gene\\
         &\sqsubseteq~ \leq~ 1~ hasEcotype\\
         &\sqsubseteq~ \leq~ 1~ hasSubspecies\\
         &\cdots\\
Project  & \sqsubseteq \forall hasGenotype.Genotype\\
Material & \sqsubseteq \forall hasPhenotype.Phenotype\\
         & \sqsubseteq \forall refersToGenotype.Genotype
\end{align*}
\end{minipage}
\begin{minipage}[t]{.4\linewidth}
\centering
\begin{align*}
Gene & \sqsubseteq~ PODDConcept\\
     & \sqsubseteq~ \forall hasSequence.Sequence\\
     & \sqsubseteq~ \leq~ 1~ hasAlias\\
     & \sqsubseteq~ \leq~ 1~ hasChromosome\\
     & \cdots
\end{align*}
\end{minipage}

\vspace{-4pt}
\caption{Domain-specific OWL definitions.}\label{fig:phe_ont}
\end{figure}
}

\vspace{-12pt}
In developing the PODD system, we chose to employ a number of mature
technologies. (1) We use Fedora Commons for the storage and retrieval
of domain objects. Together with raw data files, the OWL (for concepts)
and RDF (for objects) definitions of each concept and object are stored
in a versioned datastream PODD, which is used by the PODD system in
various tasks such as object creation, rendering, validation, update
and visualization. (2) We incorporate
the Sesame triple store\footnote{\url{http://www.openrdf.org/}} to
support complex query answering using SPARQL. Sesame contexts are used
to give scope to the RDF triples for each domain object. As described
in Section~\ref{sec:arch}, access control needs to be enforced on a per
project level. Similarly, it also needs to be enforced on query
answering in the triple store. By identifying triples of individual
objects, we are able to control contexts a user can access through
query expansion. (3) Lastly, we use the Lucene and Solr open-source search
engine platform\footnote{\url{http://lucene.apache.org/}} to provide
full-text search and faceted browsing capabilities. Similar to the
structure of the Sesame triple store, there is a one-to-one
correspondence between domain objects in the repository and the Solr
documents, the logical indexing units.

\vspace{-8pt}
\begin{figure}[htb]
\centering
\includegraphics[trim = 5mm 0mm 5mm 0mm, clip,height=76mm]{browser.png}
\vspace{-8pt} \caption{The browser view of a plant project in the PODD
repository.}\label{fig:browser}
\end{figure}

\vspace{-8pt}
Although the architecture and the system are based on ontologies, the
interface is designed to hide ontology-related complexity from the user
and present information in an easy to use manner for all repository
functions. For example, Figure~\ref{fig:browser} shows the browser view
of a plant phenomics project that investigates salt tolerance of wheat. 
In this view, the objects are shown in a tree-like structure by following
property assertions of subproperties of contains defined in  the base
and domain ontologies.

We have started to deploy the PODD system in Australian phenomics
research centers including APPF and APN and begun engaging users in the
evaluation of the performance, flexibility, usability and scalability
of the system. User feedback to date has shown that the system is
intuitive and efficient.

\section{Conclusion}\label{sec:concl}
In summary, our contribution to scientific data management is
three-fold: firstly, the proposal of the ontology-centric architecture
for developing data management systems; secondly, the development of a
base ontology that defines essential domain knowledge; and thirdly, the
development of the PODD data management system (based on both existing
and new technologies) that validates the feasibility of the proposed
approach.

We have identified a number of future work directions that we would
like to pursue. Firstly, we will investigate integration with existing
domain ontologies such as the Gene Ontology and the Plant Ontology. One
possibility would be to use terms defined in these ontologies to
annotate metadata objects. Secondly, we would like to investigate the
generalization of the ontology-centric approach so that it can be
applied to other areas such as workflow management systems. Thirdly, we
will continue the development of the PODD system to provide additional
functionalities such as data visualization, automated data integration
and Linked Data-style data discovery and publication.

%\section*{Acknowledgment}
%The authors wish to acknowledge the support of the National eResearch
%Architecture Taskforce (NeAT) and the Integrated Biological Sciences
%Steering Committee (IBSSC) Australia. The authors wish to thank Dr
%Xavier Sirault, Dr. Kai Xu and Mr. Philip Wu for the discussion and
%assistance in the development of the domain ontology and the PODD
%system.

\bibliography{all}
\bibliographystyle{abbrv}
\end{document}

### 2012-03-02 epncanberra2012/Podd-20120313.tex
first canberra presentation commit

\documentclass[12pt]{beamer}

\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\usepackage{beamerthemenirma}

\usepackage[australian]{babel}
\usepackage[T1]{fontenc}
\usepackage{graphicx}

\title{PODD Ontology Driven Database}
\author{Dr Peter Ansell}
\institute{University of Queensland}
\date{13 March 2012}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

\section{Scientific experiments}

\begin{frame}
\frametitle{Scientific experiments}

\begin{itemize}
 \item Aim to test hypotheses in known, partially controlled situations
\pause
 \item Contain many variables
\pause
 \item Experiment variables are noted in workbooks
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Scientific workbooks}

\begin{itemize}
 \item Some common headings
\pause
 \item Structure can vary
\end{itemize}

\end{frame}

\section{Structured data}

\begin{frame}
\frametitle{Tables}

\begin{itemize}
 \item Two dimensional structure
\pause
 \item Suit simple relationships
\pause
 \item Single table not ideal for entire experiment
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Trees}

\begin{itemize}
 \item Everything has a common root
\pause
 \item Branches are independent of each other
\end{itemize}

\end{frame}

\section{PODD Ontology Driven Database}

\begin{frame}
\frametitle{PODD as a workbook}

\begin{itemize}
 \item Vital metadata about each experiment
\pause
 \item Variable number of headings and subheadings
\pause
 \item Acceptable use of headings and subheadings is configurable
\pause
 \item Each workbook is independent of other workbooks
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Ontologies to define structure}

\begin{itemize}
 \item Ontologies can be arbitrarily structured
\pause
 \item PODD has basic constraints
\begin{enumerate}
 \item Each ontology must have exactly one Project
 \item Each node must be connected to the Project through a series of ``part\_of'' links
 \item Each node can reference other nodes using ``refers\_to'' links
\end{enumerate}
\pause
\item Ontology defines acceptable structure without imposing a single schema on every workbook

\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Phenomics ontology structure}

\begin{center}
 \includegraphics[scale=0.40,keepaspectratio=true]{./podd_ont.pdf}
 % podd_ont.pdf: 720x540 pixel, 72dpi, 25.40x19.05 cm, bb=0 0 720 540
\end{center}

\end{frame}

\end{document}

### 2012-03-08 epncanberra2012/Podd-20120313.tex
more changes to presentation

@@
 \usepackage[utf8x]{inputenc}
-\usepackage{beamerthemenirma}
+\usepackage{beamerthemeBerkeley}
 
@@
 \begin{itemize}
- \item Aim to test hypotheses in known, partially controlled situations
+ \item Aim to test hypotheses in known, partially controlled, situations
 \pause
@@
 \pause
- \item Experiment variables are noted in workbooks
+ \item Experiment variables and results are noted and used in analysis.
+\pause
+ \item Analysed results are used to make conclusions, which are published.
 \end{itemize}
-
 \end{frame}
@@
 \end{itemize}
-
 \end{frame}
@@
 \begin{frame}
-\frametitle{Tables}
+\frametitle{Relational databases}
 
@@
  \item Single table not ideal for entire experiment
+\pause
+ \item Not ideal for multiple different experiments
 \end{itemize}
-
 \end{frame}
@@
 \begin{frame}
-\frametitle{Trees}
+\frametitle{Structured ontologies}
 
 \begin{itemize}
- \item Everything has a common root
+ \item A common root
 \pause
- \item Branches are independent of each other
+ \item Branches and leaves are independent of each other
 \end{itemize}
-
 \end{frame}
@@
 \begin{itemize}
- \item Vital metadata about each experiment
+ \item Vital metadata about each experiment in a single ontology
 \pause
@@
 \pause
- \item Each workbook is independent of other workbooks
+ \item Each experiment is independent of other experiments
 \end{itemize}
-
 \end{frame}
@@
 \begin{itemize}
- \item Ontologies can be arbitrarily structured
+ \item Ontologies allow different structures for different experiments without changing the underlying database schema
 \pause
@@
  \item Each ontology must have exactly one Project
- \item Each node must be connected to the Project through a series of ``part\_of'' links
+ \item Each node must be connected to the Project through a unique series of ``part\_of'' links
  \item Each node can reference other nodes using ``refers\_to'' links
@@
 \pause
-\item Ontology defines acceptable structure without imposing a single schema on every workbook
-
+\item A single ontology can be used with different unique experiments if it is designed well
 \end{itemize}
-
 \end{frame}
@@
 \begin{center}
- \includegraphics[scale=0.40,keepaspectratio=true]{./podd_ont.pdf}
+ \includegraphics[scale=0.36,keepaspectratio=true]{./podd_ont.pdf}
  % podd_ont.pdf: 720x540 pixel, 72dpi, 25.40x19.05 cm, bb=0 0 720 540
 \end{center}
+\end{frame}
+
+\begin{frame}
+\frametitle{Publication}
+
+\begin{itemize}
+ \item Projects can be published on completion, both in HTML and RDF
+\pause
+ \item Published projects designed to be immutable
+\pause
+ \item Each data item has a URI
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Questions}
+
+Code can be found online at:
+
+\url{https://github.com/podd}
 

### 2012-03-08 epncanberra2012/Podd-20120313.tex
final draft before todays presentation

@@
 \pause
- \item Experiment variables and results are noted and used in analysis.
+ \item Experiment variables and results are noted and used in analysis
 \pause
- \item Analysed results are used to make conclusions, which are published.
+ \item Analysed results are used to make conclusions, which are then published
 \end{itemize}
@@
 \pause
- \item Not ideal for multiple different experiments
+ \item Not ideal for multiple structurally different experiments
 \end{itemize}
@@
 \begin{itemize}
- \item Ontologies allow different structures for different experiments without changing the underlying database schema
+ \item Different structures for different experiments without changing the database schema
 \pause
- \item PODD has basic constraints
+ \item PODD has basic ontology constraints
 \begin{enumerate}
@@
  \item Each node must be connected to the Project through a unique series of ``part\_of'' links
- \item Each node can reference other nodes using ``refers\_to'' links
+ \item Each node can reference other eligible nodes using ``refers\_to'' links
 \end{enumerate}
 \pause
-\item A single ontology can be used with different unique experiments if it is designed well
+\item A single ontology can be used with different experiments if it is designed well
 \end{itemize}
@@
 \pause
- \item Published projects designed to be immutable
+ \item Published projects may be locked from future changes
 \pause
- \item Each data item has a URI
+ \item Each data item has a URI for references
 \end{itemize}

### 2012-03-09 epncanberra2012/Podd-20120313.tex
new cutdown skeleton structure as the old one didn't appeal to eresearch group

--- 2012-03-02 epncanberra2012/Podd-20120313.tex
@@
 \usepackage[utf8x]{inputenc}
-\usepackage{beamerthemenirma}
+
+%\usepackage{beamerthemeBerkeley}
+\usetheme{Berkeley}
 
@@
 
-\begin{frame}{Outline}
-  \tableofcontents
-  % You might wish to add the option [pausesections]
-\end{frame}
-
-\section{Scientific experiments}
-
 \begin{frame}
-\frametitle{Scientific experiments}
-
-\begin{itemize}
- \item Aim to test hypotheses in known, partially controlled situations
-\pause
- \item Contain many variables
-\pause
- \item Experiment variables are noted in workbooks
-\end{itemize}
+\frametitle{Background}
 
@@
 \begin{frame}
-\frametitle{Scientific workbooks}
-
-\begin{itemize}
- \item Some common headings
-\pause
- \item Structure can vary
-\end{itemize}
-
-\end{frame}
-
-\section{Structured data}
-
-\begin{frame}
-\frametitle{Tables}
-
-\begin{itemize}
- \item Two dimensional structure
-\pause
- \item Suit simple relationships
-\pause
- \item Single table not ideal for entire experiment
-\end{itemize}
+\frametitle{Motivation}
 
@@
 \begin{frame}
-\frametitle{Trees}
-
-\begin{itemize}
- \item Everything has a common root
-\pause
- \item Branches are independent of each other
-\end{itemize}
-
-\end{frame}
-
-\section{PODD Ontology Driven Database}
-
-\begin{frame}
-\frametitle{PODD as a workbook}
-
-\begin{itemize}
- \item Vital metadata about each experiment
-\pause
- \item Variable number of headings and subheadings
-\pause
- \item Acceptable use of headings and subheadings is configurable
-\pause
- \item Each workbook is independent of other workbooks
-\end{itemize}
+\frametitle{Example}
 
@@
 \begin{frame}
-\frametitle{Ontologies to define structure}
+\frametitle{Design}
 
-\begin{itemize}
- \item Ontologies can be arbitrarily structured
-\pause
- \item PODD has basic constraints
-\begin{enumerate}
- \item Each ontology must have exactly one Project
- \item Each node must be connected to the Project through a series of ``part\_of'' links
- \item Each node can reference other nodes using ``refers\_to'' links
-\end{enumerate}
-\pause
-\item Ontology defines acceptable structure without imposing a single schema on every workbook
+\end{frame}
 
-\end{itemize}
+\begin{frame}[t]
+\frametitle{Phenomics ontology for PODD}
+\invisible
+
+\begin{center}
+ \includegraphics[scale=0.36,keepaspectratio=true]{./podd_ont.pdf}
+ % podd_ont.pdf: 720x540 pixel, 72dpi, 25.40x19.05 cm, bb=0 0 720 540
+\end{center}
+\end{frame}
+
+\begin{frame}
+\frametitle{Demo}
 
@@
 \begin{frame}
-\frametitle{Phenomics ontology structure}
+\frametitle{Evaluation}
 
-\begin{center}
- \includegraphics[scale=0.40,keepaspectratio=true]{./podd_ont.pdf}
- % podd_ont.pdf: 720x540 pixel, 72dpi, 25.40x19.05 cm, bb=0 0 720 540
-\end{center}
+\end{frame}
+
+\begin{frame}
+\frametitle{Next steps}
+
+\end{frame}
+
+\begin{frame}
+
+\url{http://github.com/podd}
+
+peter.ansell@uq.edu.au
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Questions}
+
+Open source code can be found online at:
+
+\url{https://github.com/podd}
+
+My email is: peter.ansell@uq.edu.au
 

### 2012-03-09 epncanberra2012/Podd-20120313.tex
make a blank slide with a bigger version of the podd ontology which should scale with screen size
fix second to last slide
add architecture diagram as svg and png

@@
 
+\begin{itemize}
+ \item PODD implemented between 2009 and 2011
+ \item Design by Yuan Fang Li and Gavin Kennedy
+\end{itemize}
+
 \end{frame}
@@
 
-\begin{frame}
-\frametitle{Design}
+% PODD Design/Architecture as a blank slide
+\bgroup
+\usebackgroundtemplate{\includegraphics[
+width=\paperwidth,
+height=\paperheight,
+keepaspectratio=true
+]{podd_arch.png}}
+\begin{frame}[plain]{}
+\end{frame}
+\egroup
 
+% PODD Phenomics Ontology as a blank slide
+\bgroup
+\usebackgroundtemplate{\includegraphics[
+width=\paperwidth,
+height=\paperheight,
+keepaspectratio=true
+]{podd_ont.png}}
+\begin{frame}[plain]{}
 \end{frame}
-
-\begin{frame}[t]
-\frametitle{Phenomics ontology for PODD}
-\invisible
-
-\begin{center}
- \includegraphics[scale=0.36,keepaspectratio=true]{./podd_ont.pdf}
- % podd_ont.pdf: 720x540 pixel, 72dpi, 25.40x19.05 cm, bb=0 0 720 540
-\end{center}
-\end{frame}
+\egroup
 
@@
 
+Good:
+
+\begin{enumerate}
+ \item Flexible : Simultaneously support different experiments
+ \item Adaptable : Additions and changes to schema ontologies
+\end{enumerate}
+
+Bad:
+
+\begin{enumerate}
+ \item Current implementation does not scale
+ \item Object oriented, pulling document into memory
+\end{enumerate}
+
 \end{frame}
@@
 \begin{frame}
-
-\url{http://github.com/podd}
-
-peter.ansell@uq.edu.au
-
-\end{frame}
-
-\begin{frame}
 \frametitle{Questions}
 
+\begin{center}
 Open source code can be found online at:
@@
 
+\vskip 12pt
+
 My email is: peter.ansell@uq.edu.au
-
+\end{center}
 \end{frame}

### 2012-03-12 epncanberra2012/Podd-20120313.tex
more changes to canberra presentation
add example and demo, although the example is currently sited in the demo

@@
  \item PODD implemented between 2009 and 2011
- \item Design by Yuan Fang Li and Gavin Kennedy
+ \item Design by Yuan Fang Li and Gavin Kennedy for High Resolution Plant Phenomics Centre
 \end{itemize}
@@
 
+\begin{itemize}
+ \item Flexible scientific experiment management
+ \item Use RDF and OWL technology to support science
+\end{itemize}
+
 \end{frame}
@@
 \frametitle{Example}
+
+Simultaneous Phenotyping of Drought Stress Tolerance in Arabidopsis OST1-2 Mutant and Wild Type
+
+\vskip 12pt
+
+Credit to Xueqin Wang for the example design : \url{http://podd.plantphenomics.org.au/podd/object/poddObject:838}
 
@@
 
+\url{http://podd.plantphenomics.org.au/podd}
+
 \end{frame}
@@
 \begin{enumerate}
- \item Current implementation does not scale
- \item Object oriented, pulling document into memory
+ \item Current implementation does not scale with experiment size
+ \item Only supports OWL-1.1
+ \item Uses old versions of Fedora and Spring
 \end{enumerate}
@@
 \frametitle{Next steps}
+
+\begin{itemize}
+ \item Use upcoming SPARQL 1.1 Query and Update standards
+ \item Use OWL-2 and pure SPARQL access to databases
+ \item Support links from experiments to other RDF documents
+\end{itemize}
 

### 2012-03-12 epncanberra2012/Podd-20120313.tex
more minor edits
grammar check

@@
 
-%\usepackage{beamerthemeBerkeley}
-\usetheme{Berkeley}
-
 \usepackage[australian]{babel}
@@
 \usepackage{graphicx}
+
+\usetheme{Berkeley}
 
@@
  \item PODD implemented between 2009 and 2011
- \item Design by Yuan Fang Li and Gavin Kennedy for High Resolution Plant Phenomics Centre
+ \item Design by Yuan Fang Li and Gavin Kennedy for the High Resolution Plant Phenomics Centre
 \end{itemize}
@@
 
-Good:
+Pros:
 
 \begin{enumerate}
- \item Flexible : Simultaneously support different experiments
- \item Adaptable : Additions and changes to schema ontologies
+ \item Flexible : Simultaneously supports different experiments
+ \item Adaptable : Supports additions and changes to schema ontologies
 \end{enumerate}
 
-Bad:
+Cons:
 
@@
  \item Use upcoming SPARQL 1.1 Query and Update standards
- \item Use OWL-2 and pure SPARQL access to databases
+ \item Support OWL-2
+ \item Pure SPARQL access using a single database
  \item Support links from experiments to other RDF documents

### 2012-03-25 csiroala2012/Podd-20120402.tex
add csiro presentation based on hrppc presentation

--- 2012-03-12 epncanberra2012/Podd-20120313.tex
@@
 \institute{University of Queensland}
-\date{13 March 2012}
+\date{2 April 2012}
 
@@
 \begin{itemize}
- \item PODD implemented between 2009 and 2011
- \item Design by Yuan Fang Li and Gavin Kennedy for the High Resolution Plant Phenomics Centre
+ \item PODD was designed by Yuan-Fang Li and Gavin Kennedy
+ \item Implemented between 2009 and 2011
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Next steps}
+\frametitle{Redesign}
 
@@
  \item Support OWL-2
- \item Pure SPARQL access using a single database
- \item Support links from experiments to other RDF documents
+ \item Pure SPARQL access using a single database, including for management information to enable unified queries
+ \item Support links between objects in PODD
 \end{itemize}

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
initial submission tex structure

\documentclass{llncs}

\begin{document}

\title{Flexible Scientific Data Management for Plant Phenomics Research}

%Autors: Peter Ansell1, Robert Furbank2, Kutila Gunasekera1, Jianming Guo2, David Benn3, Gareth Williams3, Xavier Sirault2
\author{Peter Ansell\inst{1}, Robert Furbank\inst{2}, Kutila Gunasekera\inst{1}, Jianming Guo\inst{2}, David Benn\inst{3}, Gareth Williams\inst{3}, Xavier Sirault\inst{2}}

%Affiliation:
\institution{
      eResearch Group, School of Information Technology and Electronic Engineering,
             University of Queensland, Brisbane, Australia
 \and CSIRO Plant industry, High Resolution Plant Phenomics Centre, Canberra, Australia
 \and CSIRO IMT Advanced Scientific Computing and Research Data Services, Melbourne, Australia}

\begin{abstract}

In this paper, we expand on the design and implementation of the Phenomics
Ontology Driven Data repository [13] (PODD) with respect to the capture, storage
and retrieval of data and metadata generated at the High Resolution Plant
Phenomics Centre (Canberra, Australia). PODD is a schema-driven Semantic Web
database which uses the Resource Description Framework (RDF) model to store
semi-structured information. RDF allows PODD to process information about a
range of phenomics experiments without needing to define a universal schema for
all of the different structures. To illustrate the process, exemplar datasets
were generated using a medium throughput, high resolution, three-dimensional
digitisation system purposely built for studying plant structure and function
simultaneously under specific environmental conditions. The High Performance
Compute (HPC), storage and data collection publication aspects of the workflow
and their realisation in CSIRO infrastructure are also discussed along with
their relationship to PODD.

\end{abstract}

%\keywords{eResearch, Semantic Web, RDF, OWL, Data collection citation, BagIt, Data Access Portal}

\section{Introduction}
Since the genomics era, biology has become a data-driven science. Advances in
robotics, automation and imaging, in combination with high performance computing
have permitted the rapid production of large and complex biological datasets.
Currently, high volumes of heterogeneous image data, physiological and
morphological measurements are being acquired by a range of new phenotyping
platforms located in purpose built phenomics centres across the world. These
large datasets of phenotypic characters such as growth rate, plant architecture,
photosynthetic performance, yield must be stored and related to genotype. This
link is essential for expediting the identification of genetic variation in
natural and derived genetic populations (e.g. germplasm collections, association
genetic panels, recombinant inbred lines), understanding the dynamic
relationship between phenotype, genotype and environment and delivering the
increase in productivity necessary for feeding the world.

To achieve this goal, the vast array of phenotypic data collected from a variety
of phenomics platforms must be combined with metadata explaining how the raw
information was collected. This process forms structured digital objects
containing the raw information. These objects are then delivered to a range of
analysis pipelines, which transform the raw information into  aggregated
multi-phase datasets, each phase representing a new aggregation or inference
from the original raw data. This reduction process converts the raw
multi-dimensional data into information which is conceptually interpretable by a
human being, i.e. new knowledge. The details about the steps taken are recorded
to give context to the data. These links enable scientists to generate novel
conclusions about the plants being phenotyped, including conclusions based on
the genotype, plant growth environments, treatments, phenotyping platforms and
the phenotyping and analysis processes.

To make sense of this large amount of information, sophisticated storage,
archiving, searching and analysis capabilities are required. To date solutions
to this problem have been handled essentially by private companies, and no
suitable solution exists in the public domain. Lack of systems, both to manage
linked metadata, and controlled vocabularies to describe plant growth and
experimental conditions, have severely hampered sharing of plant phenomics data,
comparison of results between laboratories and the capacity to carry out
meta-analysis of existing data sets.

Thus, to support publicly-funded phenomics activities in Australia, the
Phenomics Ontology Driven Data repository (PODD) has been developed as a
repository for data produced by the variety of plant imaging and phenotyping
platforms available at the High Resolution Plant Phenomics Centre, as well as
for recording the contextual metadata associated with plant genotypes,
treatments and environmental conditions [13]. PODD manages the links between the
raw image files generated by plant imaging platforms, the results from analyses,
and conclusions by scientists.

In this paper, we describe the workflow management that the High Resolution
Plant Phenomics Centre (HRPPC) has implemented for keeping track of its
phenomics data, metadata and experimental processes. This complex challenge was
addressed by building a multi-disciplinary group of information technology
experts and embedding users of phenomics technologies into it. The result of the
approach is a state of the art computational and data mining environment,
optimised for data access, data discovery and data sharing, which also provides
the flexibility for linking genomic information through the use of RDF triples.
In this context, we also describe the role of the CSIRO Data Access Portal (DAP,
http://data.csiro.au) to annotate and store raw and processed datasets. DAP also
provides long term secure storage for data collections and the ability to search
for, control access to, and cite them via Digital Object Identifiers. PODD
manages the mapping of collections located in DAP to PODD projects, providing
for the storage of large images and documents unsuited to RDF databases. Figure 1. shows
the relationship between components and key data flows.

%% FIXME: TODO: Insert Figure 1 here

Figure 1. HRPPC component relationships and data flow

\section{Phenomics Ontology Driven Data repository}
\subsection{Semantic science for phenomics data management}
Scientists have focused on including semantics into datasets, typically using
the foundations of RDF and OWL, from two main directions. Some focus on defining
ontologies based on hierarchies of scientific concepts and properties, while
others have focused on mapping complex scientific datasets to RDF using syntax
transformations without initially defining the semantic meaning of the results.
In reality, most efforts fall somewhere in the middle, with ontological
annotations attached to some data points while other nearby data points are
syntactically represented using RDF, without links to ontologies of scientific
concepts.

Increasingly however, providers of scientific datasets are focusing on enhancing
their datasets using curated scientific concepts from ontologies. For example,
scientists have used the Gene Ontology [4] to link well known concepts to
represent common elements across genomics datasets. If these annotations are
given using recognised URIs, and not as key-value string tuples, various RDF
documents can be merged automatically based on RDF semantics to discover novel
conclusions. Initially, different providers were not able to decide on URIs to
use in Linked Data. This resulted in various third parties such as Bio2RDF [12]
creating RDF versions of scientific datasets so that the RDF documents could be
easily merged based on common URIs. However, recently more providers have given
authoritative URIs for each object in their dataset, making it easy to integrate
data from various locations.

\subsection{Implementation of the Phenomics Ontology Driven Data repository}
As described earlier, the PODD repository relies on semantic web technologies to
manage phenomics data and metadata. Although both ontologies and mappings are
essential, in PODD it was necessary to build the system with a relaxed
ontological vocabulary to allow scientists to generate datasets before they were
certain about the semantic meaning of their work. In addition, it was necessary
for PODD to allow scientists to continue to maintain projects containing curated
scientific concepts alongside raw experimental data.

The basic PODD project is structured as a tree, which, when used with
PlantScanTM, represents the different components of a scientific project as
top-level branches. These include a branch for raw data, along with separate
branches for results, analysis, and publications related to the project. In the
case of raw data, the semantics are not clear and cannot be defined by the
platforms collecting the data. Scientists are able to sparsely attach meaning
using horizontal links between different branches, using the OWL Open World
Assumption to allow extra semantic annotations to be attached in future. For
example, they may annotate images of a plant with a trait in only one case,
without annotating other images. These horizontal links are not constrained by
the basic scientific project structure, so they are able to entail novel
meanings independent of the ontology, including untested hypotheses, process
annotations and meaningful scientific results.

\subsection{Semantic validation}
PODD validates experiments using independently configurable rules based on OWL
(Web Ontology Language) schemas. Although PODD currently only supports OWL as a
rules language, it could be easily extended in other cases to use different
systems such as N3, RDFS, SPARQL, or SPIN as rules languages.

OWL is used to determine whether projects are both internally consistent, with
all objects having an explicit RDF type, and whether they are consistent with
the ontologies that they import. The main points that PODD focuses on for
validation is determining that mappings are consistent with the basic PODD
predicates and classes, so that they can be used to render both static HTML
pages and provide HTML based editing facilities. For example, any OWL object
property that has been defined to link from image acquisition runs to images can
be used to define the purpose of an image. This may include any combination of
the defined properties, without requiring all of the properties to be used.

The parent object types and properties used by PODD to establish basic ground
rules for rendering and validation are defined in a small base schema ontology.
General scientific properties and phenotype specific properties are then defined
in larger extension ontologies as illustrated in Figure 2.

%%% FIXME: TODO: Insert Figure 2 here

Figure 2. PODD schema ontology hierarchy

An ontology is a common controlled vocabulary about a particular topic, in this
case plant phenomics. The PODD Plant ontology therefore becomes an ontology for
plant phenomics that extends the generally applicable PODD Science ontology with
plant phenomics specific details [13]. This approach allows PODD to model
multiple phenomics processes without enforcing a rigid data structure on
objects, either now, in the past, or in the future. In addition, other community
ontologies, such as the plant and gene ontologies [3, 4], can be used to
annotate the data. This not only supports data discovery, but provides common
reference points through which PODD managed data can be integrated with data in
other research databases.

The PODD application handles the instantiation and validation of objects based
on the versions of ontologies that are specified by the user as being relevant.
This gives PODD a unique extensibility as the domain knowledge is entirely
captured in the ontology, and the operational aspects of the system are derived
purely from the content of the ontology. In particular, PODD is designed around
the ability to support multiple versions of ontologies concurrently. Previously
published artifacts can be loaded without modification into PODD without
affecting current studies that may be using newer, incompatible, versions of the
schema ontologies.

PODD is able to manage high volume, high resolution, heterogeneous datasets
using a tree based structure, which in the PODD Science ontology is based on the
concept of a ``Project'' (http://podd.plantphenomics.org.au/podd/). Data
references are inserted into projects using subproperties of the PODD
``contains'' predicate. When results are available they are inserted
independently and cross-referenced with the data that they were derived from
using a subproperty of the PODD ``refers to'' property. Each of the
subproperties have constraints on the types of objects that are valid as sources
and destinations, making it easy to both validate existing uses, and suggest
potential targets based on a source and a property.

The genotype of the plant alone does not determine the phenotype; it is the
combination of the organism’s genome with the environmental conditions that
determine the phenotype. PODD captures metadata on the genotype of the
organism(s) under investigation. PODD also gathers comprehensive metadata on the
environmental conditions and treatments that the biological samples are exposed
to. This structure makes PODD suitable to queries by web agents thus  linking
information contained in other data repositories, e.g genomics, transcriptomics,
metabolomics to phenomics information.

\section{The PlantScanTM digitisation platform}
% See unresolved comment at the end

%System description and Processes

\subsection{Heterogeneous data streams}

PlantScanTM is a medium throughput high resolution phenotyping platform, which
brings together a number of imaging sensors (Light detection and ranging, Far
infrared imaging, multi-wavelength imaging) to non-invasively measure plant
growth and function using in-silico approaches [REF?]. Raw data is captured with
its contextual information (e.g. system configuration, time of acquisition,
batch number and project) and is stored in a purpose-built database as the data
is being generated. The various data streams are collated and used to produce
full 3D representation of each plant with overlaid spectral information. The
metadata collected during image acquisition are necessary inputs for the
computer vision techniques which are used to create the 3D representation of the
plant (Figure 3)[e]. The 3D meshes are then automatically segmented in order to
semantically identify the different parts of the plants (Paproki et al, 2012). A
longitudinal 3D matching pipeline for plant mesh parts is then used to evaluate 
temporal changes at the whole plant and/or organ level. After initial
processing, i.e. 3D reconstruction, the raw data for each plant is packaged into
one location using BagIt, an hierarchical file [6]. All bags belonging to the
same specific PODD project are grouped into collections on the CSIRO Data Access
Portal. These bags have a permanent address which is recorded into PODD. Any
additional processes on these digital objects (such as statistical information
on features like plant colour, volume,  growth etc. and on elements such as
petiole, stem, leaf) are also captured by PODD (Figure 1). Metadata
Each acquisition on PlantScanTM includes metadata (in addition to the raw data
streams), such as plant genus and species, project and experiment metadata, a
unique identifier for each image (Globally Unique Identifier), imaging angle,
environmental temperature of the imaging chamber, location of optical and colour
calibration datasets for each acquisition run, LiDAR (Light Detection and
Ranging Sensors) calibration files. The metadata associated with each
acquisition is automatically generated when setting up the configuration on the
platform. This information is paramount to validate and process the raw image
data, and for the post-processing phases. This information currently resides in
the PlantScanTM database.

\subsection{Data volume}
Digitisation systems such as PlantScanTM generate huge amounts of data including
raw image data, registration metadata, sensor configurations and plant metadata.
For example, PlantScanTM generates around 500GB of raw image data, representing
in excess of 200,000 database records, per day. Sufficient storage space
(usually at remote locations) and fast network transfer rates were thus
necessary to facilitate data movement for processing using high performance
computers (HPC). Because an RDF database structure is not suitable for handling
large data sets of images, it was necessary to package the raw information into
elementary units with permanent addresses which could be retrieved using PODD. 
The CSIRO Data Access Portal (DAP) \footnote{http://data.csiro.au} and Advanced
Scientific Computing (ASC) storage and compute facilities
\footnote{https://wiki.csiro.au/display/ASC} are key resources used by PlantScanTM to
process and store bulk data.

\section{CSIRO High Performance Compute Infrastructure}
CSIRO ASC shared facilities [10] are used to process the raw PlantScanTM data to
derive data products (meshes). Raw data and meshes are ``bagged'' and stored in
the ASC archival system. ASC High Performance Compute (HPC) hosts (systems with
high processor count and large memory) are taken advantage of to create and
verify bags more rapidly than would be possible on conventional computer
systems.

\subsection{BagIt}

BagIt is defined by an Internet Engineering Task Force (IETF) document as an
``hierarchical file packaging format for storage and transfer of arbitrary
digital content''[6]. A payload manifest details content and MD5 or SHA hashes
for content integrity verification. Data file related metadata can be stored in
pre-defined files as key-value pairs. For PlantScanTM, file-level metadata
includes plant barcodes, batch numbers, and plant type.

The BagIt specification does not mandate a particular archiving strategy, the
focus being upon the directory structure, special files, and integrity checking.
Bag creation and payload manifest integrity checking, tools can take advantage
of the multi-core capabilities of modern computing hardware. To suit current
needs, BagIt-conforming tools [7] [8] [9] were assessed and where necessary,
improvements were implemented and tested. For example, the implementation of [8]
assumed that as many threads as possible should be created for multi-threaded
operations such as bag creation and payload manifest verification. On a HPC host
with hundreds of processors, this can lead to runaway thread creation, so this
needed to be addressed. Neither [7] nor [8] permitted payload files in a bag to
be listed. This has since been added to [8]. Some functionality (not discussed
here) present in [8] but not in [7] was added to the latter.

\section{CSIRO Data Access Portal}
CSIRO's Research Data Service (RDS) has developed the Data Access Portal (DAP),
an open source web application that enables research data to be discovered,
managed and shared. [5]

Researchers can describe a data collection, deposit data, choose a license, and
add attribution details. Access to a collection’s description and/or data can be
restricted to CSIRO or a set of individuals (within CSIRO or partner
organisations) or it can be made public, becoming searchable by anyone via the
Internet. In the case where a collection and its data are public, a Digital
Object Identifier (DOI) is issued and can be used to formally cite the
collection in a publication. Collections can either be one-time, self-serve
deposits or automated, ongoing deposits and can be changed over time through
versioning. Whole datasets or specific files within a collection can be
downloaded. A machine-to-machine REST interface permits programmatic search and
collection download.

DAP is the visible part of the system. Other than standard web servers,
databases and application servers, additional required server-side
infrastructure includes petabyte-level, multi-decade, backed-up disk storage,
integrity checking, and automated migration of data collections to and from tape
storage.

\subsection{Bag preparation for a DAP collection}
CSIRO’s HRPPC makes use of DAP to store collections of PlantScanTM raw images
and processed mesh data as bags [6]. Raw data from PlantScanTM local storage
(HRPPC-Store) and data processed on HPC hosts are transferred to ASC bulk
storage where image and mesh files are organised in folders by batch, then
barcode number, then subfolders for each image file type (e.g. RGB images, IR
images, LiDAR) and meshes. Bag creation is carried out via an allocated ASC HPC
job.

Currently, one bag is equivalent to a single batch scanned on the PlantScanTM
local software system, which usually means the same kind of plant with different
genotypes scanned under one  experiment configuration profile.  However the
naming, content and amount of data in a bag are flexible and depend upon the
intention of the bag. For example, multiple batches can be packed into one bag
when the batches come from one plant’s scans at different time points. The
multi-threaded BagIt tool is called on the ASC HPC host to create a bag that is
then compressed into a single file containing a top-level data folder, a
manifest file that contains checksums for every data file, and tag files that
contain the metadata description of the bag just packed.

The metadata required for a DAP publication is created and the bag transferred
to the DAP staging area via SFTP (SSH File Transfer Protocol). After publication
of the DAP collection, the data from PlantScanTM for the given project becomes
discoverable via DAP. In addition, experiment reports, published papers, and
sensor configurations can either be made accessible via a DAP collection's
``related materials'' links, other metadata fields, or within the collection's
data (e.g. bag).

\section{Semantic integration}
The PODD ontology enables plant phenomics researchers to link from mesh results
to the raw data that they were generated from. It also allows researchers to
link from both mesh results and their recorded conclusions to shared phenomics
ontologies which describe specific features of the plants. When used together,
this enables scientists to trace the provenance of their results and conclusions
based on well known concepts in phenomics ontologies.

The PODD ontology allows sparse annotation of data and results, so scientists
are not required to annotate all of their data with links to shared ontologies
by default. However, the system is extensible, so in future such a requirement
could be added in a new version of the PODD ontology while maintaining backwards
compatibility with previous experiments that may be published or immutable.

External ontologies must currently be mapped into the PODD system to define the
expected status of the classes and properties with respect to the basic PODD
concepts. The basic concepts are used to both verify that each PODD project has
a single valid tree structure, and render HTML pages for both browsing and
editing. PODD also supports an RDF-based HTTP REST interface that clients may
use to fetch and update objects without using the HTML pages.

\section{Semantic publication}
PODD provides a secure mechanism for publishing both human and machine readable
descriptions of scientific experiments. It utilises the well-known DOI mechanism
for publishing raw data files using DAP, and uses HTTP URIs to publish
experiments using the PODD web interface.

PODD implements support for Content Negotiation, with common RDF serialisations
supported in addition to HTML pages that are encoded with RDFa (RDF in
Attributes). Although most non-human agents will receive an RDF formatted
document when they resolve a PODD URI, if a PODD HTML page is saved the
annotations can still be retrieved using RDFa.

Scientific journals increasingly require the data and provenance for articles to
be available in a machine readable format. The DOI registrar that DAP uses,
DataCite [11], was setup to provide unique identifiers for data items that can
be attached to publications, which in turn may have their own DOIs.

By providing machine readable descriptions of scientific experiments, including
semantic references to shared ontologies where possible, PODD enables the output
from PlantScanTM to be interpreted and extended by others. The use of PODD URIs
in other RDF documents enables scientists to extend the initial work using the
Linked Data paradigm [14].

\section{Conclusion}
This paper described how the Phenomics Ontology Driven Data repository
integrates with the PlantScanTM platform and CSIRO Data Access Portal to manage
the complex workflows at the High Resolution Plant Phenomics Centre. This
workflow keeps track of phenomics data, metadata and experimental processes and
also provides a secure mechanism to share and publish scientific experiments in
both human and machine readable formats.

References
[1] http://www.iucr.org/\_\_data/iucr/lists/epc-l/pdf00003.pdf
[2] http://topbraid.org/spin/api/
[3] http://www.plantontology.org/
[4] http://www.geneontology.org/
[5] https://wiki.csiro.au/confluence/display/dmsdoc/Home
[6] http://datatracker.ietf.org/doc/draft-kunze-bagit/
[7] https://pypi.python.org/pypi/bagit/
[8] http://sourceforge.net/projects/loc-xferutils/
[9] http://sourceforge.net/projects/loc-xferutils/files/loc-bagger/
[10] https://wiki.csiro.au/display/ASC
[11] doi:10.1109/COINFO.2009.66 
[12] doi:10.1016/j.jbi.2008.03.004
[13] doi:10.1007/978-3-642-13654-2\_22
[14] http://www.w3.org/DesignIssues/LinkedData.html

\end{document}
% Unresolved comments from Google docs:
%
% [a]Xavier Sirault:
% NOTE:Submitted papers should not exceed 15 pages for long papers and 8 pages for
% short papers and must be formatted according to the LNCS guidelines. 
%  
% Proceedings of the workshop will be published online at CEUR Workshop
% Proceedings (www.CEUR-WS.org ). The 3 best papers presented at the workshop will
% be included in the supplementary Springer LNCS (Lecture Notes in Computer
% Science ) proceedings of the ESWC2013 conference.
% ________________
% ansell.peter:
% I will reformat the document tommorrow morning using Latex and distribute the
% PDF to others for a final check.
% [b]Xavier Sirault:
% currently 78 words
% ________________
% Xavier Sirault:
% now 126 words
% ________________
% Xavier Sirault:
% 136 words
% [c]David Benn:
% "DAP" has been removed from the abstract. I know it looms large in the paper but
% is it worth adding something like "and data collection publication"? I've added
% this in but I understand if it's not okay because we're space challenged.
% ________________
% Xavier Sirault:
% No worries, I think the paper reads quite well. I would look at it tomorrow
% again
% ________________
% Xavier Sirault:
% to try to link some loose ends
% ________________
% David Benn:
% Okay, good. I'll do the same.
% ________________
% David Benn:
% I've read through, made suggestions for correction, chatted with Peter about
% them. I'll check back later tonight.
% [d]Xavier Sirault:
% I do not like this title. I do not want the paper to describe plantscan per se
% but more the acquisition processes and how it is captured.
% ________________
% David Benn:
% How about something like: "Plant phenomics imaging, processing, metadata capture
% and publication". Too verbose I know.
% [e]Xavier Sirault:
% Jianming, could you please put a 3D model of tobacco with overlaid RGB signal?
% [f]Xavier Sirault:
% This figure is figure 1. I think we should put it there rather than at the end
% of the introduction
% ________________
% David Benn:
% Do we want to simply change "Figure 4" to "Figure 1" here, or are you suggesting
% we move the diagram?
% ________________
% David Benn:
% Looks like this has been resolved.
% ________________
% David Benn:
% Not sure whether this is all we want to do or to physically move Fig 1 to here.

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
fix tex syntax so it compiles
add figures to pdf

@@
 
-\begin{document}
+\usepackage{graphicx}
 
@@
 %Affiliation:
-\institution{
+\institute{
       eResearch Group, School of Information Technology and Electronic Engineering,
@@
  \and CSIRO IMT Advanced Scientific Computing and Research Data Services, Melbourne, Australia}
+
+\begin{document}
+
+\maketitle
 
@@
 
-Figure 1. HRPPC component relationships and data flow
+%%Figure 1. HRPPC component relationships and data flow
+\begin{figure}
+\begin{center}
+ \includegraphics[width=12cm,keepaspectratio=true]{plantscan-workflow-figure1.png}
+ % plantscan-workflow-figure1.png: 598x330 pixel, 72dpi, 21.10x11.64 cm, bb=0 0 598 330
+
+\caption{HRPPC component relationships and data flow}
+\label{hrppccomponents}
+\end{center}
+\end{figure}
 
@@
 General scientific properties and phenotype specific properties are then defined
-in larger extension ontologies as illustrated in Figure 2.
+in larger extension ontologies as illustrated in Figure~\ref{poddontologies}.
 
-%%% FIXME: TODO: Insert Figure 2 here
+%Figure 2. 
+\begin{figure}
+\begin{center}
+ \includegraphics[width=12cm,keepaspectratio=true]{podd-ontologies-figure2.png}
+ % podd-ontologies-figure2.png: 5097x2360 pixel, 600dpi, 21.58x9.99 cm, bb=0 0 612 283
 
-Figure 2. PODD schema ontology hierarchy
+\caption{PODD ontology hierarchy}
+\label{poddontologies}
+\end{center}
+\end{figure}
 

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
add bibliography and first reference

@@
 the relationship between components and key data flows.
-
-%% FIXME: TODO: Insert Figure 1 here
 
@@
 be available in a machine readable format. The DOI registrar that DAP uses,
-DataCite [11], was setup to provide unique identifiers for data items that can
+DataCite \cite{Brase2009}, was setup to provide unique identifiers for data items that can
 be attached to publications, which in turn may have their own DOIs.
@@
 both human and machine readable formats.
+
+\bibliographystyle{splncs}
+\bibliography{podd-bibliography}
 

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
add bio2rdf reference

@@
 conclusions. Initially, different providers were not able to decide on URIs to
-use in Linked Data. This resulted in various third parties such as Bio2RDF [12]
+use in Linked Data. This resulted in various third parties such as Bio2RDF \cite{Belleau2008}
 creating RDF versions of scientific datasets so that the RDF documents could be
@@
 [10] https://wiki.csiro.au/display/ASC
-[11] doi:10.1109/COINFO.2009.66 
-[12] doi:10.1016/j.jbi.2008.03.004
+%[11] doi:10.1109/COINFO.2009.66 
+%[12] doi:10.1016/j.jbi.2008.03.004
 [13] doi:10.1007/978-3-642-13654-2\_22

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
add original PODD citation

@@
 In this paper, we expand on the design and implementation of the Phenomics
-Ontology Driven Data repository [13] (PODD) with respect to the capture, storage
+Ontology Driven Data repository \cite{Li2010} (PODD) with respect to the capture, storage
 and retrieval of data and metadata generated at the High Resolution Plant
@@
 %[12] doi:10.1016/j.jbi.2008.03.004
-[13] doi:10.1007/978-3-642-13654-2\_22
+%[13] doi:10.1007/978-3-642-13654-2\_22
 [14] http://www.w3.org/DesignIssues/LinkedData.html

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
convert plantscanTM to latex

@@
 The basic PODD project is structured as a tree, which, when used with
-PlantScanTM, represents the different components of a scientific project as
+PlantScan\textsuperscript{\texttrademark}, represents the different components of a scientific project as
 top-level branches. These include a branch for raw data, along with separate
@@
 
-\section{The PlantScanTM digitisation platform}
+\section{The PlantScan\textsuperscript{\texttrademark} digitisation platform}
 % See unresolved comment at the end
@@
 
-PlantScanTM is a medium throughput high resolution phenotyping platform, which
+PlantScan\textsuperscript{\texttrademark} is a medium throughput high resolution phenotyping platform, which
 brings together a number of imaging sensors (Light detection and ranging, Far
@@
 petiole, stem, leaf) are also captured by PODD (Figure 1). Metadata
-Each acquisition on PlantScanTM includes metadata (in addition to the raw data
+Each acquisition on PlantScan\textsuperscript{\texttrademark} includes metadata (in addition to the raw data
 streams), such as plant genus and species, project and experiment metadata, a
@@
 data, and for the post-processing phases. This information currently resides in
-the PlantScanTM database.
+the PlantScan\textsuperscript{\texttrademark} database.
 
 \subsection{Data volume}
-Digitisation systems such as PlantScanTM generate huge amounts of data including
+Digitisation systems such as PlantScan\textsuperscript{\texttrademark} generate huge amounts of data including
 raw image data, registration metadata, sensor configurations and plant metadata.
-For example, PlantScanTM generates around 500GB of raw image data, representing
+For example, PlantScan\textsuperscript{\texttrademark} generates around 500GB of raw image data, representing
 in excess of 200,000 database records, per day. Sufficient storage space
@@
 Scientific Computing (ASC) storage and compute facilities
-\footnote{https://wiki.csiro.au/display/ASC} are key resources used by PlantScanTM to
+\footnote{https://wiki.csiro.au/display/ASC} are key resources used by PlantScan\textsuperscript{\texttrademark} to
 process and store bulk data.
@@
 \section{CSIRO High Performance Compute Infrastructure}
-CSIRO ASC shared facilities [10] are used to process the raw PlantScanTM data to
+CSIRO ASC shared facilities [10] are used to process the raw PlantScan\textsuperscript{\texttrademark} data to
 derive data products (meshes). Raw data and meshes are ``bagged'' and stored in
@@
 for content integrity verification. Data file related metadata can be stored in
-pre-defined files as key-value pairs. For PlantScanTM, file-level metadata
+pre-defined files as key-value pairs. For PlantScan\textsuperscript{\texttrademark}, file-level metadata
 includes plant barcodes, batch numbers, and plant type.
@@
 \subsection{Bag preparation for a DAP collection}
-CSIRO’s HRPPC makes use of DAP to store collections of PlantScanTM raw images
-and processed mesh data as bags [6]. Raw data from PlantScanTM local storage
+CSIRO’s HRPPC makes use of DAP to store collections of PlantScan\textsuperscript{\texttrademark} raw images
+and processed mesh data as bags [6]. Raw data from PlantScan\textsuperscript{\texttrademark} local storage
 (HRPPC-Store) and data processed on HPC hosts are transferred to ASC bulk
@@
 
-Currently, one bag is equivalent to a single batch scanned on the PlantScanTM
+Currently, one bag is equivalent to a single batch scanned on the PlantScan\textsuperscript{\texttrademark}
 local software system, which usually means the same kind of plant with different
@@
 to the DAP staging area via SFTP (SSH File Transfer Protocol). After publication
-of the DAP collection, the data from PlantScanTM for the given project becomes
+of the DAP collection, the data from PlantScan\textsuperscript{\texttrademark} for the given project becomes
 discoverable via DAP. In addition, experiment reports, published papers, and
@@
 semantic references to shared ontologies where possible, PODD enables the output
-from PlantScanTM to be interpreted and extended by others. The use of PODD URIs
+from PlantScan\textsuperscript{\texttrademark} to be interpreted and extended by others. The use of PODD URIs
 in other RDF documents enables scientists to extend the initial work using the
-Linked Data paradigm [14].
+Linked Data paradigm \cite{BernersLee}.
 
@@
 This paper described how the Phenomics Ontology Driven Data repository
-integrates with the PlantScanTM platform and CSIRO Data Access Portal to manage
+integrates with the PlantScan\textsuperscript{\texttrademark} platform and CSIRO Data Access Portal to manage
 the complex workflows at the High Resolution Plant Phenomics Centre. This
@@
 %[13] doi:10.1007/978-3-642-13654-2\_22
-[14] http://www.w3.org/DesignIssues/LinkedData.html
+%[14] http://www.w3.org/DesignIssues/LinkedData.html
 

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
fine tuning

@@
 
-%Autors: Peter Ansell1, Robert Furbank2, Kutila Gunasekera1, Jianming Guo2, David Benn3, Gareth Williams3, Xavier Sirault2
-\author{Peter Ansell\inst{1}, Robert Furbank\inst{2}, Kutila Gunasekera\inst{1}, Jianming Guo\inst{2}, David Benn\inst{3}, Gareth Williams\inst{3}, Xavier Sirault\inst{2}}
+%Autors: Peter Ansell1, Robert Furbank2, Kutila Gunasekera1, Jianming Guo2,
+%David Benn3, Gareth Williams3, Xavier Sirault2
+\author{Peter Ansell\inst{1}, Robert Furbank\inst{2}, Kutila Gunasekera\inst{1},
+Jianming Guo\inst{2}, David Benn\inst{3}, Gareth Williams\inst{3}, Xavier
+Sirault\inst{2}}
 
@@
 \institute{
-      eResearch Group, School of Information Technology and Electronic Engineering,
+      eResearch Group, School of Information Technology and Electronic
+Engineering,
              University of Queensland, Brisbane, Australia
- \and CSIRO Plant industry, High Resolution Plant Phenomics Centre, Canberra, Australia
- \and CSIRO IMT Advanced Scientific Computing and Research Data Services, Melbourne, Australia}
+ \and CSIRO Plant industry, High Resolution Plant Phenomics Centre, Canberra,
+Australia
+ \and CSIRO IMT Advanced Scientific Computing and Research Data Services,
+Melbourne, Australia}
 
@@
 In this paper, we expand on the design and implementation of the Phenomics
-Ontology Driven Data repository \cite{Li2010} (PODD) with respect to the capture, storage
+Ontology Driven Data repository \cite{Li2010} (PODD) with respect to the
+capture, storage
 and retrieval of data and metadata generated at the High Resolution Plant
@@
 
-%\keywords{eResearch, Semantic Web, RDF, OWL, Data collection citation, BagIt, Data Access Portal}
+%\keywords{eResearch, Semantic Web, RDF, OWL, Data collection citation, BagIt,
+% Data Access Portal}
 
@@
 manages the mapping of collections located in DAP to PODD projects, providing
-for the storage of large images and documents unsuited to RDF databases. Figure 1. shows
+for the storage of large images and documents unsuited to RDF databases. Figure
+1. shows
 the relationship between components and key data flows.
@@
  \includegraphics[width=12cm,keepaspectratio=true]{plantscan-workflow-figure1.png}
- % plantscan-workflow-figure1.png: 598x330 pixel, 72dpi, 21.10x11.64 cm, bb=0 0 598 330
+ % plantscan-workflow-figure1.png: 598x330 pixel, 72dpi, 21.10x11.64 cm, bb=0 0
+598 330
 
@@
 conclusions. Initially, different providers were not able to decide on URIs to
-use in Linked Data. This resulted in various third parties such as Bio2RDF \cite{Belleau2008}
+use in Linked Data. This resulted in various third parties such as Bio2RDF
+\cite{Belleau2008}
 creating RDF versions of scientific datasets so that the RDF documents could be
@@
 The basic PODD project is structured as a tree, which, when used with
-PlantScan\textsuperscript{\texttrademark}, represents the different components of a scientific project as
+PlantScan\textsuperscript{\texttrademark}, represents the different components
+of a scientific project as
 top-level branches. These include a branch for raw data, along with separate
@@
  \includegraphics[width=12cm,keepaspectratio=true]{podd-ontologies-figure2.png}
- % podd-ontologies-figure2.png: 5097x2360 pixel, 600dpi, 21.58x9.99 cm, bb=0 0 612 283
+ % podd-ontologies-figure2.png: 5097x2360 pixel, 600dpi, 21.58x9.99 cm, bb=0 0
+612 283
 
@@
 
-PlantScan\textsuperscript{\texttrademark} is a medium throughput high resolution phenotyping platform, which
-brings together a number of imaging sensors (Light detection and ranging, Far
-infrared imaging, multi-wavelength imaging) to non-invasively measure plant
-growth and function using in-silico approaches [REF?]. Raw data is captured with
+PlantScan\textsuperscript{\texttrademark} is a medium throughput high resolution
+phenotyping platform, which brings together a number of imaging sensors--light 
+detection and ranging, far-infrared imaging, and multi-wavelength imaging--to non-invasively measure plant growth and function using in-silico approaches.
+% FIXME: TODO: [REF?] for PlantScan. 
+Raw data is captured with
 its contextual information (e.g. system configuration, time of acquisition,
@@
 petiole, stem, leaf) are also captured by PODD (Figure 1). Metadata
-Each acquisition on PlantScan\textsuperscript{\texttrademark} includes metadata (in addition to the raw data
+Each acquisition on PlantScan\textsuperscript{\texttrademark} includes metadata
+(in addition to the raw data
 streams), such as plant genus and species, project and experiment metadata, a
@@
 \subsection{Data volume}
-Digitisation systems such as PlantScan\textsuperscript{\texttrademark} generate huge amounts of data including
+Digitisation systems such as PlantScan\textsuperscript{\texttrademark} generate
+huge amounts of data including
 raw image data, registration metadata, sensor configurations and plant metadata.
-For example, PlantScan\textsuperscript{\texttrademark} generates around 500GB of raw image data, representing
+For example, PlantScan\textsuperscript{\texttrademark} generates around 500GB of
+raw image data, representing
 in excess of 200,000 database records, per day. Sufficient storage space
@@
 Scientific Computing (ASC) storage and compute facilities
-\footnote{https://wiki.csiro.au/display/ASC} are key resources used by PlantScan\textsuperscript{\texttrademark} to
+\footnote{https://wiki.csiro.au/display/ASC} are key resources used by
+PlantScan\textsuperscript{\texttrademark} to
 process and store bulk data.
@@
 \section{CSIRO High Performance Compute Infrastructure}
-CSIRO ASC shared facilities [10] are used to process the raw PlantScan\textsuperscript{\texttrademark} data to
+CSIRO ASC shared facilities [10] are used to process the raw
+PlantScan\textsuperscript{\texttrademark} data to
 derive data products (meshes). Raw data and meshes are ``bagged'' and stored in
@@
 for content integrity verification. Data file related metadata can be stored in
-pre-defined files as key-value pairs. For PlantScan\textsuperscript{\texttrademark}, file-level metadata
+pre-defined files as key-value pairs. For
+PlantScan\textsuperscript{\texttrademark}, file-level metadata
 includes plant barcodes, batch numbers, and plant type.
@@
 \subsection{Bag preparation for a DAP collection}
-CSIRO’s HRPPC makes use of DAP to store collections of PlantScan\textsuperscript{\texttrademark} raw images
-and processed mesh data as bags [6]. Raw data from PlantScan\textsuperscript{\texttrademark} local storage
+CSIRO’s HRPPC makes use of DAP to store collections of
+PlantScan\textsuperscript{\texttrademark} raw images
+and processed mesh data as bags [6]. Raw data from
+PlantScan\textsuperscript{\texttrademark} local storage
 (HRPPC-Store) and data processed on HPC hosts are transferred to ASC bulk
@@
 
-Currently, one bag is equivalent to a single batch scanned on the PlantScan\textsuperscript{\texttrademark}
+Currently, one bag is equivalent to a single batch scanned on the
+PlantScan\textsuperscript{\texttrademark}
 local software system, which usually means the same kind of plant with different
@@
 to the DAP staging area via SFTP (SSH File Transfer Protocol). After publication
-of the DAP collection, the data from PlantScan\textsuperscript{\texttrademark} for the given project becomes
+of the DAP collection, the data from PlantScan\textsuperscript{\texttrademark}
+for the given project becomes
 discoverable via DAP. In addition, experiment reports, published papers, and
@@
 be available in a machine readable format. The DOI registrar that DAP uses,
-DataCite \cite{Brase2009}, was setup to provide unique identifiers for data items that can
+DataCite \cite{Brase2009}, was setup to provide unique identifiers for data
+items that can
 be attached to publications, which in turn may have their own DOIs.
@@
 semantic references to shared ontologies where possible, PODD enables the output
-from PlantScan\textsuperscript{\texttrademark} to be interpreted and extended by others. The use of PODD URIs
+from PlantScan\textsuperscript{\texttrademark} to be interpreted and extended by
+others. The use of PODD URIs
 in other RDF documents enables scientists to extend the initial work using the
@@
 This paper described how the Phenomics Ontology Driven Data repository
-integrates with the PlantScan\textsuperscript{\texttrademark} platform and CSIRO Data Access Portal to manage
+integrates with the PlantScan\textsuperscript{\texttrademark} platform and CSIRO
+Data Access Portal to manage
 the complex workflows at the High Resolution Plant Phenomics Centre. This
@@
 % [a]Xavier Sirault:
-% NOTE:Submitted papers should not exceed 15 pages for long papers and 8 pages for
+% NOTE:Submitted papers should not exceed 15 pages for long papers and 8 pages
+for
 % short papers and must be formatted according to the LNCS guidelines. 
@@
 % Proceedings of the workshop will be published online at CEUR Workshop
-% Proceedings (www.CEUR-WS.org ). The 3 best papers presented at the workshop will
+% Proceedings (www.CEUR-WS.org ). The 3 best papers presented at the workshop
+will
 % be included in the supplementary Springer LNCS (Lecture Notes in Computer
@@
 % [c]David Benn:
-% "DAP" has been removed from the abstract. I know it looms large in the paper but
-% is it worth adding something like "and data collection publication"? I've added
+% "DAP" has been removed from the abstract. I know it looms large in the paper
+but
+% is it worth adding something like "and data collection publication"? I've
+added
 % this in but I understand if it's not okay because we're space challenged.
@@
 % David Benn:
-% How about something like: "Plant phenomics imaging, processing, metadata capture
+% How about something like: "Plant phenomics imaging, processing, metadata
+capture
 % and publication". Too verbose I know.
@@
 % David Benn:
-% Do we want to simply change "Figure 4" to "Figure 1" here, or are you suggesting
+% Do we want to simply change "Figure 4" to "Figure 1" here, or are you
+suggesting
 % we move the diagram?
@@
 % David Benn:
-% Not sure whether this is all we want to do or to physically move Fig 1 to here.
+% Not sure whether this is all we want to do or to physically move Fig 1 to
+here.

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
add ASC reference

@@
 the flexibility for linking genomic information through the use of RDF triples.
-In this context, we also describe the role of the CSIRO Data Access Portal (DAP,
-http://data.csiro.au) to annotate and store raw and processed datasets. DAP also
+In this context, we also describe the role of the CSIRO Data Access Portal (DAP) \cite{DAP}
+%http://data.csiro.au) 
+to annotate and store raw and processed datasets. DAP also
 provides long term secure storage for data collections and the ability to search
@@
 manages the mapping of collections located in DAP to PODD projects, providing
-for the storage of large images and documents unsuited to RDF databases. Figure
-1. shows
-the relationship between components and key data flows.
+for the storage of large images and documents unsuited to RDF databases. Figure~\ref{hrppccomponents}
+ shows the relationship between components and key data flows.
 
@@
  % plantscan-workflow-figure1.png: 598x330 pixel, 72dpi, 21.10x11.64 cm, bb=0 0
-598 330
+%598 330
 
@@
  % podd-ontologies-figure2.png: 5097x2360 pixel, 600dpi, 21.58x9.99 cm, bb=0 0
-612 283
+%612 283
 
@@
 computer vision techniques which are used to create the 3D representation of the
-plant (Figure 3)[e]. The 3D meshes are then automatically segmented in order to
+plant.
+% FIXME: TODO: Create Figure 3
+%(Figure 3)[e]. 
+The 3D meshes are then automatically segmented in order to
 semantically identify the different parts of the plants (Paproki et al, 2012). A
@@
 temporal changes at the whole plant and/or organ level. After initial
-processing, i.e. 3D reconstruction, the raw data for each plant is packaged into
+processing to recreate the 3D structure, the raw data for each plant is packaged into
 one location using BagIt, an hierarchical file [6]. All bags belonging to the
@@
 elementary units with permanent addresses which could be retrieved using PODD. 
-The CSIRO Data Access Portal (DAP) \footnote{http://data.csiro.au} and Advanced
-Scientific Computing (ASC) storage and compute facilities
-\footnote{https://wiki.csiro.au/display/ASC} are key resources used by
+The CSIRO Data Access Portal (DAP) \cite{DAP} and Advanced
+Scientific Computing (ASC) storage and compute facilities \cite{ASC} are key resources used by
 PlantScan\textsuperscript{\texttrademark} to
@@
 % NOTE:Submitted papers should not exceed 15 pages for long papers and 8 pages
-for
+%for
 % short papers and must be formatted according to the LNCS guidelines. 
@@
 % Proceedings (www.CEUR-WS.org ). The 3 best papers presented at the workshop
-will
+%will
 % be included in the supplementary Springer LNCS (Lecture Notes in Computer
@@
 % "DAP" has been removed from the abstract. I know it looms large in the paper
-but
+%but
 % is it worth adding something like "and data collection publication"? I've
-added
+%added
 % this in but I understand if it's not okay because we're space challenged.
@@
 % How about something like: "Plant phenomics imaging, processing, metadata
-capture
+%capture
 % and publication". Too verbose I know.
@@
 % Do we want to simply change "Figure 4" to "Figure 1" here, or are you
-suggesting
+%suggesting
 % we move the diagram?
@@
 % Not sure whether this is all we want to do or to physically move Fig 1 to
-here.
+%here.

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
add podd citation that was not converted initially

@@
 for recording the contextual metadata associated with plant genotypes,
-treatments and environmental conditions [13]. PODD manages the links between the
+treatments and environmental conditions \cite{Li2010}. PODD manages the links between the
 raw image files generated by plant imaging platforms, the results from analyses,

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
add gene ontology reference

@@
 their datasets using curated scientific concepts from ontologies. For example,
-scientists have used the Gene Ontology [4] to link well known concepts to
+scientists have used the Gene Ontology \cite{Ashburner2000} to link well known concepts to
 represent common elements across genomics datasets. If these annotations are
@@
 plant phenomics that extends the generally applicable PODD Science ontology with
-plant phenomics specific details [13]. This approach allows PODD to model
+plant phenomics specific details \cite{Li2010}. This approach allows PODD to model
 multiple phenomics processes without enforcing a rigid data structure on
@@
 [3] http://www.plantontology.org/
-[4] http://www.geneontology.org/
+%[4] http://www.geneontology.org/
 [5] https://wiki.csiro.au/confluence/display/dmsdoc/Home

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
more ref cleanup

@@
 \section{CSIRO High Performance Compute Infrastructure}
-CSIRO ASC shared facilities [10] are used to process the raw
+CSIRO ASC shared facilities \cite{ASC} are used to process the raw
 PlantScan\textsuperscript{\texttrademark} data to
@@
 [9] http://sourceforge.net/projects/loc-xferutils/files/loc-bagger/
-[10] https://wiki.csiro.au/display/ASC
+%[10] https://wiki.csiro.au/display/ASC
 %[11] doi:10.1109/COINFO.2009.66 

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
add plant ontology reference
add SPIN/OWL comparison reference

@@
 rules language, it could be easily extended in other cases to use different
-systems such as N3, RDFS, SPARQL, or SPIN as rules languages.
+systems such as N3, RDFS, SPARQL, or SPIN as rules languages \cite{Fuerber2010}.
 
@@
 objects, either now, in the past, or in the future. In addition, other community
-ontologies, such as the plant and gene ontologies [3, 4], can be used to
+ontologies, such as the plant and gene ontologies \cite{Ashburner2000,Avraham2008}, can be used to
 annotate the data. This not only supports data discovery, but provides common
@@
 References
-[1] http://www.iucr.org/\_\_data/iucr/lists/epc-l/pdf00003.pdf
-[2] http://topbraid.org/spin/api/
-[3] http://www.plantontology.org/
+%[1] http://www.iucr.org/\_\_data/iucr/lists/epc-l/pdf00003.pdf
+%[2] http://topbraid.org/spin/api/
+%[3] http://www.plantontology.org/
 %[4] http://www.geneontology.org/

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
add BagIt reference

@@
 processing to recreate the 3D structure, the raw data for each plant is packaged into
-one location using BagIt, an hierarchical file [6]. All bags belonging to the
+one location using BagIt, an hierarchical file \cite{Kunze2011}. All bags belonging to the
 same specific PODD project are grouped into collections on the CSIRO Data Access
@@
 ``hierarchical file packaging format for storage and transfer of arbitrary
-digital content''[6]. A payload manifest details content and MD5 or SHA hashes
+digital content''\cite{Kunze2011}. A payload manifest details content and MD5 or SHA hashes
 for content integrity verification. Data file related metadata can be stored in
@@
 an open source web application that enables research data to be discovered,
-managed and shared. [5]
+managed and shared. \cite{DAP}
 
@@
 PlantScan\textsuperscript{\texttrademark} raw images
-and processed mesh data as bags [6]. Raw data from
+and processed mesh data as bags \cite{Kunze2011}. Raw data from
 PlantScan\textsuperscript{\texttrademark} local storage
@@
 %[4] http://www.geneontology.org/
-[5] https://wiki.csiro.au/confluence/display/dmsdoc/Home
-[6] http://datatracker.ietf.org/doc/draft-kunze-bagit/
+%[5] https://wiki.csiro.au/confluence/display/dmsdoc/Home
+%[6] http://datatracker.ietf.org/doc/draft-kunze-bagit/
 [7] https://pypi.python.org/pypi/bagit/

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
add python bagit reference

@@
 of the multi-core capabilities of modern computing hardware. To suit current
-needs, BagIt-conforming tools [7] [8] [9] were assessed and where necessary,
+needs, BagIt-conforming tools \cite{Summers} [8] [9] were assessed and where necessary,
 improvements were implemented and tested. For example, the implementation of [8]
@@
 with hundreds of processors, this can lead to runaway thread creation, so this
-needed to be addressed. Neither [7] nor [8] permitted payload files in a bag to
+needed to be addressed. Neither \cite{Summers} nor [8] permitted payload files in a bag to
 be listed. This has since been added to [8]. Some functionality (not discussed
-here) present in [8] but not in [7] was added to the latter.
+here) present in [8] but not in \cite{Summers} was added to the latter.
 
@@
 %[6] http://datatracker.ietf.org/doc/draft-kunze-bagit/
-[7] https://pypi.python.org/pypi/bagit/
+%[7] https://pypi.python.org/pypi/bagit/
 [8] http://sourceforge.net/projects/loc-xferutils/

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
add reference for java bagit software from library of congress

@@
 of the multi-core capabilities of modern computing hardware. To suit current
-needs, BagIt-conforming tools \cite{Summers} [8] [9] were assessed and where necessary,
-improvements were implemented and tested. For example, the implementation of [8]
+needs, BagIt-conforming tools \cite{Summers} \cite{LoCBagger} were assessed and where necessary,
+improvements were implemented and tested. For example, the implementation of \cite{LoCBagger}
 assumed that as many threads as possible should be created for multi-threaded
@@
 with hundreds of processors, this can lead to runaway thread creation, so this
-needed to be addressed. Neither \cite{Summers} nor [8] permitted payload files in a bag to
-be listed. This has since been added to [8]. Some functionality (not discussed
-here) present in [8] but not in \cite{Summers} was added to the latter.
+needed to be addressed. Neither \cite{Summers} nor \cite{LoCBagger} permitted payload files in a bag to
+be listed. This has since been added to \cite{LoCBagger}. 
+%Some functionality (not discussed
+%here) present in [8] but not in \cite{Summers} was added to the latter.
 
@@
 
-References
+%References
 %[1] http://www.iucr.org/\_\_data/iucr/lists/epc-l/pdf00003.pdf
@@
 %[7] https://pypi.python.org/pypi/bagit/
-[8] http://sourceforge.net/projects/loc-xferutils/
-[9] http://sourceforge.net/projects/loc-xferutils/files/loc-bagger/
+%[8] http://sourceforge.net/projects/loc-xferutils/
+%[9] http://sourceforge.net/projects/loc-xferutils/files/loc-bagger/
 %[10] https://wiki.csiro.au/display/ASC

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
add citation for original podd to emphasise that it is being redesigned

@@
 
-\subsection{Implementation of the Phenomics Ontology Driven Data repository}
+\subsection{Redesign of the Phenomics Ontology Driven Data repository}
 As described earlier, the PODD repository relies on semantic web technologies to
@@
 for PODD to allow scientists to continue to maintain projects containing curated
-scientific concepts alongside raw experimental data.
+scientific concepts alongside raw experimental data. The PODD repository was 
+redesigned based on an evaluation of the original software \cite{Li2010} that 
+found it was not able to scale sufficiently well to suit the HRPPC needs. The 
+major differences to the old software is that projects are no longer the only 
+supported top object type, and projects are not stored in multiple parts, as 
+that approach was not able to scale as was originally hypothesised.
 
@@
 using a tree based structure, which in the PODD Science ontology is based on the
-concept of a ``Project'' (http://podd.plantphenomics.org.au/podd/). Data
+concept of a ``Project''. 
+%(http://podd.plantphenomics.org.au/podd/). 
+Data
 references are inserted into projects using subproperties of the PODD

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
compress the BagIt descriptions

@@
 The CSIRO Data Access Portal (DAP) \cite{DAP} and Advanced
-Scientific Computing (ASC) storage and compute facilities \cite{ASC} are key resources used by
-PlantScan\textsuperscript{\texttrademark} to
+Scientific Computing (ASC) storage and compute facilities \cite{ASC} are key
+resources used by PlantScan\textsuperscript{\texttrademark} to
 process and store bulk data.
-
-\section{CSIRO High Performance Compute Infrastructure}
-CSIRO ASC shared facilities \cite{ASC} are used to process the raw
-PlantScan\textsuperscript{\texttrademark} data to
-derive data products (meshes). Raw data and meshes are ``bagged'' and stored in
-the ASC archival system. ASC High Performance Compute (HPC) hosts (systems with
-high processor count and large memory) are taken advantage of to create and
-verify bags more rapidly than would be possible on conventional computer
-systems.
-
-\subsection{BagIt}
-
-BagIt is defined by an Internet Engineering Task Force (IETF) document as an
-``hierarchical file packaging format for storage and transfer of arbitrary
-digital content''\cite{Kunze2011}. A payload manifest details content and MD5 or SHA hashes
-for content integrity verification. Data file related metadata can be stored in
-pre-defined files as key-value pairs. For
-PlantScan\textsuperscript{\texttrademark}, file-level metadata
-includes plant barcodes, batch numbers, and plant type.
-
-The BagIt specification does not mandate a particular archiving strategy, the
-focus being upon the directory structure, special files, and integrity checking.
-Bag creation and payload manifest integrity checking, tools can take advantage
-of the multi-core capabilities of modern computing hardware. To suit current
-needs, BagIt-conforming tools \cite{Summers} \cite{LoCBagger} were assessed and where necessary,
-improvements were implemented and tested. For example, the implementation of \cite{LoCBagger}
-assumed that as many threads as possible should be created for multi-threaded
-operations such as bag creation and payload manifest verification. On a HPC host
-with hundreds of processors, this can lead to runaway thread creation, so this
-needed to be addressed. Neither \cite{Summers} nor \cite{LoCBagger} permitted payload files in a bag to
-be listed. This has since been added to \cite{LoCBagger}. 
-%Some functionality (not discussed
-%here) present in [8] but not in \cite{Summers} was added to the latter.
 
@@
 
+\subsection{BagIt}
+
+BagIt is defined by an Internet Engineering Task Force (IETF) document as an
+``hierarchical file packaging format for storage and transfer of arbitrary
+digital content''\cite{Kunze2011}. A payload manifest details content and MD5 or SHA hashes
+for content integrity verification. Data file related metadata can be stored in
+pre-defined files as key-value pairs.
+
+For PlantScan\textsuperscript{\texttrademark}, file-level metadata
+includes plant barcodes, batch numbers, and plant type, although the BagIt
+specification does not mandate a particular archiving strategy, with the
+focus being upon the directory structure, special files, and integrity checking.
+% Bag creation and payload manifest integrity checking, tools can take advantage
+% of the multi-core capabilities of modern computing hardware. 
+To suit current
+needs, BagIt-conforming tools \cite{Summers} \cite{LoCBagger} were assessed and where necessary,
+improvements were implemented and tested to ensure that the tools were scalable with the size of the collections being created. 
+% For example, the implementation of \cite{LoCBagger}
+% assumed that as many threads as possible should be created for multi-threaded
+% operations such as bag creation and payload manifest verification. On a HPC host
+% with hundreds of processors, this can lead to runaway thread creation, so this
+% needed to be addressed. Neither \cite{Summers} nor \cite{LoCBagger} permitted payload files in a bag to
+% be listed. This has since been added to \cite{LoCBagger}. 
+%Some functionality (not discussed
+%here) present in [8] but not in \cite{Summers} was added to the latter.
+
 \subsection{Bag preparation for a DAP collection}
-CSIRO’s HRPPC makes use of DAP to store collections of
+
+CSIRO ASC shared facilities \cite{ASC} are used to process the raw
+PlantScan\textsuperscript{\texttrademark} data to
+derive data products (meshes). Raw data and meshes are collected using the BagIt format \cite{Kunze2011} and stored in
+the ASC archival system. ASC High Performance Compute (HPC) hosts (systems with
+high processor count and large memory) are taken advantage of to create and
+verify bags more rapidly than would be possible on conventional computer
+systems. The DAP to store collections of
 PlantScan\textsuperscript{\texttrademark} raw images
-and processed mesh data as bags \cite{Kunze2011}. Raw data from
+and processed mesh data as bags . Raw data from
 PlantScan\textsuperscript{\texttrademark} local storage

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
reduction of detail about the BagIt to DAP process

@@
 CSIRO ASC shared facilities \cite{ASC} are used to process the raw
-PlantScan\textsuperscript{\texttrademark} data to
-derive data products (meshes). Raw data and meshes are collected using the BagIt format \cite{Kunze2011} and stored in
-the ASC archival system. ASC High Performance Compute (HPC) hosts (systems with
-high processor count and large memory) are taken advantage of to create and
-verify bags more rapidly than would be possible on conventional computer
-systems. The DAP to store collections of
-PlantScan\textsuperscript{\texttrademark} raw images
-and processed mesh data as bags . Raw data from
-PlantScan\textsuperscript{\texttrademark} local storage
-(HRPPC-Store) and data processed on HPC hosts are transferred to ASC bulk
-storage where image and mesh files are organised in folders by batch, then
-barcode number, then subfolders for each image file type (e.g. RGB images, IR
-images, LiDAR) and meshes. Bag creation is carried out via an allocated ASC HPC
-job.
+PlantScan\textsuperscript{\texttrademark} data to derive data products (meshes).
+Raw data and meshes are collected using the BagIt format \cite{Kunze2011} and
+stored in the ASC archival system. ASC High Performance Compute (HPC) hosts
+(systems with high processor count and large memory) are taken advantage of to
+create and verify bags more rapidly than would be possible on conventional
+computer systems. The DAP to store collections of
+PlantScan\textsuperscript{\texttrademark} 
+raw images and processed mesh data as bags . Raw data from
+PlantScan\textsuperscript{\texttrademark} 
+local storage (HRPPC-Store) and data processed on HPC hosts are transferred to
+ASC bulk storage where image and mesh files are organised in folders by batch,
+then barcode number, then subfolders for each image file type (e.g. RGB images,
+IR images, LiDAR) and meshes. Bag creation is carried out via an allocated ASC
+HPC job.
 
@@
 local software system, which usually means the same kind of plant with different
-genotypes scanned under one  experiment configuration profile.  However the
-naming, content and amount of data in a bag are flexible and depend upon the
-intention of the bag. For example, multiple batches can be packed into one bag
-when the batches come from one plant’s scans at different time points. The
-multi-threaded BagIt tool is called on the ASC HPC host to create a bag that is
-then compressed into a single file containing a top-level data folder, a
-manifest file that contains checksums for every data file, and tag files that
-contain the metadata description of the bag just packed.
-
+genotypes scanned under one  experiment configuration profile.  
+% However the
+% naming, content and amount of data in a bag are flexible and depend upon the
+% intention of the bag. For example, multiple batches can be packed into one bag
+% when the batches come from one plant’s scans at different time points. The
+% multi-threaded BagIt tool is called on the ASC HPC host to create a bag that is
+% then compressed into a single file containing a top-level data folder, a
+% manifest file that contains checksums for every data file, and tag files that
+% contain the metadata description of the bag just packed. 
 The metadata required for a DAP publication is created and the bag transferred

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
rearrange to separate the plant phenomics details from the technical details
reduce details about PODD

@@
 PlantScan\textsuperscript{\texttrademark} 
-raw images and processed mesh data as bags . Raw data from
+raw images and processed mesh data as bags. Currently, one bag is equivalent to
+a single batch scanned on the
+PlantScan\textsuperscript{\texttrademark}
+local software system, which usually means the same kind of plant with different
+genotypes scanned under one  experiment configuration profile.
+
+Raw data from
 PlantScan\textsuperscript{\texttrademark} 
@@
 HPC job.
-
-Currently, one bag is equivalent to a single batch scanned on the
-PlantScan\textsuperscript{\texttrademark}
-local software system, which usually means the same kind of plant with different
-genotypes scanned under one  experiment configuration profile.  
 % However the
@@
 
-The PODD ontology allows sparse annotation of data and results, so scientists
-are not required to annotate all of their data with links to shared ontologies
-by default. However, the system is extensible, so in future such a requirement
-could be added in a new version of the PODD ontology while maintaining backwards
-compatibility with previous experiments that may be published or immutable.
+% The PODD ontology allows sparse annotation of data and results, so scientists
+% are not required to annotate all of their data with links to shared ontologies
+% by default. However, the system is extensible, so in future such a requirement
+% could be added in a new version of the PODD ontology while maintaining backwards
+% compatibility with previous experiments that may be published or immutable.
 
@@
 
-PODD implements support for Content Negotiation, with common RDF serialisations
-supported in addition to HTML pages that are encoded with RDFa (RDF in
-Attributes). Although most non-human agents will receive an RDF formatted
-document when they resolve a PODD URI, if a PODD HTML page is saved the
-annotations can still be retrieved using RDFa.
+% PODD implements support for Content Negotiation, with common RDF serialisations
+% supported in addition to HTML pages that are encoded with RDFa (RDF in
+% Attributes). Although most non-human agents will receive an RDF formatted
+% document when they resolve a PODD URI, if a PODD HTML page is saved the
+% annotations can still be retrieved using RDFa.
 

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
more cutbacks to reduce length

@@
 scientists have used the Gene Ontology \cite{Ashburner2000} to link well known concepts to
-represent common elements across genomics datasets. If these annotations are
+represent common elements across genomics datasets, while the Plant Ontology \cite{Avraham2008} 
+allows the description of plant based datasets. If these annotations are
 given using recognised URIs, and not as key-value string tuples, various RDF
@@
 \subsection{Redesign of the Phenomics Ontology Driven Data repository}
-As described earlier, the PODD repository relies on semantic web technologies to
+The PODD repository relies on semantic web technologies to
 manage phenomics data and metadata. Although both ontologies and mappings are
@@
 
-An ontology is a common controlled vocabulary about a particular topic, in this
-case plant phenomics. The PODD Plant ontology therefore becomes an ontology for
-plant phenomics that extends the generally applicable PODD Science ontology with
-plant phenomics specific details \cite{Li2010}. This approach allows PODD to model
-multiple phenomics processes without enforcing a rigid data structure on
-objects, either now, in the past, or in the future. In addition, other community
-ontologies, such as the plant and gene ontologies \cite{Ashburner2000,Avraham2008}, can be used to
-annotate the data. This not only supports data discovery, but provides common
-reference points through which PODD managed data can be integrated with data in
-other research databases.
+% An ontology is a common controlled vocabulary about a particular topic, in this
+% case plant phenomics. The PODD Plant ontology therefore becomes an ontology for
+% plant phenomics that extends the generally applicable PODD Science ontology with
+% plant phenomics specific details \cite{Li2010}. This approach allows PODD to model
+% multiple phenomics processes without enforcing a rigid data structure on
+% objects, either now, in the past, or in the future. In addition, other community
+% ontologies, such as the plant and gene ontologies \cite{Ashburner2000,Avraham2008}, can be used to
+% annotate the data. This not only supports data discovery, but provides common
+% reference points through which PODD managed data can be integrated with data in
+% other research databases.
 
-The PODD application handles the instantiation and validation of objects based
-on the versions of ontologies that are specified by the user as being relevant.
-This gives PODD a unique extensibility as the domain knowledge is entirely
-captured in the ontology, and the operational aspects of the system are derived
-purely from the content of the ontology. In particular, PODD is designed around
-the ability to support multiple versions of ontologies concurrently. Previously
-published artifacts can be loaded without modification into PODD without
-affecting current studies that may be using newer, incompatible, versions of the
-schema ontologies.
+% The PODD application handles the instantiation and validation of objects based
+% on the versions of ontologies that are specified by the user as being relevant.
+% This gives PODD a unique extensibility as the domain knowledge is entirely
+% captured in the ontology, and the operational aspects of the system are derived
+% purely from the content of the ontology. In particular, PODD is designed around
+% the ability to support multiple versions of ontologies concurrently. Previously
+% published artifacts can be loaded without modification into PODD without
+% affecting current studies that may be using newer, incompatible, versions of the
+% schema ontologies.
 
-PODD is able to manage high volume, high resolution, heterogeneous datasets
-using a tree based structure, which in the PODD Science ontology is based on the
-concept of a ``Project''. 
-%(http://podd.plantphenomics.org.au/podd/). 
-Data
-references are inserted into projects using subproperties of the PODD
-``contains'' predicate. When results are available they are inserted
-independently and cross-referenced with the data that they were derived from
-using a subproperty of the PODD ``refers to'' property. Each of the
-subproperties have constraints on the types of objects that are valid as sources
-and destinations, making it easy to both validate existing uses, and suggest
-potential targets based on a source and a property.
+% PODD is able to manage high volume, high resolution, heterogeneous datasets
+% using a tree based structure, which in the PODD Science ontology is based on the
+% concept of a ``Project''. 
+% %(http://podd.plantphenomics.org.au/podd/). 
+% Data
+% references are inserted into projects using subproperties of the PODD
+% ``contains'' predicate. When results are available they are inserted
+% independently and cross-referenced with the data that they were derived from
+% using a subproperty of the PODD ``refers to'' property. Each of the
+% subproperties have constraints on the types of objects that are valid as sources
+% and destinations, making it easy to both validate existing uses, and suggest
+% potential targets based on a source and a property.
 
-The genotype of the plant alone does not determine the phenotype; it is the
-combination of the organism’s genome with the environmental conditions that
-determine the phenotype. PODD captures metadata on the genotype of the
-organism(s) under investigation. PODD also gathers comprehensive metadata on the
-environmental conditions and treatments that the biological samples are exposed
-to. This structure makes PODD suitable to queries by web agents thus  linking
-information contained in other data repositories, e.g genomics, transcriptomics,
-metabolomics to phenomics information.
-
-\section{The PlantScan\textsuperscript{\texttrademark} digitisation platform}
-% See unresolved comment at the end
-
-%System description and Processes
-
-\subsection{Heterogeneous data streams}
-
-PlantScan\textsuperscript{\texttrademark} is a medium throughput high resolution
-phenotyping platform, which brings together a number of imaging sensors--light 
-detection and ranging, far-infrared imaging, and multi-wavelength imaging--to non-invasively measure plant growth and function using in-silico approaches.
-% FIXME: TODO: [REF?] for PlantScan. 
-Raw data is captured with
-its contextual information (e.g. system configuration, time of acquisition,
-batch number and project) and is stored in a purpose-built database as the data
-is being generated. The various data streams are collated and used to produce
-full 3D representation of each plant with overlaid spectral information. The
-metadata collected during image acquisition are necessary inputs for the
-computer vision techniques which are used to create the 3D representation of the
-plant.
-% FIXME: TODO: Create Figure 3
-%(Figure 3)[e]. 
-The 3D meshes are then automatically segmented in order to
-semantically identify the different parts of the plants (Paproki et al, 2012). A
-longitudinal 3D matching pipeline for plant mesh parts is then used to evaluate 
-temporal changes at the whole plant and/or organ level. After initial
-processing to recreate the 3D structure, the raw data for each plant is packaged into
-one location using BagIt, an hierarchical file \cite{Kunze2011}. All bags belonging to the
-same specific PODD project are grouped into collections on the CSIRO Data Access
-Portal. These bags have a permanent address which is recorded into PODD. Any
-additional processes on these digital objects (such as statistical information
-on features like plant colour, volume,  growth etc. and on elements such as
-petiole, stem, leaf) are also captured by PODD (Figure 1). Metadata
-Each acquisition on PlantScan\textsuperscript{\texttrademark} includes metadata
-(in addition to the raw data
-streams), such as plant genus and species, project and experiment metadata, a
-unique identifier for each image (Globally Unique Identifier), imaging angle,
-environmental temperature of the imaging chamber, location of optical and colour
-calibration datasets for each acquisition run, LiDAR (Light Detection and
-Ranging Sensors) calibration files. The metadata associated with each
-acquisition is automatically generated when setting up the configuration on the
-platform. This information is paramount to validate and process the raw image
-data, and for the post-processing phases. This information currently resides in
-the PlantScan\textsuperscript{\texttrademark} database.
-
-\subsection{Data volume}
-Digitisation systems such as PlantScan\textsuperscript{\texttrademark} generate
-huge amounts of data including
-raw image data, registration metadata, sensor configurations and plant metadata.
-For example, PlantScan\textsuperscript{\texttrademark} generates around 500GB of
-raw image data, representing
-in excess of 200,000 database records, per day. Sufficient storage space
-(usually at remote locations) and fast network transfer rates were thus
-necessary to facilitate data movement for processing using high performance
-computers (HPC). Because an RDF database structure is not suitable for handling
-large data sets of images, it was necessary to package the raw information into
-elementary units with permanent addresses which could be retrieved using PODD. 
-The CSIRO Data Access Portal (DAP) \cite{DAP} and Advanced
-Scientific Computing (ASC) storage and compute facilities \cite{ASC} are key
-resources used by PlantScan\textsuperscript{\texttrademark} to
-process and store bulk data.
+% The genotype of the plant alone does not determine the phenotype; it is the
+% combination of the organism’s genome with the environmental conditions that
+% determine the phenotype. PODD captures metadata on the genotype of the
+% organism(s) under investigation. PODD also gathers comprehensive metadata on the
+% environmental conditions and treatments that the biological samples are exposed
+% to. This structure makes PODD suitable to queries by web agents thus  linking
+% information contained in other data repositories, e.g genomics, transcriptomics,
+% metabolomics to phenomics information.
 
@@
 
+\section{The PlantScan\textsuperscript{\texttrademark} digitisation platform}
+% See unresolved comment at the end
+
+%System description and Processes
+
 \subsection{BagIt}
@@
 % of the multi-core capabilities of modern computing hardware. 
-To suit current
-needs, BagIt-conforming tools \cite{Summers} \cite{LoCBagger} were assessed and where necessary,
+% To suit current
+% needs, 
+BagIt-conforming tools \cite{Summers} \cite{LoCBagger} were assessed and where necessary,
 improvements were implemented and tested to ensure that the tools were scalable with the size of the collections being created. 
@@
 data (e.g. bag).
+
+\subsection{Heterogeneous data streams}
+
+PlantScan\textsuperscript{\texttrademark} is a medium throughput high resolution
+phenotyping platform, which brings together a number of imaging sensors--light 
+detection and ranging, far-infrared imaging, and multi-wavelength imaging--to non-invasively measure plant growth and function using in-silico approaches.
+% FIXME: TODO: [REF?] for PlantScan. 
+Raw data is captured with
+its contextual information (e.g. system configuration, time of acquisition,
+batch number and project) and is stored in a purpose-built database as the data
+is being generated. The various data streams are collated and used to produce
+full 3D representation of each plant with overlaid spectral information. The
+metadata collected during image acquisition are necessary inputs for the
+computer vision techniques which are used to create the 3D representation of the
+plant.
+% FIXME: TODO: Create Figure 3
+%(Figure 3)[e]. 
+The 3D meshes are then automatically segmented in order to
+semantically identify the different parts of the plants (Paproki et al, 2012). A
+longitudinal 3D matching pipeline for plant mesh parts is then used to evaluate 
+temporal changes at the whole plant and/or organ level.
+
+% After initial processing to recreate the 3D structure, the raw data for each plant is packaged into
+% one location using BagIt. All bags belonging to the
+% same specific PODD project are grouped into collections on the CSIRO Data Access
+% Portal. These bags have a permanent address which is recorded into PODD. Any
+% additional processes on these digital objects (such as statistical information
+% on features like plant colour, volume,  growth etc. and on elements such as
+% petiole, stem, leaf) are also captured by PODD using the properties in the 
+% PODD Plant Ontology (Figure~\ref{poddontologies}).
+
+\subsection{Metadata}
+Each acquisition on PlantScan\textsuperscript{\texttrademark} includes metadata
+(in addition to the raw data
+streams), such as plant genus and species, project and experiment metadata, a
+unique identifier for each image (Globally Unique Identifier), imaging angle,
+environmental temperature of the imaging chamber, location of optical and colour
+calibration datasets for each acquisition run, LiDAR (Light Detection and
+Ranging Sensors) calibration files. The metadata associated with each
+acquisition is automatically generated when setting up the configuration on the
+platform. This information is paramount to validate and process the raw image
+data, and for the post-processing phases. This information currently resides in
+the PlantScan\textsuperscript{\texttrademark} database.
+
+\subsection{Data volume}
+Digitisation systems such as PlantScan\textsuperscript{\texttrademark} generate
+huge amounts of data including
+raw image data, registration metadata, sensor configurations and plant metadata.
+For example, PlantScan\textsuperscript{\texttrademark} generates around 500GB of
+raw image data, representing
+in excess of 200,000 database records, per day. Sufficient storage space
+(usually at remote locations) and fast network transfer rates were thus
+necessary to facilitate data movement for processing using high performance
+computers (HPC). Because an RDF database structure is not suitable for handling
+large data sets of images, it was necessary to package the raw information into
+elementary units with permanent addresses which could be retrieved using PODD. 
+The CSIRO Data Access Portal (DAP) \cite{DAP} and Advanced
+Scientific Computing (ASC) storage and compute facilities \cite{ASC} are key
+resources used by PlantScan\textsuperscript{\texttrademark} to
+process and store bulk data.
 

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
more cutbacks

@@
 represent common elements across genomics datasets, while the Plant Ontology \cite{Avraham2008} 
-allows the description of plant based datasets. If these annotations are
-given using recognised URIs, and not as key-value string tuples, various RDF
-documents can be merged automatically based on RDF semantics to discover novel
-conclusions. Initially, different providers were not able to decide on URIs to
-use in Linked Data. This resulted in various third parties such as Bio2RDF
-\cite{Belleau2008}
-creating RDF versions of scientific datasets so that the RDF documents could be
-easily merged based on common URIs. However, recently more providers have given
-authoritative URIs for each object in their dataset, making it easy to integrate
-data from various locations.
+allows the description of plant based datasets. 
+% If these annotations are
+% given using recognised URIs, and not as key-value string tuples, various RDF
+% documents can be merged automatically based on RDF semantics to discover novel
+% conclusions. 
+% Initially, different providers were not able to decide on URIs to
+% use in Linked Data. This resulted in various third parties such as Bio2RDF
+% \cite{Belleau2008}
+% creating RDF versions of scientific datasets so that the RDF documents could be
+% easily merged based on common URIs. However, recently more providers have given
+% authoritative URIs for each object in their dataset, making it easy to integrate
+% data from various locations.
 
@@
 found it was not able to scale sufficiently well to suit the HRPPC needs. The 
-major differences to the old software is that projects are no longer the only 
-supported top object type, and projects are not stored in multiple parts, as 
-that approach was not able to scale as was originally hypothesised.
+major differences to the software implemented by \cite{Li2010} is that projects 
+are no longer the only supported top object type, and projects are not stored 
+in multiple parts, as that approach was not able to scale as was originally hypothesised.
 
@@
 case of raw data, the semantics are not clear and cannot be defined by the
-platforms collecting the data. Scientists are able to sparsely attach meaning
-using horizontal links between different branches, using the OWL Open World
-Assumption to allow extra semantic annotations to be attached in future. For
-example, they may annotate images of a plant with a trait in only one case,
+platforms collecting the data. 
+% Scientists are able to sparsely attach meaning
+% using horizontal links between different branches, using the OWL Open World
+% Assumption to allow extra semantic annotations to be attached in future. 
+For example, they may annotate images of a plant with a trait in only one case,
 without annotating other images. These horizontal links are not constrained by
@@
 the defined properties, without requiring all of the properties to be used.
-
-The parent object types and properties used by PODD to establish basic ground
-rules for rendering and validation are defined in a small base schema ontology.
-General scientific properties and phenotype specific properties are then defined
+% The parent object types and properties used by PODD to establish basic ground
+% rules for rendering and validation are defined in a small base schema ontology.
+General scientific properties and phenotype specific properties are defined
 in larger extension ontologies as illustrated in Figure~\ref{poddontologies}.

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
cut back to 8 pages by reducing details in the introduction

@@
 human being, i.e. new knowledge. The details about the steps taken are recorded
-to give context to the data. These links enable scientists to generate novel
-conclusions about the plants being phenotyped, including conclusions based on
-the genotype, plant growth environments, treatments, phenotyping platforms and
-the phenotyping and analysis processes.
+to give context to the data. 
+% These links enable scientists to generate novel
+% conclusions about the plants being phenotyped, including conclusions based on
+% the genotype, plant growth environments, treatments, phenotyping platforms and
+% the phenotyping and analysis processes.
 
@@
 for recording the contextual metadata associated with plant genotypes,
-treatments and environmental conditions \cite{Li2010}. PODD manages the links between the
-raw image files generated by plant imaging platforms, the results from analyses,
-and conclusions by scientists.
+treatments and environmental conditions \cite{Li2010}. 
+% PODD manages the links between the
+% raw image files generated by plant imaging platforms, the results from analyses,
+% and conclusions by scientists.
 

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
fix minor typo - remove an 's'

@@
 all objects having an explicit RDF type, and whether they are consistent with
-the ontologies that they import. The main points that PODD focuses on for
+the ontologies that they import. The main point that PODD focuses on for
 validation is determining that mappings are consistent with the basic PODD

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
fix sentence to be meaningful
keywords added in again

@@
 
-%\keywords{eResearch, Semantic Web, RDF, OWL, Data collection citation, BagIt,
-% Data Access Portal}
+\keywords{eResearch, Semantic Web, RDF, OWL, Data collection citation, BagIt, Data Access Portal}
 
@@
 create and verify bags more rapidly than would be possible on conventional
-computer systems. The DAP to store collections of
+computer systems. DAP is used to store collections of
 PlantScan\textsuperscript{\texttrademark} 

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
apply changed based on David's review

@@
 Researchers can describe a data collection, deposit data, choose a license, and
-add attribution details. Access to a collection’s description and/or data can be
+add attribution details. Access to a collection's description and/or data can be
 restricted to CSIRO or a set of individuals (within CSIRO or partner
@@
 BagIt-conforming tools \cite{Summers} \cite{LoCBagger} were assessed and where necessary,
-improvements were implemented and tested to ensure that the tools were scalable with the size of the collections being created. 
+improvements were implemented and tested to ensure that the tools were fit for purpose in the ASC HPC environment.
 % For example, the implementation of \cite{LoCBagger}
@@
 create and verify bags more rapidly than would be possible on conventional
-computer systems. DAP is used to store collections of
-PlantScan\textsuperscript{\texttrademark} 
-raw images and processed mesh data as bags. Currently, one bag is equivalent to
+computer systems. 
+CSIRO's HRPPC makes use of DAP to store collections of PlantScan\textsuperscript{\texttrademark} raw images and processed mesh data as bags.
+%DAP is used to store collections of
+%PlantScan\textsuperscript{\texttrademark} 
+%raw images and processed mesh data as bags. 
+Currently, one bag is equivalent to
 a single batch scanned on the

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
use acronyms that have been previously defined to reduce space

@@
 elementary units with permanent addresses which could be retrieved using PODD. 
-The CSIRO Data Access Portal (DAP) \cite{DAP} and Advanced
-Scientific Computing (ASC) storage and compute facilities \cite{ASC} are key
+The CSIRO DAP \cite{DAP} and ASC storage and compute facilities \cite{ASC} are key
 resources used by PlantScan\textsuperscript{\texttrademark} to

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
comment out reference to PlantScan database
more fine tuning

@@
 Australia
- \and CSIRO IMT Advanced Scientific Computing and Research Data Services,
+ \and CSIRO IM\&T Advanced Scientific Computing and Research Data Services,
 Melbourne, Australia}
@@
 platforms located in purpose built phenomics centres across the world. These
-large datasets of phenotypic characters such as growth rate, plant architecture,
-photosynthetic performance, yield must be stored and related to genotype. This
-link is essential for expediting the identification of genetic variation in
-natural and derived genetic populations (e.g. germplasm collections, association
-genetic panels, recombinant inbred lines), understanding the dynamic
-relationship between phenotype, genotype and environment and delivering the
-increase in productivity necessary for feeding the world.
+large datasets of phenotypic characteristics such as growth rate, plant architecture,
+photosynthetic performance, yield must be stored and correlated with genotypes.
+These factors provide evidence of genetic variation in natural and derived genetic populations (e.g.
+germplasm collections, association genetic panels, recombinant inbred lines).
+They also enable a deeper understanding the dynamic relationship between phenotype, genotype and environment which is necessary to continue delivering the increase in productivity necessary for feeding the world.
 
-To achieve this goal, the vast array of phenotypic data collected from a variety
+The vast array of phenotypic data collected from a variety
 of phenomics platforms must be combined with metadata explaining how the raw
-information was collected. This process forms structured digital objects
-containing the raw information. These objects are then delivered to a range of
-analysis pipelines, which transform the raw information into  aggregated
+data was collected. This combination of raw data and metadata are then delivered to a range of analysis pipelines, which transform the raw data into aggregated
 multi-phase datasets, each phase representing a new aggregation or inference
@@
 multi-dimensional data into information which is conceptually interpretable by a
-human being, i.e. new knowledge. The details about the steps taken are recorded
-to give context to the data. 
+human being, i.e. new knowledge. The additional metadata describing the steps taken are recorded to give context to the data. 
 % These links enable scientists to generate novel
@@
 redesigned based on an evaluation of the original software \cite{Li2010} that 
-found it was not able to scale sufficiently well to suit the HRPPC needs. The 
-major differences to the software implemented by \cite{Li2010} is that projects 
+found it was not able to scale sufficiently to suit the HRPPC needs. The 
+major differences to the software implemented by \cite{Li2010} are that projects 
 are no longer the only supported top object type, and projects are not stored 
@@
 platform. This information is paramount to validate and process the raw image
-data, and for the post-processing phases. This information currently resides in
-the PlantScan\textsuperscript{\texttrademark} database.
+data, and for the post-processing phases. 
+% This information currently resides in
+% the PlantScan\textsuperscript{\texttrademark} database.
 

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
cut sentences to get back to 8 pages
final draft for submission

@@
 Object Identifier (DOI) is issued and can be used to formally cite the
-collection in a publication. Collections can either be one-time, self-serve
-deposits or automated, ongoing deposits and can be changed over time through
-versioning. Whole datasets or specific files within a collection can be
-downloaded. A machine-to-machine REST interface permits programmatic search and
-collection download.
+collection in a publication.
+
+%Collections can either be one-time, self-serve
+%deposits or automated, ongoing deposits and can be changed over time through
+%versioning. Whole datasets or specific files within a collection can be
+%downloaded. A machine-to-machine REST interface permits programmatic search and
+%collection download.
 
@@
 BagIt-conforming tools \cite{Summers} \cite{LoCBagger} were assessed and where necessary,
-improvements were implemented and tested to ensure that the tools were fit for purpose in the ASC HPC environment.
+improvements were implemented and tested to ensure that the tools were fit for purpose in the CSIRO Advanced Scientific Computing (ASC) HPC environment.
 % For example, the implementation of \cite{LoCBagger}
@@
 ASC bulk storage where image and mesh files are organised in folders by batch,
-then barcode number, then subfolders for each image file type (e.g. RGB images,
-IR images, LiDAR) and meshes. Bag creation is carried out via an allocated ASC
-HPC job.
+then barcode number, then subfolders for each image file type, including RGB images,
+IR images, and LiDAR (Light Detection and
+Ranging Sensors, and their related meshes. Bag creation is carried out via an allocated ASC HPC job.
 % However the
@@
 environmental temperature of the imaging chamber, location of optical and colour
-calibration datasets for each acquisition run, LiDAR (Light Detection and
-Ranging Sensors) calibration files. The metadata associated with each
+calibration datasets for each acquisition run, and LiDAR calibration files. The metadata associated with each
 acquisition is automatically generated when setting up the configuration on the
@@
 in excess of 200,000 database records, per day. Sufficient storage space
-(usually at remote locations) and fast network transfer rates were thus
+(usually at remote locations) and fast network transfer rates are thus
 necessary to facilitate data movement for processing using high performance
 computers (HPC). Because an RDF database structure is not suitable for handling
-large data sets of images, it was necessary to package the raw information into
+large data sets of images, it is necessary to package the raw information into
 elementary units with permanent addresses which could be retrieved using PODD. 

### 2013-03-11 s4biodiv2013/PODD-S4BIODIV-Submission.tex
add paproki reference to bibtex and cleanup so it goes back to 8 pages.

@@
 The 3D meshes are then automatically segmented in order to
-semantically identify the different parts of the plants (Paproki et al, 2012). A
+semantically identify the different parts of the plants \cite{Paproki2012}. A
 longitudinal 3D matching pipeline for plant mesh parts is then used to evaluate 
@@
 editing. PODD also supports an RDF-based HTTP REST interface that clients may
-use to fetch and update objects without using the HTML pages.
+use to fetch and update objects.
+% without using the HTML pages.
 

### 2013-04-12 s4biodiv2013/Rebuttals.txt
add initial Rebuttals

 
----------------------- REVIEW 1 ---------------------
PAPER: 1
TITLE: Flexible Scientific Data Management for Plant Phenomics Research
AUTHORS: Peter Ansell, Robert Furbank, Kutila Gunasekera, Jianming Guo, David Benn, Gareth Williams and Xavier Sirault

OVERALL EVALUATION: 3 (strong accept)
REVIEWER'S CONFIDENCE: 5 (expert)
What is your expertise level for the paper?: 5 (excellent) For this paper do you suggest a long presentation (20 min) ?: 1 (yes) For this paper do you suggest a flashtalk (5 min) ?: 3 (maybe) Is this paper follows LNCS guideline?: 1 (yes) Would you consider the paper to submit an extended paper for publication in a journal?: 3 (maybe) Do you think the paper will open or drive a fruitful discussion during the workshop?: 1 (yes) APPROPRIATENESS for the workshop : from 1 (lowest) to 5 (highest): 5 (excellent)
ORIGINALITY: from 1 (lowest) to 5 (highest): 5 (excellent) TECHNICAL CORRECTNESS: from 1 (lowest) to 5 (highest): 5 (excellent)

----------- REVIEW -----------
This paper demonstrate several inovative application of ontologies in Phenomic like : (i) ontology driven data store, (ii) data consistency check, (iii) relation between multiple references ontologies.

For more than ten years, storage and archiving biological data was achieved by first designing a data model representing what was to be stored and then implementing it, generaly in a database. In this paper, we find those two steps but with two major improvment. First, being based on an ontology, the data model is richer that what can be commonly achieved with traditional UML/Database modeling. Secondly, the implementation is not purely based on a database but relies on semantic web technologies like OWL, and are thus very likely to fully preserve the semantic richness of the model.
Furthermore, this datastore is designed for Big Data, and it is of very high interesst to see how ontologies scales and what solutions can be implemented for high volumes of data.

The ontologies are also used for a two level consistency check. There is a core set of rule to ensure the integrability of the data in the system, but it seems to allow the neccessary level of flexibility to ensure data quality without restricting data submission and dataset elaboration.
Furthermore, PODD allows to use several reference ontologies thus easing data interoperability.

This paper shows extensive use of semantic web technologies applied to an important biological field, Phenomics applied to the study of phenotype, genotype and environment relationship in plants. For me, it is therefore perfectly in the scope of S4BioDiv.

It is a very original work, as there no or little phenomic datastore curently relying on ontologies like PODD. Furthermore, it is based on recognized technologies like OWL, RDF and implemented with recognized frameworks like BagIt (cited in the paper) or Hibernate ORM (used in source code). This indicate an excelent technical correcteness.

The experience of PODD is very interresting for every group managing an information system, and this talk should lead to long and fruitfull discussions.

----------------------- REVIEW 2 ---------------------
PAPER: 1
TITLE: Flexible Scientific Data Management for Plant Phenomics Research
AUTHORS: Peter Ansell, Robert Furbank, Kutila Gunasekera, Jianming Guo, David Benn, Gareth Williams and Xavier Sirault

OVERALL EVALUATION: -1 (weak reject)
REVIEWER'S CONFIDENCE: 3 (medium)
What is your expertise level for the paper?: 2 (poor) For this paper do you suggest a long presentation (20 min) ?: 2 (no) For this paper do you suggest a flashtalk (5 min) ?: 2 (no) Is this paper follows LNCS guideline?: 1 (yes) Would you consider the paper to submit an extended paper for publication in a journal?: 2 (no) Do you think the paper will open or drive a fruitful discussion during the workshop?: 2 (no) APPROPRIATENESS for the workshop : from 1 (lowest) to 5 (highest): 1 (very poor)
ORIGINALITY: from 1 (lowest) to 5 (highest): 4 (good) TECHNICAL CORRECTNESS: from 1 (lowest) to 5 (highest): 3 (fair)

----------- REVIEW -----------
The authors describe PODD - an ontology-driven repository to contain data coming from a high-throughput plant phenotyping information.

Though I have few criticisms of the paper itself (and I will raise those in a moment), my main concern is that the authors don't describe how this system is "for biodiversity", and moreover, there is also only passing reference to the semantics of their system in the manuscript.

*****

Rebuttal: We believe that a high level relationship between the semantics of the system and the plant imaging platforms has been described in the paper. We did not set out to describe in intimate details all of the classes and properties that we used to perform this integration in this short workshop paper.

*****

The majority of the manuscript was spent describing neither semantics, nor biodiversity.

*****

Rebuttal: While the majority of the paper describes technical aspects of the system, we believe that its contains direct relationships to both semantics and plant phenomics, the latter of which is a subset of biodiversity.

*****

As such, while the manuscript itself is, in principle, high-quality, it would need to be significantly re-written to be appropriate for this venue.  In an attempt to give the manuscript a more fair review, I attempted to find the connection for myself.  A Google search for HRPPC + biodiversity found several mentions of how the output from the center might be used for biodiversity, but these were extremely generic statements and there didn't appear to be any active project in biodiversity (and if there is, the authors need to describe this in the manuscript).

*****

Rebuttal: The references contain a publication written by HRPPC describing the links between the mesh processing algorithms which generate data which is managed by the PODD system and plant phenomics diversity, which is a subset of biodiversity.

*****

The use of semantics, as described, is difficult to !
 follow.  I understand that they use concepts from, e.g. the Plant Ontology, to annotate their datasets, but that in itself is fairly common and isn't really a publishable achievement.  There is a difficult-to-understand description of the use of OWL as a "rules language", but no clear explanation of why one would use a Description Logic in this way, as opposed to using... for example... a Rules Language.

 *****

 Rebuttal: OWL has been previously used for constraint verification, aka, "Rules", and its approach may be superior to other possible languages. Searching Google for "OWL constraint verifications" returns a range of publications describing its use in this way.

 *****

 As such, the "semantic" aspect of the manuscript isn't particularly compelling either.  The authors didn't discuss if/how semantics is used in their capture and representation of metadata.

 *****

 Rebuttal: Section 2.2 and 2.3 have been updated to include clarifications of the relationship between the (metadata) semantics of the project structure, and the (science) semantics of the links between data, descriptions of processes, results, and conclusions. The project structure semantics are heavily constrained, while the science semantics are less strictly constrained. The science semantics do not attempt to generate high quality upper ontologies, rather they attempt to represent the scientists intentions in a slightly more structured way than otherwise would be possible.

 *****

 The final aspect where semantics was discussed was with respect to "semantic publishing", and that section of the manuscript also wasn't clear.  Presumably by "semantic publishing" the authors mean that it is possible to retrieve fragments of RDF that describe the datasets and their ontology-grounded annotations, but again, this (in and of itself) isn't particularly novel, and the use o!
 f these PODD URIs by other scientists to extend the initial work seems to be the very definition of Linked Data.

 *****

 *****

 I would have been far more supportive of the manuscript if it had described some of their models; explained how the models take advantage of other ontologies (and which); explained which aspects of the data they semantically annotate, why, and how successfully; and how this relates to biodiversity.

The manuscript also feels like it was written by someone extremely familiar with the system, as it uses phraseology that likely only has meaning to them.  For example "External ontologies must currently be mapped into the PODD system to define the expected status of the classes and properties with respect to the basic PODD concepts" - I have read that sentence multiple times and still cannot decypher what it means to "define the expected status of the classes and properties", or what the authors mean by "mapping" external ontologies "into" the PODD system.  Other examples of this include:

"allow scientists to generate datasets before they were certain about the semantic meaning of their work" - what does that mean?

Most of the paragraph beginning "The basic PODD project...", with statements like "annotate images of a plant with a trait in only one case, without annotating other images" and "these horizontal links are not constrained by the basic scientific project structure" (what horizontal links are you referring to, and what are "horizontal links"?)

"any OWL object property that has been defined to link from image acquisition runs to images can be used to define the purpose of an image" - are you saying that images are properties of image acquisition runs?  and that the link between an image acquisition process, and its resulting image, defines the purpose of the image?  How so??

Overall, I thought the manuscript was unnecessarily opaque, didn't emphasise the primary themes of "biodiversity" and "semantics", and over-emphasised many other aspects of their phenotype acquisition and processing workflow.  It's not a bad manuscript, at all, and is likely very well-suited for submission elsewhere (if made slighly less "jargony")!  However, I don't see the fit with this meeting, unless it is extensively re-written.

Best wishes to all the authors!

*Reviewer Name Redacted*

----------------------- REVIEW 3 ---------------------
PAPER: 1
TITLE: Flexible Scientific Data Management for Plant Phenomics Research
AUTHORS: Peter Ansell, Robert Furbank, Kutila Gunasekera, Jianming Guo, David Benn, Gareth Williams and Xavier Sirault

OVERALL EVALUATION: 2 (accept)
REVIEWER'S CONFIDENCE: 3 (medium)
What is your expertise level for the paper?: 4 (good) For this paper do you suggest a long presentation (20 min) ?: 1 (yes) For this paper do you suggest a flashtalk (5 min) ?: 3 (maybe) Is this paper follows LNCS guideline?: 1 (yes) Would you consider the paper to submit an extended paper for publication in a journal?: 1 (yes) Do you think the paper will open or drive a fruitful discussion during the workshop?: 1 (yes) APPROPRIATENESS for the workshop : from 1 (lowest) to 5 (highest): 4 (good)
ORIGINALITY: from 1 (lowest) to 5 (highest): 4 (good) TECHNICAL CORRECTNESS: from 1 (lowest) to 5 (highest): 5 (excellent)

----------- REVIEW -----------
This paper provides an overall view of the Plant Ontology Driven Data repository architecture building to support the High-throughput data generated at the High Resolution Plant Phenomics Centre (Camberra, Australia).
Authors clearly answer major issues of plant Phenomics research applied to the study of phenotype, genotype and environment relationship by providing a flexible and scalable architecture. Moreover the paper shows extensive use of semantic web technologies applied to this field of research. This paper falls completely in the scope of the workshop.

The main contribution of PODD is an OWL ontology which both structure the database model and  check consistency for data. Authors address skillfully the issue of data storage by using robust technologies (e.g. BagIt).

Minor Comment:

I wish appreciate that Authors provide a state of the art for ontology-driven architectures (e.g. iPlant collaborative) and how PODD is distinguised itself from the others.

----------------------- REVIEW 4 ---------------------
PAPER: 1
TITLE: Flexible Scientific Data Management for Plant Phenomics Research
AUTHORS: Peter Ansell, Robert Furbank, Kutila Gunasekera, Jianming Guo, David Benn, Gareth Williams and Xavier Sirault

OVERALL EVALUATION: 2 (accept)
REVIEWER'S CONFIDENCE: 3 (medium)
What is your expertise level for the paper?: 4 (good) For this paper do you suggest a long presentation (20 min) ?: 1 (yes) For this paper do you suggest a flashtalk (5 min) ?: 3 (maybe) Is this paper follows LNCS guideline?: 1 (yes) Would you consider the paper to submit an extended paper for publication in a journal?: 3 (maybe) Do you think the paper will open or drive a fruitful discussion during the workshop?: 1 (yes) APPROPRIATENESS for the workshop : from 1 (lowest) to 5 (highest): 4 (good)
ORIGINALITY: from 1 (lowest) to 5 (highest): 4 (good) TECHNICAL CORRECTNESS: from 1 (lowest) to 5 (highest): 4 (good)

----------- REVIEW -----------
The authors present an intersting approach dedicated to high throughput phenotyping data.
This paper mainly focuses on PODD (Phenomics Ontology Driven Data) properties and the data workflow management. In my opinion, the gap between this paper and state of the art on workflow aspects could be a little more detailled. Overall, the paper is well presented and I believe this work will be relevant for the workshop.

### 2013-04-12 s4biodiv2013/Rebuttals.txt
make changes based on reviewer comments

@@
 The project structure semantics are heavily constrained, while the science semantics are less strictly constrained.
-The science semantics do not attempt to generate high quality upper ontologies, rather they attempt to represent the scientists intentions in a slightly more structured way than otherwise would be possible.
+The science semantics do not set out to generate high quality upper ontologies, rather they attempt to represent the scientists intentions in a slightly more structured way than otherwise would be possible.
+
+ The capture and representation of metadata is highly platform specific, and while the current reference platform, PlantScan, does not generate scientific semantic metadata, other platforms are free to add semantics to metadata without being biased against their use of metadata semantics.
 
@@
 
+ Rebuttal: Unfortunately the use of semantic metadata in scientific publications is very limited in practice.
+Section 6 merely aims to give a quick description of how PODD enables the long-term stability of projects so they can be used as references in static peer reviewed publications, while ontologies are simultaneously updated for use in other projects.
+This enables scientists to have more control than a raw Linked Data platform where there are no guarantees relating to the stability of the system in general.
+
  *****
@@
  I would have been far more supportive of the manuscript if it had described some of their models; explained how the models take advantage of other ontologies (and which); explained which aspects of the data they semantically annotate, why, and how successfully; and how this relates to biodiversity.
+
+ *****
+
+ Rebuttal: There are no measures for the successful use of external ontologies, as the definitions of external ontologies are always in part defined by the community based on both discussions and based on the way they are used.
+The set of ontologies that are currently in use is described in Figure 2.
+
+ *****
 
@@
 
+ *****
+
+ Rebuttal : TODO: Fix this
+
+ *****
+
 Most of the paragraph beginning "The basic PODD project...", with statements like "annotate images of a plant with a trait in only one case, without annotating other images" and "these horizontal links are not constrained by the basic scientific project structure" (what horizontal links are you referring to, and what are "horizontal links"?)
+
+ *****
+
+ Rebuttal : TODO: Fix this
+
+ *****
 
@@
 
+ *****
+
+ Rebuttal : TODO: Fix this
+
+ *****
+
 Overall, I thought the manuscript was unnecessarily opaque, didn't emphasise the primary themes of "biodiversity" and "semantics", and over-emphasised many other aspects of their phenotype acquisition and processing workflow.
@@
 However, I don't see the fit with this meeting, unless it is extensively re-written.
+
+ *****
+
+ Rebuttal : We feel that the reviewer did not recognise the attempt to provide a high level view of the integrated PODD and PlantScan system and how it is being used by HRPPC.
+
+ The reviewer requests that the manuscript use less jargon, but also detail more specifically what ontologies are in use and how they are being practically used.
+
+ We broadly agree with the comments and have made changes to clarify various points.
+
+ *****
 
@@
 
+ *****
+
+ Rebuttal : TODO: Fix this
+
+ *****
+
 ----------------------- REVIEW 4 ---------------------

### 2013-04-12 s4biodiv2013/PODD-S4BIODIV-Submission.tex
make changes based on reviewer comments

@@
 
-The basic PODD project is structured as a tree, which, when used with
-PlantScan\textsuperscript{\texttrademark}, represents the different components
-of a scientific project as
-top-level branches. These include a branch for raw data, along with separate
+In the case of PlantScan\textsuperscript{\texttrademark} a PODD project contains top level branches describing the 
+various parts of a scientific project. These include a branch for raw data, along with separate
 branches for results, analysis, and publications related to the project. In the
-case of raw data, the semantics are not clear and cannot be defined by the
-platforms collecting the data. 
+case of raw data, the semantics are not necessarily clear and are not easily defined by the
+automated platforms collecting the data. 
 % Scientists are able to sparsely attach meaning
@@
 % Assumption to allow extra semantic annotations to be attached in future. 
-For example, they may annotate images of a plant with a trait in only one case,
-without annotating other images. These horizontal links are not constrained by
-the basic scientific project structure, so they are able to entail novel
-meanings independent of the ontology, including untested hypotheses, process
-annotations and meaningful scientific results.
+The scientist may then semantically link the data with results, conclusions, and external ontologies.
+For example, a scientist may annotate the data objects representing images of a plant with a trait 
+that is defined in an external community ontology. They may also annotate the image with a link 
+to a trait that is defined inside of the project, particularly in cases where the trait is novel 
+and not represented in a community ontology.
+%These horizontal links are not constrained by
+%the basic scientific project structure, so they are able to entail novel
+%meanings independent of the ontology, including untested hypotheses, process
+%annotations and meaningful scientific results.
 
 \subsection{Semantic validation}
-PODD validates experiments using independently configurable rules based on OWL
-(Web Ontology Language) schemas. Although PODD currently only supports OWL as a
-rules language, it could be easily extended in other cases to use different
+PODD validates scientific project descriptions using independently configurable constraints based on OWL
+(Web Ontology Language) ontologies. Although PODD currently solely supports OWL for 
+constraint verification, it could be easily extended in other cases to use different
 systems such as N3, RDFS, SPARQL, or SPIN as rules languages \cite{Fuerber2010}.
@@
 all objects having an explicit RDF type, and whether they are consistent with
-the ontologies that they import. The main point that PODD focuses on for
-validation is determining that mappings are consistent with the basic PODD
-predicates and classes, so that they can be used to render both static HTML
-pages and provide HTML based editing facilities. For example, any OWL object
+the ontologies that they import. 
+%The main point that PODD focuses on for
+%validation is determining that mappings are consistent with the basic PODD
+%predicates and classes, so that they can be used to render both static HTML
+%pages and provide HTML based editing facilities. 
+For example, any OWL object
 property that has been defined to link from image acquisition runs to images can
@@
 General scientific properties and phenotype specific properties are defined
-in larger extension ontologies as illustrated in Figure~\ref{poddontologies}.
+in optional extension ontologies as illustrated in Figure~\ref{poddontologies}.
+These are used by scientists to annotate their projects with concepts specific 
+to their field, without requiring other scientists using the system to use phenotype 
+properties to annotate their projects.
 

### 2013-04-15 s4biodiv2013/Rebuttals.txt
rewrite Section 5 p2 to include Crop Ontology example

@@
 
- Rebuttal: There are no measures for the successful use of external ontologies, as the definitions of external ontologies are always in part defined by the community based on both discussions and based on the way they are used.
+ Rebuttal: There are no measures to define the successful use of external ontologies, as the definitions of external ontologies are always in part defined by the community based on both discussions and based on the way they are used.
 The set of ontologies that are currently in use is described in Figure 2.
@@
 
- Rebuttal : TODO: Fix this
+ Rebuttal : Rewrote Section 5 Paragraph 2 to include an example of mapping the Crop Ontology into PODD.
 

### 2013-04-15 s4biodiv2013/PODD-S4BIODIV-Submission.tex
rewrite Section 5 p2 to include Crop Ontology example

@@
 
-External ontologies must currently be mapped into the PODD system to define the
-expected status of the classes and properties with respect to the basic PODD
-concepts. The basic concepts are used to both verify that each PODD project has
-a single valid tree structure, and render HTML pages for both browsing and
-editing. PODD also supports an RDF-based HTTP REST interface that clients may
-use to fetch and update objects.
+Subsets of phenomics ontologies such as the Plant Ontology and the Crop Ontology 
+were mapped into PODD by adding OWL constraints. These constraints enable PODD to
+verify that the use of classes and properties from these ontologies was consistent 
+with the PODD ontology. For example, the Crop Ontology contains a class defining 
+soil as ``Sandy Loam'', giving it the identifier ``0000104''. This was mapped into 
+PODD to define a particular soil sample as being Sandy Loam using the triple:
+$poddSampleSandyLoamSoil\ a\ cropOntology:\_0000104 .$
+
+%External ontologies must currently be mapped into the PODD system to define the
+%expected status of the classes and properties with respect to the basic PODD
+%concepts. The basic concepts are used to both verify that each PODD project has
+%a single valid tree structure, and render HTML pages for both browsing and
+%editing. PODD also supports an RDF-based HTTP REST interface that clients may
+%use to fetch and update objects.
 % without using the HTML pages.

### 2013-04-15 s4biodiv2013/PODD-S4BIODIV-Submission.tex
rewrite Section 2.2 based on review

@@
 essential, in PODD it was necessary to build the system with a relaxed
-ontological vocabulary to allow scientists to generate datasets before they were
-certain about the semantic meaning of their work. In addition, it was necessary
-for PODD to allow scientists to continue to maintain projects containing curated
+ontological vocabulary. This enables scientists to sparsely populate their datasets 
+and sparsely link to community defined upper ontologies as necessary. This 
+allows scientists to continue to maintain projects containing curated
 scientific concepts alongside raw experimental data. The PODD repository was 
 redesigned based on an evaluation of the original software \cite{Li2010} that 
-found it was not able to scale sufficiently to suit the HRPPC needs. The 
-major differences to the software implemented by \cite{Li2010} are that projects 
-are no longer the only supported top object type, and projects are not stored 
-in multiple parts, as that approach was not able to scale as was originally hypothesised.
+found it was not able to scale sufficiently to suit the HRPPC needs due to design 
+and implementation deficiencies. The major differences to the software implemented 
+by \cite{Li2010} are that projects are no longer the only supported top object type, 
+and projects are not stored in multiple parts, as that approach was not able to 
+scale as was originally hypothesised.
 

### 2013-04-15 s4biodiv2013/Rebuttals.txt
rewrite Section 2.2 based on review
address comments for Section 2.2

@@
  Rebuttal : Rewrote Section 5 Paragraph 2 to include an example of mapping the Crop Ontology into PODD.
+Rewrote Section 2.2 Paragraph 1 to clarify that scientists are able to sparsely populate PODD projects with links to community defined upper ontologies.
 
@@
 
- Rebuttal : TODO: Fix this
+ Rebuttal : Rewrote Section 2.2 Paragraph 2 to remove reference to horizontal links. The sparse linking explanation is given in Section 2.2 Paragraph 1.
 

### 2013-04-15 s4biodiv2013/PODD-S4BIODIV-Submission.tex
address comments for Section 2.2

@@
 found it was not able to scale sufficiently to suit the HRPPC needs due to design 
-and implementation deficiencies. The major differences to the software implemented 
+and implementation deficiencies. The major design differences to the software implemented 
 by \cite{Li2010} are that projects are no longer the only supported top object type, 
@@
 
-In the case of PlantScan\textsuperscript{\texttrademark} a PODD project contains top level branches describing the 
+A PODD project in PlantScan\textsuperscript{\texttrademark} contains top level branches describing the 
 various parts of a scientific project. These include a branch for raw data, along with separate
@@
 % Assumption to allow extra semantic annotations to be attached in future. 
-The scientist may then semantically link the data with results, conclusions, and external ontologies.
-For example, a scientist may annotate the data objects representing images of a plant with a trait 
-that is defined in an external community ontology. They may also annotate the image with a link 
-to a trait that is defined inside of the project, particularly in cases where the trait is novel 
+The scientist may later semantically link the data with results, conclusions, and external ontologies.
+For example, a scientist may annotate the data objects representing images of a plant with a link to a trait 
+that is defined in the Plant Ontology. They may also annotate the image with a link 
+to a trait that is defined inside of the project, such as when the trait is novel 
 and not represented in a community ontology.

### 2013-04-15 s4biodiv2013/Rebuttals.txt
update figure

@@
 
- Rebuttal : We feel that the reviewer did not recognise the attempt to provide a high level view of the integrated PODD and PlantScan system and how it is being used by HRPPC.
+ Rebuttal : We feel that the reviewer did not recognise the attempt to provide a high level view of the integrated PODD and PlantScan system and how it is being used by HRPPC to process plant phenomics data (Section 2).
+Plant phenomics is a legitimate subset of biodiversity.
+It highlights differences between, and inside, multiple plant species based on feature comparisons.
+This paper provides a description of the high level use of ontologies to structure and enrich a semantic data model based on RDF and OWL.
+OWL is used in PODD for both syntactic validation and sparse semantic enrichment purposes.
 
- The reviewer requests that the manuscript use less jargon, but also detail more specifically what ontologies are in use and how they are being practically used.
+ The reviewer requests that the manuscript use less jargon, but also detail more specifically what ontologies are in use and how they are being practically used, which entails using more jargon related to those specific ontologies.
 

### 2013-04-15 s4biodiv2013/PODD-S4BIODIV-Submission.tex
update figure

@@
 photosynthetic performance, yield must be stored and correlated with genotypes. 
-These factors provide evidence of genetic variation in natural and derived genetic populations (e.g.
-germplasm collections, association genetic panels, recombinant inbred lines).
-They also enable a deeper understanding the dynamic relationship between phenotype, genotype and environment which is necessary to continue delivering the increase in productivity necessary for feeding the world.
+These factors provide evidence of genetic variation in natural and derived genetic 
+populations (e.g. germplasm collections, association genetic panels, recombinant inbred 
+lines). They also enable a deeper understanding of the dynamic relationship between 
+phenotype, genotype and environment which is necessary to continue delivering the 
+increase in productivity necessary for feeding the world.
 
@@
 of phenomics platforms must be combined with metadata explaining how the raw
-data was collected. This combination of raw data and metadata are then delivered to a range of analysis pipelines, which transform the raw data into aggregated
+data was collected. This combination of raw data and metadata are then delivered 
+to a range of analysis pipelines, which transform the raw data into aggregated
 multi-phase datasets, each phase representing a new aggregation or inference
@@
 multi-dimensional data into information which is conceptually interpretable by a
-human being, i.e. new knowledge. The additional metadata describing the steps taken are recorded to give context to the data. 
+human being, i.e. new knowledge. The additional metadata describing the steps taken 
+are recorded to give context to the data. 
 % These links enable scientists to generate novel
@@
 These are used by scientists to annotate their projects with concepts specific 
-to their field, without requiring other scientists using the system to use phenotype 
-properties to annotate their projects.
+to their field, without requiring other scientists using the same PODD installation 
+to use phenotype properties to annotate their projects.
 

### 2013-04-15 s4biodiv2013/Rebuttals.txt

@@
 
- Rebuttal : TODO: Fix this
+ Rebuttal : Rewrote this sentence so it refers to provenance, not properties.
 
@@
 
- The reviewer requests that the manuscript use less jargon, but also detail more specifically what ontologies are in use and how they are being practically used, which entails using more jargon related to those specific ontologies.
+ The reviewer requests that the manuscript use less jargon, while also requiring more detail about what ontologies are in use (See Figure and how they are being practically used, which entails using more jargon related to those specific ontologies.
 

### 2013-04-15 s4biodiv2013/PODD-S4BIODIV-Submission.tex

@@
 %pages and provide HTML based editing facilities. 
-For example, any OWL object
-property that has been defined to link from image acquisition runs to images can
-be used to define the purpose of an image. This may include any combination of
-the defined properties, without requiring all of the properties to be used.
+For example, any OWL object property that has been defined to link from image 
+acquisition runs to images defines the provenance of an image.
+
+%This may include 
+%any combination of the defined properties, without requiring all of the properties 
+%to be used.
 % The parent object types and properties used by PODD to establish basic ground
 % rules for rendering and validation are defined in a small base schema ontology.
+
 General scientific properties and phenotype specific properties are defined

### 2013-04-15 s4biodiv2013/PODD-S4BIODIV-Submission.tex
re-export the figure in different proportions
back down to 8 pages

@@
 \begin{center}
- \includegraphics[width=12cm,keepaspectratio=true]{plantscan-workflow-figure1.png}
+ \includegraphics[width=11.5cm,keepaspectratio=true]{podd-workflow-figure1.png}
  % plantscan-workflow-figure1.png: 598x330 pixel, 72dpi, 21.10x11.64 cm, bb=0 0
@@
 
-DAP is the visible part of the system. Other than standard web servers,
-databases and application servers, additional required server-side
-infrastructure includes petabyte-level, multi-decade, backed-up disk storage,
-integrity checking, and automated migration of data collections to and from tape
-storage.
+%DAP is the visible part of the system. Other than standard web servers,
+%databases and application servers, additional required server-side
+%infrastructure includes petabyte-level, multi-decade, backed-up disk storage,
+%integrity checking, and automated migration of data collections to and from tape
+%storage.
 
